### 2019-09-14 Draft/draft.tex
QIP version

\documentclass[11pt,letterpaper]{article}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
                                      {1ex \@plus1ex \@minus.2ex}%
                                      {-1em}%
                                      {\normalfont\normalsize\bfseries}}
\makeatother

%%%%%%%%%%%%%%%%%%%%%
%  P A C K A G E S  %
%%%%%%%%%%%%%%%%%%%%%

% Authors
\usepackage{authblk}

% Page margins
\usepackage[margin=1in]{geometry}

% Nicer math font
\usepackage{mathpazo}

% More fancy lists
\usepackage{enumerate}

% Microtype
\usepackage{microtype}

% TikZ
\usepackage{tikz}
%\usetikzlibrary{calc,shapes.geometric}
\usetikzlibrary{backgrounds,fit,decorations.pathreplacing,calc}

% Highlights
\usepackage{soul}

% Figure
\usepackage{float}

% Hypertext package
\usepackage[colorlinks = true]{hyperref}
% Title and authors
%\hypersetup{
%  pdftitle = {},
%  pdfauthor = {}
%}
% Color definitions
\definecolor{darkred}  {rgb}{0.5,0,0}
\definecolor{darkblue} {rgb}{0,0,0.5}
\definecolor{darkgreen}{rgb}{0,0.5,0}
% Color links
\hypersetup{
  urlcolor   = blue,         % color of external links
  linkcolor  = darkblue,     % color of internal links
  citecolor  = darkgreen,    % color of links to bibliography
  filecolor  = darkred       % color of file links
}

% AMS
\usepackage{amsmath,amssymb,amsfonts,amsthm,amstext}
\usepackage{bbm}

%% Restating theorems
%\usepackage{thm-restate}

% Powerful macros
\usepackage{etoolbox}

% Fixes for amsmath
\usepackage{mathtools}
\mathtoolsset{centercolon}
\makeatletter
\protected\def\tikz@nonactivecolon{\ifmmode\mathrel{\mathop\ordinarycolon}\else:\fi}
\makeatother

% Daw boxes
\usepackage{tcolorbox}

% Code
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}

% Clever references

\usepackage{cleveref}%[nameinlink]

\crefname{lemma}{Lemma}{Lemmas}
\crefname{proposition}{Proposition}{Propositions}
\crefname{definition}{Definition}{Definitions}
\crefname{theorem}{Theorem}{Theorems}
\crefname{conjecture}{Conjecture}{Conjectures}
\crefname{corollary}{Corollary}{Corollaries}
\crefname{claim}{Claim}{Claims}
\crefname{section}{Section}{Sections}
\crefname{appendix}{Appendix}{Appendices}
\crefname{figure}{Fig.}{Figs.}
\crefname{table}{Table}{Tables}

% \crefname{algorithm}{Algorithm}{Algorithms}

% IEEE tools
\usepackage[retainorgcmds]{IEEEtrantools}

% table of contents
\usepackage{tocloft}

% Table with multi-row
\usepackage{multirow}

% TikZ
\usepackage{tikz}	
\usetikzlibrary{backgrounds,fit,decorations.pathreplacing}

%%%%%%%%%%%%%%%%%%%%%%%%%
%  N E W C O M M A N D  %
%%%%%%%%%%%%%%%%%%%%%%%%%

% Standard quantum notation

\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\braket}[2]{\langle#1|#2\rangle}
\newcommand{\ketbra}[2]{|#1\rangle\langle#2|}
\newcommand{\proj}[1]{|#1\rangle\langle#1|}

\newcommand{\x}{\otimes}
\newcommand{\xp}[1]{^{\otimes #1}}
\newcommand{\op}{\oplus}

\newcommand{\ct}{^{\dagger}}
\newcommand{\tp}{^\intercal}

% Linear algebra

%\newcommand{\1}{\mathbb{1}} % identity matrix
%\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\End}{End}
%\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Lin}{L} % all linear maps
\newcommand{\Mat}[1]{\mathrm{M}(#1)} % all matrices
%\newcommand{\Mat}[1]{\mathrm{M}_{#1}(\C)}

% Paired delimiters

\DeclarePairedDelimiter{\set}{\lbrace}{\rbrace}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\DeclarePairedDelimiter{\cnorm}{\lvert}{\rvert}
\DeclarePairedDelimiter{\of}{\lparen}{\rparen}
\DeclarePairedDelimiter{\sof}{\lbrack}{\rbrack}
\DeclarePairedDelimiter{\ip}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

% Operators

\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}
\DeclareMathOperator{\vc}{vec}
\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\spec}{spec}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\hook}{hook}
\DeclareMathOperator*{\E}{\mathbb{E}}
\DeclareMathOperator{\supp}{supp}

% Sets

\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\calH}{\mathcal{H}}
\newcommand{\calX}{\mathcal{X}}
\newcommand{\calY}{\mathcal{Y}}
\newcommand{\calA}{\mathcal{A}}
\newcommand{\calB}{\mathcal{B}}

% Group
\newcommand{\Zd}{\Z_d^{\times}}

% Identity operator
\newcommand{\1}{\mathbb{1}}

% Pauli Group
\newcommand{\Pg}{\mathcal{P}}
\newcommand{\J}{\mathcal{J}}

% Special notation

\newcommand{\CHSH}{CHSH^{(d)}}
\newcommand{\MS}{MS}
\newcommand{\EXT}{EXT}
\newcommand{\LS}{LS}
\newcommand{\COMM}{COMM}
\newcommand{\EPR}[1]{\Sigma^{(#1)}}
\newcommand{\paulix}{\sigma_x}
\newcommand{\pauliz}{\sigma_z}
\newcommand{\tO}{\tilde{O}}
\newcommand{\tP}{\tilde{P}}
\newcommand{\tQ}{\tilde{Q}}
\newcommand{\tM}{\tilde{M}}
\newcommand{\tN}{\tilde{N}}
\newcommand{\tA}{\tilde{A}}
\newcommand{\tB}{\tilde{B}}
\newcommand{\tX}{\tilde{X}}
\newcommand{\tZ}{\tilde{Z}}
\newcommand{\tU}{\tilde{U}}
\newcommand{\tW}{\tilde{W}}
\newcommand{\tx}{\tilde{x}}
\newcommand{\tpsi}{\tilde{\psi}}
\newcommand{\tri}{\Delta}
\newcommand{\lB}{\overline{B}}
\newcommand{\dr}[1]{d^{(#1)}}
\newcommand{\nr}{n(r)}
\newcommand{\mr}{m(r)}
\newcommand{\ux}{\underline{x}}
\newcommand{\uc}{\underline{c}}
\newcommand{\ua}{\underline{a}}
\newcommand{\ub}{\underline{b}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\ba}{\pmb{a}}
\newcommand{\bb}{\pmb{b}}
\newcommand{\bc}{\pmb{c}}
\newcommand{\bS}{\mathbbmtt{S}}

% Probabilities
\newcommand{\pr}[2]{P(#1|#2)}
\newcommand{\pa}[2]{P_A(#1|#2)}
\newcommand{\pb}[2]{P_B(#1|#2)}
\newcommand{\tpr}[2]{\tilde{P}(#1|#2)}

% Bell Ineqaulities
\newcommand{\I}{\mathcal{I}}

% Square root of epsilon
\newcommand{\ep}{\epsilon}
\newcommand{\se}{\sqrt{\epsilon}}
\newcommand{\qe}{\epsilon^{1/4}}
\newcommand{\sd}{\sqrt{d}}
\newcommand{\sr}{\sqrt{r}}
\newcommand{\qr}{r^{1/4}}
\newcommand{\qd}{d^{1/4}}

% Approximately equatli
\newcommand{\appd}[1]{\simeq_{#1}}

% Comments
\def\carl#1{{\color{blue} #1}}
\newcommand{\hf}[1]{\textcolor{red}{#1}}
\newcommand{\hfc}[1]{\textcolor{red}{#1 -H.F.}}

%%%%%%%%%%%%%%%%%%%%%%%%%
%  N E W T H E O R E M  %
%%%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{claim}[theorem]{Claim}
\newtheorem*{conjecture*}{Conjecture}
\newtheorem*{problem}{Problem}
\newtheorem*{example}{Example}

\theoremstyle{definition}
\newtheorem*{remark}{Remark}

%%%%%%%%%%%%%%%%
%   Document   %
%%%%%%%%%%%%%%%%

\begin{document}

\title{Constant-sized correlations are sufficient to 
robustly self-test maximally entangled states with unbounded dimension}

\author[1]{Honghao Fu}

\renewcommand\Affilfont{\itshape\small}

\affil[1]{Department of Computer Science, Institute for Advanced Computer Studies, and Joint Center for Quantum \break Information and Computer Science, University of Maryland, College Park, MD 20742, USA}
%\affil[2]{National Institute of Standards and Technology, 100 Bureau Dr., Gaithersbug, MD 20899, USA}
\maketitle

\begin{abstract}
	We show that for any prime odd integer $d$, there exists a correlation of size $\Theta(r)$ 
	that can robustly self-test a maximally entangled state of dimension $4d-4$,
	where $r$ is the smallest multiplicative generator of $\Zd$.
	The construction of the correlation uses the embedding procedure
	proposed by Slofstra (\textit{Forum of Mathematics, Pi.} Vol.~$7$ Cambridge University Press, $2019$).
	Since there are 
	infinitely many prime numbers whose smallest multiplicative generator
	is at most $5$ (\textit{The Mathematical Intelligencer} $10.4$ ($1988$)), 
	our result implies that constant-sized correlations are sufficient for robust self-testing of maximally entangled states
	with unbounded local dimensions.
\end{abstract}
%========================================
\section{Introduction}
\label{sec:intro}
%========================================
The certification of a quantum device is an important building block
for many quantum information processing tasks, especially
when the devices are provided by some untrusted vendor.
We would like such certification to be done based solely on
the observed measurement statistics and with the only assumption that
any local device cannot communicate their own
inputs to the other local devices. 
The measurement statistics are referred to as correlations.

It has been shown that certain quantum correlations require the distant parties to share
a particular entangled state up to some local isometry. 
This phenomenon is referred to as self-testing.  
If we can further prove that, when the observed correlation deviates 
from the ideal one to some extent, 
the shared state is close to the ideal state up to some local isometry,
then we say the self-test is robust.

Self-testing is powerful because only classical interactions are performed.
Hence, it allows a classical party to delegate quantum computations to some untrusted service provider
and verify that the computations are performed
honestly and correctly \cite{ruv2013,coladan2017verifier}.
It also becomes a critical component of the security proofs of device-independent quantum cryptographic protocols
\cite{mayersyao,vv2014,miller2017,fu2018,eat2018}.
Since the setting of interactive proof systems is similar to a self-test, self-testing arguments also help to prove
lower bounds of the computational power of entanglement-assisted
multiprover interactive proof systems \cite{fitzsimons2019, neexp}.

The case of robust self-testing of the EPR pair, 
\begin{align*}
    \ket{EPR} = \frac{1}{\sqrt{2}}(\ket{00}+\ket{11}),
\end{align*}
is fully understood.
The techniques for this case are 
first introduced in \cite{mckague2012}, then improved in \cite{bamps2015}. 
Robust self-testing of partially entangled qubits is proved later in
\cite{yang2013}, which also proposed an generalized swap-isometry
that can be used to robustly self-test large dimensional maximally entangled states.
Robust self-testings of tensor product of maximally and partially entangled qubits 
are proved in \cite{coladan2016parallel,natarajan2017,lowdegree},
with the last one being the one with the smallest input and output alphabets.
Self-testing of all pure bipartite entangled states is proved in \cite{coladan2017all}.
The authors in \cite{coladan2017all} constructed a family of correlations such
that an entangled state of any local dimension has
a self-testing correlation in this family.
% One can also self-test many copies of the $2$-dimensional EPR
% pair in parallel \cite{mckague2016,natarajan2017,coladan2016parallel}.
% The most efficient way uses $O(\log n)$ bits question and
% answers to self-test $n$ copies of $2$-dimensional EPR
% pair \cite{lowdegree}.
% Self-testing general $d$-dimensional EPR pairs is a harder task.
% Coladangelo, Goh and Scarani has shown that all pure bipartite entangled states can be self-tested \cite{coladan2017all}
To generate a correlation in the family, Alice and Bob
use constant-sized input alphabets but $\Theta(d)$-sized output alphabets, 
where $d$ is the local dimension of the entangled state. 
This is the only result about self-testing the general $d$-dimensional entangled state so far.
However, proving the robustness of this family of correlations remains an open problem.

Our work aims at answering two open questions.
The first one is whether maximally entangled states with large local dimension
can be self-tested with constant-sized input and output alphabets, 
which are independent of the local dimension of the self-tested state.
For comparisons, all the correlations used in the results listed above
have either input or output alphabets with sizes dependent on the
local dimension of the entangled state.
The second one is how to robustly self-test maximally entangled states
with large local dimension and no tensor product structure. 
We give an affirmative answers to the first question and
make significant progress in answering the second question
by proving the following theorem.
\begin{theorem}[Informal]
\label{thm:inf}
	There exists an infinite-sized set $D$ of odd prime numbers such that, for any $d \in D$, 
	the maximally entangled state of local dimension $4(d-1)$ can be robustly self-tested 
	with constant-sized question and answer alphabets.
\end{theorem}
The formal statement of this theorem is given in \cref{thm:self-test} and \cref{thm:infty}.

The set $D$ is, specifically, the set of all 
odd prime numbers whose smallest primitive root 
is either $2$, $3$, or $5$, 
We say that $r$ is a primitive root of $d$ if $r$ is the 
multiplicative generator of the group $\Zd$.
It has been shown that $D$ has infinitely many elements \cite{murty1988}.

To prove \cref{thm:inf}, we construct a correlation with $\Theta(r)$-sized
input and constant-sized output alphabets
for each odd prime number $d$ whose smallest primitive root is $r$.
This correlation is denoted by $\fC(d,r)$.
Note that although the size of the input or output alphabet of $\fC(d,r)$ does not depend on $d$, 
the distributions in $\fC(d,r)$ do. 
Once we prove $\fC(d,r)$ can robustly self-test a maximally entangled
state of dimension $4(d-1)$, we apply it to all $d \in D$
and prove \cref{thm:inf}.
%Then the self-testing property of this correlation is 
%given in the following theorem.
%\begin{theorem}[Informal]
%\label{thm:pr_2}
%	All maximally entangled state with local dimension $d-1$, where $d$ is prime and has
%	primitive root $r \in \{2, 3, 5\}$, can be self-tested by a constant-sized correlation.
%\end{theorem}

The correlation $\fC(d,r)$ contains a winning correlation of a specially-designed binary linear system game.
In a general linear system game, Alice is given a random equation 
from a linear system and she
should give an assignment to each variable of the equation.
Bob is asked to assign a value to a random variable of that equation.
They win the game if Alice's assignment satisfies the equation and Bob's
assignment to the variable matches Alice's assignment.
A widely-used and thoroughly-studied example of a linear system game is
the Magic Square game \cite{magic_square}.
In this game, the linear system is
\begin{align*}
    &x_1 + x_2 + x_3 = 0 && x_4 + x_5 + x_6 = 0 &&
    x_7 + x_8 + x_9 = 0 \\
    &x_1 + x_4 + x_7 = 0 && x_2 + x_5 + x_8 = 1 &&
    x_3 + x_6 + x_9 = 0.
\end{align*}
Using two copies of $\ket{EPR}$
the Magic square game can be won with certainty.
It has been shown that if a strategy that can win this game almost perfectly, 
the shared state must be close to $\ket{EPR}^{\x 2}$ up to some local isometry \cite{wu2016}, 
so we can say that the Magic Square game can robustly self-test $\ket{EPR}^{\x 2}$.
The key observation that leads to the self-testing proof is that 
a winning strategy must contain binary observables $X$ and $Z$ that 
satisfy the anti-commutation relation 
\begin{align*}
    X Z = - Z X.
\end{align*}
Intuitively, we can think of the Magic Square game as a way to enforce the
anti-commutation relation \cite{coladan2017}.

Our linear system game is designed to enforce the relation 
\begin{align*}
    U O U\ct = O^r,
\end{align*}
for unitaries $U$ and $O$, and some integer $r$.
It implies that the winning strategy contains binary observables
$M_1, M_2, M_3, M_4$ on Alice's side and 
$N_1, N_2, N_3, N_4$ on Bob's side such that 
\begin{align*}
    U = M_3M_4 = N_3N_4 && O = M_1M_2 = N_1N_2.
\end{align*}
This relation implies the eigenvalues of $O$ must be $\omega_d^j = e^{ij2\pi/d}$ for some $j \neq 0$ and some prime $d$ whose primitive root is $r$. 
It also implies that $U$ is a permutation of the eigenbasis of $O$.
Winning this linear system game implies that Alice and Bob share a state
$\ket{\psi}$ such that
\begin{align*}
    O \x O \ket{\psi} = \ket{\psi}, && 
    U \x U \ket{\psi} = \ket{\psi}.
\end{align*}
These two relations are important for the self-testing argument.
The relation $O \x O \ket{\psi} = \ket{\psi}$ implies that $\ket{\psi}$
is spanned by tensor products of states in the eigenbasis of $O$.
The relation $U \x U \ket{\psi} = \ket{\psi}$ implies that $\ket{\psi}$ is invariant
under the permutation induced by $U$. 
At this point, the eigenvalues of $O$ are unknown, so
certifying the eigenvalues of $O$ requires another test, 
which will be introduced later.

The special feature of our linear system game is that the sizes
of the question sets for Alice and Bob are of order $\Theta(r)$,
and the answer sets are of size $\Theta(1)$.
Restricting $r \leq 5$ as in \cref{thm:inf} implies that our linear system 
game only uses constant-sized question sets and constant-sized answer sets.

% The number of variables and the number of equations of the linear system game depend linearly on $r$,
% which makes our correlation of size $\Theta(r^2)$.
% Unlike previous self-testing proofs, which try to enforce the anti-commutation relation
% $XZ = \omega_d^{-1} ZX$ between Pauli operators of dimension $d \geq 2$, 
% our approach is more succinct in terms of the size of the correlation but at a cost that
% it only works for odd prime $d$.
% The reason why we focus on this relation is that if we can certify that $O$ has one eigenvalue $\omega_d$
% for some $d$ with primitive $r$,
% then it guarantees that $O$ has eigenvalues $\omega_d^j$ for $j = 1 \dots d-1$ and $U$ only
% permutes the eigenvectors of $O$ in certain way.
% After knowing the eigenvalues of $O$, we can use infer the structure of the shared state $\ket{\psi}$
% based on the two relation of $\ket{\psi}$ imposed by the linear system game --
% $U_A \x U_B \ket{\psi} = \ket{\psi}$ and $O_A \x O_B \ket{\psi} = \ket{\psi}$.
% The two relations imply that $\ket{\psi}$ is spanned by eigenvectors of $O$ and its
% Schmidt coefficient follows some pattern that is invariant under the permutation by $U$.
% At this point, we can use the generalized swap isometry \cite{yang2013} to extract a copy of the 
% $(d-1)$-dimensional maximally entangled state. Note that our robust self-testing proof 
% follows a different line of arguments from what was originally
% proposed in Ref.~\cite{yang2013}.

The remaining question is how to certify eigenvalues of $O$, so we
design a nonlocal test called the extended weighted CHSH test.
% Before going to the answer,
% we remark that 
% the other specialty of the linear system game is that it forces Alice and Bob to have binary observables
% whose products are the special unitaries $U$ and $O$, because the answers to the questions of the 
% linear system games are binary. 
% The idea of decompose the unitaries of arbitrary order into binary observables was also first introduced by Slofstra \cite{slofstra2017}.
% We use a different way to decompose the unitaries, which can be seen as the inverse of Jordan's lemma, 
% The fact that Alice and Bob only have binary observables make it easier for us 
% to compose the linear system game with another nonlocal test, which we call the extended weighted CHSH test.
The difference between a nonlocal game and a nonlocal test will be discussed later.
The combination of a linear system game with 
a nonlocal test in the construction of $\fC(d,r)$ 
is a new way to use a linear system game for robust self-testing.
It distinguishes our approach from all the previous ones.

The extended weighted CHSH test is designed based on the $\alpha$-weighted CHSH inequality, 
\begin{align*}
    \I_\alpha = \alpha\ip{M_1N_1}+\alpha\ip{M_1N_2} + \ip{M_2N_1} - \ip{M_2N_2}\leq 2|\alpha|,
\end{align*}
where $M_x$ and $N_y$ for $x,y \in \{1, 2\}$ are some binary observables.
Define $\mu = \arctan(1/\alpha)$,
\begin{align*}
    \pauliz = \ketbra{0}{0} - \ketbra{1}{1} && \paulix = \ketbra{0}{1} + \ketbra{1}{0},
\end{align*}
which are the Pauli-Z and Pauli-X operators.
This inequality is maximally violated when the shared state is $\ket{EPR}$,
Alice's binary observables are
\begin{align*}
    \tM_1 = \pauliz && \tM_2 = \paulix,
\end{align*}
and Bob's binary observables are
\begin{align*}
    \tN_1 = \cos(\mu) \pauliz+ \sin(\mu) \paulix  && \tN_2 = \cos(\mu) \pauliz - \sin(\mu) \paulix。
\end{align*}
We first prove that the maximal violation of $\I_\alpha$ can robustly self-test the shared
state $\ket{\psi}$ and the observables $\tM_1, \tM_2$ and $\tN_1, \tN_2$ in \cref{sec:ext_chsh}.
Then we notice that 
\begin{align*}
    \tN_1\tN_2 = \cos(2\mu) \1  - \sin(2 \mu) \pauliz \paulix 
\end{align*}
has eigenvalues $e^{i 2\mu}$ and $e^{-i 2\mu}$.

Therefore, in the extended weighted CHSH test, we set $\alpha = \cot( -\pi/d )$
for some odd prime $d$ whose primitive root is $r$,
and ask Alice and Bob
to reuse the observables $M_1, M_2$ and $N_1, N_2$ defined in $\LS(r)$. Then we can certify that
the unitary $O = M_1M_2 = N_1N_2$ has eigenvalues $e^{i2\pi/d}$ and $e^{-i2\pi/d}$.
Combining this observation with the fact that $O$ satisfy the relation $UOU\ct = O^r$,
we can conclude that all the eigenvalues of $O$ are $\{ e^{ij 2\pi /d} \}_{j=1}^{d-1}$,
and the eigenspaces of all the eigenvalues have the same cardinality.

At this point, we have the explicit structure of $O$, so we 
can use the relations $U \x U \ket{\psi} = \ket{\psi}$ and $ O \x O \ket{\psi} = \ket{\psi}$
to prove the correlation $\fC(r, d)$ can robustly self-test 
a maximally entangled state of dimension $4(d-1)$,
\begin{align*}
    \left(\frac{1}{\sqrt{d}} \sum_{j=1}^{d-1}\ket{j}\ket{d-j}\right)\x 
    \ket{EPR}^{\x 2}.
\end{align*}

In the proof, we demonstrate a new way to apply the generalized swap-isometry
proposed in \cite{yang2013} to robustly self-test a maximally entangled state of large local
dimension. This is the first application of the generalized swap-isometry in the robust self-testing proof.

Our contributions are:
\begin{enumerate}
\item We identify a family of correlations with constant-sized input and output alphabets that can be used to robustly self-test
maximally entangled states with unbounded local dimensions. Before our work, the most efficient way 
was to use $O(\log d)$-sized input and $O(\log \log d)$-sized output alphabets to robustly self-test
the maximally entangled state of dimension $d = 2^n$ \cite{lowdegree}.

\item Our result gives us deeper understandings of the set of quantum correlations.
In previous works that proved the non-closeness of the set of quantum correlations \cite{slofstra2017, slofstra2018}, 
lower bounds of the entanglement required to achieve certain correlations are identified.
We discover a large family of correlations such that we know exactly the entangled state
required to achieve each of the correlations in this family. 
\end{enumerate}

\textbf{Structure of the paper}.
We start with notations and background information in \cref{sec:prelim}.
Since the correlation we designed can win a special linear system game and satisfy
an extended weighted CHSH test, we introduce the linear system game
in \cref{sec:lsg} and the extended weighted CHSH test in \cref{sec:ext_chsh}. 
Then we give the correlation $\fC(d,r)$ and the ideal strategy to generate it in \cref{sec:full_test}.
In the end, we give the robust self-testing proof in \cref{sec:main}.

%========================================
\section{Preliminaries and notations}
\label{sec:prelim}
%========================================
%-----------------------------------
\subsection{Notations and mathematical background}
%-----------------------------------
The maximally entangled state of local dimension $d-1$ for some odd prime $d$ that we self-test is denoted by
\begin{align}
\ket{\EPR{d-1}} = \frac{1}{\sqrt{d-1}} \sum_{j = 1}^{d-1} \ket{j}\ket{d-j}.
\end{align}
The superscript $(d-1)$ stresses the dimension of the local Hilbert space and we follow this convention through this paper.
The $d$-th root of unity is denoted by $\omega_d:=e^{2\pi i/d}$.

In the qubit case, the EPR pair is denoted by 
\begin{align}
	\ket{EPR} = \frac{1}{\sqrt{2}}(\ket{00} + \ket{11}),
\end{align}
and the Pauli operators are denoted by
\begin{align}
	\paulix = \ketbra{0}{1}+\ketbra{1}{0} && \pauliz = \ketbra{0}{0} - \ketbra{1}{1}.
\end{align}
Note that we only work with finite Hilbert spaces in our work.

For quantum states and operators on different Hilbert spaces, we use subscripts to label them.
For example, $O_A$ is an operator on Alice's side and $\ket{0}_{B}$ is a state on Bob's side. 
The only exception is when we label projectors used in a quantum strategy defined below. In this case, the subscript 
is the input and the superscript is the output. For example, $M_x^a$ is Alice's projector for input $x$ and output $a$.
Operators defined by the projectors follow the same convention, for example, $M_x := M_x^0 - M_x^1$ is 
a binary observable.
If we add tilde to an operator, then this operator is fully characterized and belongs to some ideal strategy.
Otherwise, the operator is general and the only assumption is that it is finite-dimensional.

We use the following notation for the closeness between quantum states.
\begin{align}
	\ket{u} \appd{\epsilon} \ket{v} \iff \norm{\ket{u} - \ket{v}} \leq \epsilon. 
\end{align}
Similarly, for numbers $a,b \in \C$, the closeness is denoted by
$a \appd{\epsilon} b \iff \cnorm{a-b} \leq \epsilon$.

We use some basic number theory in our work. 
\begin{definition}
A primitive root of a prime number
$d$ is an integer $r$ such that $r$ has multiplicative order $d-1$
in the ring of integers modulo $d$, $\mathbb{Z}_d$.
\end{definition}
In other words,
$r$ is a multiplicative generator of $\Zd$.

We also need some basic notions of group presentation and group
representation, so 
We introduce them below.
For more contexts, please refer to \cite{dummit2004}.

Let $S$ be a set of letters. We denote by $\mathcal{F}(S)$ the \emph{free group generated by $S$}, 
which consists of all finite words made from $\set{s, s^{−1}| s \in S}$ such that no $ss^{−1}$ or $s^{−1}s$ appears as a substring for any $s$,
where $s^{-1}$ denotes the inverse of $s$. 
The group law is given by concatenation and cancellation.
For a more formal treatment, we refer to Section $6.3$ of \cite{dummit2004}.

\begin{definition}[Group presentation]
    Given a set $S$, let $\mathcal{F}(S)$ be the free group generated by $S$ and $R$ be a subset of $\mathcal{F}(S)$.
    Then $\ip{S:R}$ denotes the quotient of $\mathcal{F}(S)$
    by the normal subgroup generated by $R$.
    If the group $G$ is isomorphic to $\ip{S:R}$,
    then we say $G$ 
    has a presentation $\ip{S:R}$. 
\end{definition}
If both sets $S$ and $R$ are finite, then we say the group $G = \ip{S:R}$ is \emph{finitely-presentable}.
The elements of $S$ are the \emph{generators} and the elements
of $R$ are the \emph{relations}.
When we refer to an equation
of the form $a = b$ with $a, b \in \mathcal{F} ( S )$ as a ``relation,'' 
it is understood 
that we mean the element $ab^{-1} 
\in R$.
There are two types of relations that we will be working with.
Considering a subset of generators $\{ s_j \}_{j=1}^n \subset S$ such that $s_{j_1} \neq s_{j_2}$ if $j_1 \neq j_2$,
the relations of the form $\Pi_{j=1}^n s_j  = e$, where $e$ is the identity element, are called 
\emph{linear relations};
and the relations of the form
$s_i s_j s_i^{-1} = s_k$ for some $i \neq j$ are called \emph{conjugacy relations}.

There are two special types of groups that we will work with in \cref{sec:lsg}. We give the definitions below.
\begin{definition}
    \label{def:spe_conj_grp}
    Suppose $C \subset [n] \times [n] \times [n]$, where $[n] = \{1,2 \dots n\}$. 
    Let
    \begin{align*}
        \Gamma = \ip{\{s_i\}_{i \in [n]}  :
        \{s_i^2=e\}_{i\in [n]} \cup 
        \{ s_i s_j s_i = s_k\}_{(i,j,k)\in C}}.
    \end{align*}
    We say a group is a conjugacy group if it has a presentation of this form. 
\end{definition}
A conjugacy group is a special case of what Slofstra defined as 
a linear-plus-conjugacy group (Definition $26$ of \cite{slofstra2017}). 
The difference is that a conjugacy group does not have linear relations. 
The next definition follows the 
definition of a nice linear-plus-conjugacy group (Definition $28$ of \cite{slofstra2017}).
\begin{definition}
    \label{def:nice_conj_grp}
    Suppose $C \subset [n] \times [n] \times [n]$, where $[n] = \{1,2 \dots n\}$. 
    Let 
    \begin{align*}
        \Gamma = \ip{\{s_i\}_{i \in [n]} \cup \{s_{jk}\}_{(i,j,k)\in C} :
        \{s_i^2=e\}_{i\in [n]} \cup 
        \{s_{jk}^2 = e, s_is_js_i = s_k, s_{jk}s_js_k = e\}_{(i,j,k)\in C}}.
    \end{align*}
    We say a group is a nice conjugacy group if it has a presentation of this form. 
\end{definition}
The relation $s_{jk} s_js_k = e$ implies that $s_j$ commutes with $s_k$
because
\begin{align*}
    [s_j, s_k] = s_js_ks_js_k = s_{jk}^2 = e.
\end{align*}

Next we give the definition of a group representation. 
\begin{definition}
    A representation of a group $G$ on the finite Hilbert space $\calH$ is a homomorphism from $G$ to the group of invertible linear operators on $\calH$. 
\end{definition}
The last concept of group theory that we need is a group embedding.
\begin{definition}[Group embedding]
    An embedding of $G$ into $K$ is an injective group homomorphism: $\phi: G \rightarrow K$.
\end{definition}
%-----------------------------------------
\subsection{Nonlocal games}
%-----------------------------------------
In a nonlocal game, there are two players, Alice and Bob and each of them is requested
to give an answer for a randomly chosen question. 
We denote Alice's question set by $\calX$ and her family of answer sets by $\{\calA_x\}_x$
for each $x \in \calX$. Similarly,
Bob's question set is denoted by $\calY$ and 
his family of answer sets is denoted by $\{\calB_y\}_y$ for each $y \in \calY$. 
Note that when $\calA_x = \calA$ for each $x \in \calX$, we simply say the output alphabet is $\calA$, 
and we follow this convention on Bob's side too.
The probability distribution over the questions of
a nonlocal game 
is given by a functions $\pi: \calX \times \calY \rightarrow [0,1]$,
A nonlocal game also comes with a family of functions $\{ V_{(x,y)} : \calA_x \times \calB_y \rightarrow \R \}_{x,y}$ for each $(x,y) \in \calX \times \calY$, which is 
the family of scoring functions.
Note that when $\calA_x = \calA$ for all $x$ and $\calB_y = \calB$ for all $y$, we write the scoring function as 
$V: \calA \times \calB \times \calX \times \calY \rightarrow \R$.

We also define a nonlocal test 
to be a tuple $(\calX, \calY, \{\calA_x\}_{x \in \calX}, \{\calB_y\}_{y \in \calY}, \pi)$, where
$\calX, \calY, \{\calA_x\}_{x}, \{\calB_y\}_{y}$ and $\pi$ are as defined for a nonlocal game.
Hence, a nonlocal test does not have scoring functions.

A quantum strategy of a nonlocal game or a nonlocal test presented in terms of projective measurements
consists of projective measurements $\{\{M_x^a\}_{a \in \calA_x} \}_{x \in \calX}$ on Alice's side, 
$\{\{N_y^b\}_{b \in \calB_y} \}_{y \in \calY}$ on Bob's side, and a shared state $\ket{\psi}$. 
By projective measurements, we mean that  
$(M_x^a)^2 = M_x^a = (M_x^a)^\dagger$ and $(N_y^b)^2 = N_y^b = (N_y^b)^\dagger$.
Then Alice and Bob's quantum strategy produces the conditional probability distribution
\begin{align}
	\pr{ab}{xy} = \bra{\psi} M_x^a \x N_y^b \ket{\psi},
\end{align}
for each $(x,y) \in \calX \times \calY$ and $(a,b) \in \calA_x \times \calB_y$.
Note that the tensor product structure indicates that the
two parties cannot communicate with each other, 
which is the reason why we say such task is nonlocal.
For simplicity we omit the tensor product symbol between operators 
acting on different Hilbert spaces in the rest of the work. 
If $M_x^a : \calH_A \rightarrow \calH_A$ and $N_y^b: 
\calH_B \rightarrow \calH_B$, 
then $M_x^a N_y^b: \calH_A\x \calH_B \rightarrow \calH_A \x \calH_B$.
\begin{definition}
	Given input alphabets: $\calX$ and $\calY$, and families of output
	alphabets: $\{\calA_x\}_{x \in \calX}$ and $\{\calB_y\}_{y \in \calY}$, a correlation is a family of conditional probability distributions
	$\{\pr{ab}{xy}: a \in \calA_x, b \in \calB_y\}_{x \in \calX, y \in \calY}$.  
\end{definition}
The size of a correlation is the sum of the sizes of the conditional probability distributions in the correlation.

A particular type of nonlocal games that we are interested in is called binary linear system games.
\begin{definition}[Binary linear system game]
 Let $Hx = c$ be an $m \times n$ system of linear equations over $\Z_2 = \{0, 1\}$ (meaning that $H$ is an $m$-by-$n$ matrix with entries in $\Z_2$ and 
 $c$ is a length-$n$ vector with entries in $\Z_2$). 
 The associated linear system game involves two
 players Alice and Bob. 
 The input and output alphabets are 
 \begin{align*}
     \calX = \{I_i\}_{i=1}^m && &\calY = \{1 \dots n\} \\
     \calA_{I_i} = (\Z_2)^{I_i} && &\calB = \Z_2,
 \end{align*}
 where $I_i = \{j : H(i,j) = 1\}$.
 When Alice and Bob answer with $\ba$ and $b$ for questions $I_i$ and
 $j$ respectively,
 the scoring functions are defined by
 \begin{align}
 	V_{(I_i,j)}(\ba, b) =
	\begin{cases}
		1, \quad \text{if } \sum_{k \in I_i} H(i,k) a(k) \equiv c(i) \pmod 2 \text{ and } a(j) = b \\
		0,  \quad \text{otherwise.}
	\end{cases}
\end{align}
\end{definition}

A widely-used example of a linear system game is the Magic Square game introduced in \cite{magic_square}, which 
have equations
\begin{align*}
    &x_1 + x_2 + x_3 = 0 && x_4 + x_5 + x_6 = 0 &&
    x_7 + x_8 + x_9 = 0 \\
    &x_1 + x_4 + x_7 = 0 && x_2 + x_5 + x_8 = 1 &&
    x_3 + x_6 + x_9 = 0.
\end{align*}
Note that the equations above are taken modulo $2$.
We follow this convention in the rest of our work.
For the formulation of a general linear system games over $\Z_d$, we refer to Section $3.1$ of \cite{coladan2017}.

% \hf{In our work, a binary linear system game is conducted in a slightly different way.
% Instead of always getting an equation number, Alice gets an equation number or 
% a pair of equation number and variable number with equal probability. In the later case,
% she just needs to give assignment to the particular variable.
% By conducting a binary linear system game in this way,
% we force Alice and Bob to share
% a quantum strategy in terms of binary observables.}
% \carl{Is this modification necessary?}
\begin{definition}[Binary observable strategy of a linear system game]
\label{def:q_strat}
A quantum strategy presented in terms of binary observables for the binary linear system game $Hx = c$ consists of 
\begin{enumerate}
	\item a pair of finite-dimensional Hilbert spaces $\calH_A$ and $\calH_B$; 
	\item a collection of binary observables $N_j$, $1 \leq j \leq n$, on $\calH_B$
	such that $N_j^2 = \1$ for every $1 \leq j \leq n$; 
	\item a collection of binary observables $M_{ij}$, $1\leq i \leq m$, $1\leq j\leq n$ 
	on $\calH_A$ such that 
	\begin{enumerate}
		\item $M_{ij}^2 = \1$ for every $i,j$, and
		\item $M_{il}M_{ik} = M_{ik}M_{il}$ for every $i,l,k$ such that $H(i,l) = H(i,k) =1$;
	\end{enumerate} 
	and
	\item a quantum state $\ket{\psi} \in \calH_A \x \calH_B$.
\end{enumerate}
\end{definition}
Note that a measurement strategy of a binary linear system game can be converted to a binary observable strategy. 
Let a measurement strategy be 
$(\{\{M_i^{\ba}\}_{\ba \in X_i}\}_{i=1}^m, \{\{N_j^b\}_{b=0}^1\}_{j=1}^n, \ket{\psi})$, 
where $X_i$ is the answer set of equation $i$.
The corresponding binary observable
strategy consists of, $\ket{\psi}$, $N_j := N_j^0 - N_j^1$ and
\begin{align*}
    M_{ij} := \sum_{\ba: a(j)=0} M_i^{\ba} - \sum_{\ba: a(j) =1}
    M_i^{\ba}
\end{align*}
for $1\leq i \leq m$ and $1 \leq j \leq n$. 
Hence, in the rest of the work, we focus on binary observable strategies of linear system games.
Specifically, we assume that Alice has a binary observable for each variable
in each equation,
so when she is asked the assignment of a particular variable, she simply
measures according to the corresponding observable of that variable.

% Note that any quantum strategy presented in terms of binary observables can be 
% converted to a quantum strategy presented in terms of projective measurement by looking into
% the spectral decompositions of the observables. 
% More specifically, the fact that all the observables $\{M_{ij}\}_{j: H(i,j) = 1}$ commute with each other
% implies that they share a common eigenspace. The collection of eigenvalues associated with one eigenstate is the assignment
% to the whole equation when this eigenstate is measured in a projective measurement strategy \cite{cleve2014}.

A perfect binary observable strategy for the linear system game associated with $Hx = c$ can be constructed from
a finite-dimensional representation of a finitely presentable group
over $\Z_2$, called the solution group. 
\begin{definition}[Solution group of a linear system game]
	\label{def:presentation}
	Let $Hx = c$ be an $m \times n$ linear system. The solution group of this system
	is the group
	\begin{align*}
		\Gamma(H,c) := \ip{
		x_1,\dots x_n, J: &J^2 = x_i^2 = e, Jx_i = x_iJ \text{ for all }1 \leq i \leq n, \\
				& \Pi_{j=1}^n x_j^{H(i,j)} = J^{c(i)} \text{ for all } 1 \leq i \leq m, \text{ and } \\
				& x_l x_k = x_k x_l \text{ if } H(i,k) = H(i,l) = 1 \text{ for some } i
				}.
	\end{align*}
\end{definition}
The relation $J^2 = x_i^2 = e$ means that all the generators are of order $2$.
Let $\Psi$ be a finite-dimensional representation of $\Gamma(H, c)$ such that $\Psi(J) = -\1$ on the carrier space $\C^d$. 
Then a perfect strategy for 
the linear system game ($Hx = c$) has
\begin{enumerate}
	\item $M_{ij} = \Psi(x_j)$ for all $ 1 \leq i \leq m$ and $1 \leq j \leq n$,
	\item $N_j = \Psi(x_j)\tp$ for all $1 \leq j \leq n$, 
	\item the maximally entangled state $\ket{\psi} = \frac{1}{\sd} \sum_{k=1}^d \ket{k}\ket{k}$.
\end{enumerate}

%-------------------------------------------------------------------------------
\subsection{The weighted CHSH inequality \cite{acin2012}.}
%-------------------------------------------------------------------------------
The weighted CHSH inequality a variation of the CHSH inequality \cite{chsh}. 
We formulate it as a nonlocal game.
The input alphabets and output alphabets of Alice and Bob are
$\calX = \calY = \{1, 2\}$ and $\calA = \calB = \Z_2$.
The scoring function for the $\alpha$-weighted CHSH test, when $|\alpha|\geq 1$, is defined by
\begin{align}
	V(a,b,x,y)_{\alpha-CHSH} = 
	\begin{cases}
		\alpha (-1)^{a + b}, \quad \text{if } x = 1 \\
		(-1)^{a + b}, \quad \text{if } x = 2,\; y = 1 \\
		(-1)^{1 + a + b}, \quad \text{if } x = y = 2.
	\end{cases}
\end{align}
The input distribution is $\pi(x,y) = 1/4$ for all $(x,y) \in \calX \times \calY$.

Let Alice and Bob's quantum strategy in terms of projectors be $( \{\{M_x^a\}_a\}_x, \{\{N_y^b\}_b\}_y, \ket{\psi})$. 
We define binary observables $M_x := M_x^0 - M_x^1$ and $N_y := N_y^0 - N_y^1$ from Alice and Bob's projectors.
The weighted CHSH inequality states that 
\begin{align}
	\label{eq:chsh_op}
	\I_\alpha = \alpha\ip{M_1N_1}+\alpha\ip{M_1N_2} + \ip{M_2N_1} - \ip{M_2N_2}\leq 2|\alpha|,
\end{align}
where $\ip{M_x N_y} := \bra{\psi} M_x N_y \ket{\psi}$ is the expectation value of the observables. 
The weighted CHSH inequality is true 
if Alice and Bob share product state $\ket{\phi} = \ket{\phi_A} \x \ket{\phi_B}$.
However, if they share an entangled state $\ket{\psi}$, the value of $\I_\alpha$ can be as large as
$ \I_\alpha^{\max} = 2\sqrt{1+\alpha^2}$ \cite{acin2012}.
\begin{definition}[Ideal strategy to achieve $\I_\alpha^{\max}$ \cite{acin2012}]
	\label{def:ideal}
	Define $\mu := \arctan(1/\alpha)$.
	The ideal strategy for weighted CHSH with parameter $\alpha$, which achieves the value $\I_\alpha^{\max}$, 
	consists of the joint state $\ket{EPR}$ and observables $\tM_1 = \pauliz$, $\tM_2 = \paulix$,
	$\tN_1 = \cos(\mu) \pauliz+ \sin(\mu) \paulix$ and $\tN_2 = \cos(\mu) \pauliz - \sin(\mu) \paulix$.
\end{definition}
An interesting observation of the weighted CHSH inequality is that it admits robust self-testing of the ideal strategy given
above. 
We give the formal statement of this robust self-testing property of $\I_\alpha$ in \cref{sec:chsh}.
In \cref{sec:full_test} we design
a nonlocal test called the extended weighted CHSH test based on 
the ideal correlation that maximally violates the weighted CHSH inequality.

%-----------------------------------------------------------
\subsection{Approximation tools}
%-----------------------------------------------------------
In the rest of the paper, we use the following approximation results implicitly and explicitly.
\begin{proposition}
	\label{prop:inner_pd}
	Let $\ket{v}$ and $\ket{v'}$ be vectors such that the relation $\ket{v} \appd{\epsilon} \ket{v'}$ holds.
	Then for any unit vector $\ket{u}$,
	\begin{align*}
		 \braket{v}{u} \appd{\ep} \braket{v'}{u}.
	\end{align*}	
\end{proposition}
\begin{proof}
	We can write $\ket{v'} = \ket{v} + \ket{v''}$, where $\norm{\ket{v''}} \leq \epsilon$,
	then
	\begin{align*}
	\norm{ \braket{v}{u} - \braket{v'}{u} } = \norm{\braket{v''}{u}} \leq \norm{\ket{u}}\norm{\ket{v''}} \leq \epsilon.
	\end{align*}
\end{proof}
\begin{proposition}
	Let $\ket{u_i}$ and $\ket{v_i}$ for $i = 1,2$ be vectors with norm less than equal to $1$
	such that 
	$\ket{u_1} \appd{\ep_1} \ket{v_1}$ and $\ket{u_2} \appd{\ep_2} \ket{v_2}$.
	Then, 
	\begin{align*}
		\braket{u_1}{u_2} \appd{\ep_1+ \ep_2} \braket{v_1}{v_2}.
	\end{align*}
\end{proposition}
\begin{proof}
	By direct calculation we have
	\begin{align*}
		\norm{ \braket{u_1}{u_2} - \braket{v_1}{v_2}} \leq &\norm{\braket{u_1}{u_2} - \braket{v_1}{u_2}}
		+ \norm{\braket{v_1}{u_2} - \braket{v_1}{v_2}} \\
		\leq & \ep_1 + \ep_2,
	\end{align*}
	where we used \cref{prop:inner_pd}.
\end{proof}
\begin{proposition}
\label{prop:orthog}
	Let $U$ be a unitary matrix and $\ket{u}$ and $\ket{v}$ be two vectors
	such that
	\begin{align*}
		U\ket{u} \appd{\ep_1} a \ket{u} &&
		U\ket{v} \appd{\ep_2} b \ket{v}, 
	\end{align*}
	for some $a \neq b \in \C$ and $\norm{a} = \norm{b} = 1$. Then,
	\begin{align*}
		\norm{\braket{u}{v}} \leq \max(\norm{\ket{u}}, \norm{\ket{v}})\frac{\ep_1+\ep_2}{\norm{1-a^{-1}b}} + \frac{\ep_1\ep_2}{\norm{1-a^{-1}b}}.
	\end{align*}
\end{proposition}
\begin{proof}
	We write $U\ket{u} = a\ket{u} + \ket{u'}$ such that $\norm{\ket{u'}} \leq \ep_1$.
	Similarly, we write $U\ket{v} = b\ket{v} + \ket{v'}$ such that $\norm{\ket{v'}} \leq \ep_2$.
	Then,
	\begin{align*}
		\braket{u}{v} =& \braket{Uu}{Uv} \\
		=& (a^{-1}\bra{u}+\bra{u'})(b\ket{v}+\ket{v'})\\
		=& a^{-1}b \braket{u}{v} + a^{-1}\braket{u}{v'}
		+ b \braket{u'}{v} + \braket{u'}{v'},
	\end{align*}
	which implies that 
	\begin{align*}
		\norm{\braket{v}{u}} &= \frac{\norm{a^{-1}\braket{u}{v'} +  b\braket{u'}{v}+ \braket{u'}{v'}}}{\norm{1 - a^{-1}b}} \\
		&\leq \frac{\norm{a}\norm{\ket{v}}\norm{\ket{u'}}+
		\norm{b}\norm{\ket{u}}\norm{\ket{v'}} + 
		\norm{\ket{u'}}\norm{\ket{v'}}}{\norm{1 - a^{-1}b}} \\
		&\leq \max(\norm{\ket{u}}, \norm{\ket{v}})\frac{\ep_1+\ep_2}{\norm{1-a^{-1}b}}
		+ \frac{\ep_1\ep_2}{\norm{1-a^{-1}b}}.
	\end{align*}
\end{proof}
\begin{proposition}
\label{prop:close_vec}
	Let $\ket{u}$ and $\ket{v}$ be two vectors such that $\norm{\ket{u}} \appd{\ep_1} 1$,
	$\norm{\ket{v}} \appd{\ep_2} 1$, and $\braket{u}{v} \appd{\ep_3} 1$.
	Then
	\begin{align*}
		\ket{u} \appd{O(\sqrt{\ep_1+\ep_2+\ep_3})} \ket{v}.
	\end{align*}
\end{proposition}
\begin{proof}
	Direct calculation gives us that
	\begin{align*}
		\norm{\ket{u} - \ket{v}}^2 &= \norm{\ket{u}}^2 + \norm{\ket{v}}^2 - 2\braket{u}{v} \\
		&\leq (1+\ep_1)^2 + (1+\ep_2)^2 - 2(1 - \ep_3)\\
		&= 2(\ep_1+\ep_2+\ep_3)+ \ep_1^2 + \ep_2^2\\
		& = O(\ep_1+\ep_2+\ep_3).
	\end{align*}
\end{proof}
\begin{proposition}
	Let $\ket{u}$ be a vector with norm less than or equal to $1$ such that for a unitary $U$,
	\begin{align*}
		U\ket{u} \appd{\ep} a\ket{u},
	\end{align*}
	for $a \in \C$ and $\norm{a} = 1$.
	Then,
	\begin{align*}
		U^n \ket{u} \appd{n\ep} a^n \ket{u}
	\end{align*}
	for all $n \geq 1$.
\end{proposition}
\begin{proof}
	We can prove it by induction. 
	The $k=1$ case follows the given condition.
	Assume it holds for $k = n-1$ and consider $k = n$, then we have
	\begin{align*}
		&\norm{U^n \ket{u} - a^n\ket{u}} \\
		\leq &\norm{U} \norm{U^{n-1} \ket{\psi} - a^{n-1}\ket{\psi}} + \norm{a^{n-1}}
		\norm{U\ket{\psi} - a \ket{\psi}}\\
		\leq &(n-1)\ep + \ep = n \ep,
	\end{align*}
	where we used the fact that $\norm{U} = 1$.
\end{proof}
\begin{lemma}[Substitution Lemma]
    \label{lm:sub}
    Let $\calH_A$ and $\calH_B$ be two Hilbert spaces, and
    $\ket{\psi} \in \calH_A \x \calH_B$.
    Suppose there exist unitaries 
    $\{V\} \cup \{V_i\}_{i=1}^k \cup \{ M_i \}_{i=1}^n$ on $\calH_A)$ and
    $\{ N_i \}_{i=1}^n$ on $\calH_B$ such that
    \begin{align*}
        &M_i \ket{\psi} \appd{\ep_1} N_i \ket{\psi}
    \end{align*}
    for $1 \leq i \leq n$, and 
    \begin{align*}
        &V \ket{\psi} \appd{\ep_2} \Pi_{i=1}^k V_i \ket{\psi}.
    \end{align*}
    Then,
    \begin{align*}
        V\Pi_{i=1}^n M_i \ket{\psi} 
        \appd{(2n\ep_1 + \ep_2} \left(\Pi_{i=1}^k V_i\right) 
        \left(\Pi_{i=1}^n M_i \right)\ket{\psi}.
    \end{align*}
\end{lemma}
\begin{proof}
    We prove this lemma by induction on $n$.
    The $n = 0$ case follows the condition that $V \ket{\psi} \appd{\ep_2} \Pi_{i=1}^k V_i \ket{\psi}$.

    Assume the conclusion holds for $n = m$.
    Consider the case $n = m+1$, then
    \begin{align*}
        V\Pi_{i=1}^{m+1} M_i \ket{\psi} 
        &= V \left(\Pi_{i=1}^{m} M_i\right) M_{m+1} \ket{\psi} \\
        &\appd{\ep_1} V \left(\Pi_{i=1}^{m} M_i\right) N_{m+1} \ket{\psi} \\
        &= N_{m+1} V \left(\Pi_{i=1}^{m} M_i\right) \ket{\psi} \\
        &\appd{2m\ep_1+\ep_2} N_{m+1}\left(\Pi_{i=1}^k V_i\right) 
        \left(\Pi_{i=1}^m M_i \right)\ket{\psi} \\
        &=\left(\Pi_{i=1}^k V_i\right) 
        \left(\Pi_{i=1}^m M_i \right) N_{m+1} \ket{\psi} \\
        &\appd{\ep_1} \left(\Pi_{i=1}^k V_i\right) 
        \left(\Pi_{i=1}^{m+1} M_i \right) \ket{\psi}.
    \end{align*}
    Adding the error terms in the derivation above together gives us that
    \begin{align*}
        V\Pi_{i=1}^{m+1} M_i \ket{\psi} 
        \appd{(2m+2)\ep_1+\ep_2} \left(\Pi_{i=1}^k V_i\right) 
        \left(\Pi_{i=1}^{m+1} M_i \right) \ket{\psi}.
    \end{align*}
    By the principle of inductive proof, the proof is complete.
\end{proof}
%===============================================
\section{The binary linear system game $\LS(r)$}
\label{sec:lsg}
%===============================================
In this section, we introduce a binary linear system game that is a component of the full test.
This game is designed to enforce relations that should be satisfied by 
Alice and Bob's observables.

The inspiration comes from 
Slofstra's seminal work \cite{slofstra2017}, in which he
embeds a group with generators $x$ and $y$ satisfying the relation $xyx^{-1}=y^2$
into the solution group of a binary linear system game.  
We generalize the relation $xyx^{-1} = y^2$ and embed the 
following group
\begin{align}
    \Pg = \ip{u,o:uou^{-1} = o^r},
\label{eq:pgroup}
\end{align}
for some $r \in \N$, into the solution group, $\Gamma$, of a binary linear system game, $\LS(r)$.
The steps to embed $\Pg$ into $\Gamma$ follows the procedure detailed 
in the proofs of Proposition~$33$, Lemma~$29$ and Proposition $27$ of \cite{slofstra2017}, 
so the group $\Pg$ (\cref{eq:pgroup}) is first embedded into $\Pg_0$ (\cref{def:g0}), then $\Pg_1$ (\cref{def:g1}), and finally into the group $\Gamma$ (\cref{def:gamma}),
and we obtain a chain of embeddings:
\begin{align}
\Pg \hookrightarrow \Pg_0 \hookrightarrow \Pg_1 \hookrightarrow \Gamma.
\end{align}

This section is structured as follows.
In \cref{sec:g0}, we define the group $\Pg_0$ and prove that $\Pg$ is embedded in $\Pg_0$.
In \cref{sec:g1}, we define the group $\Pg_1$ and prove that $\Pg_0$ is embedded in $\Pg_1$.
\Cref{sec:gamma} contains the definition of the group $\Gamma$ and the proof that $\Pg_1$
is embedded in $\Gamma$.
The purposes of introducing each intermediate
group are explained along the definitions.
In \cref{sec:lsr_def}, we construct $\LS(r)$ from $\Gamma$ and give the 
formal definition of $\LS( r )$.

Note that, we defer the full definitions of the representations of $\Pg_0$, $\Pg_1$ and $\Gamma$
until \cref{sec:rep_gamma}. 
Instead, We only highlight a few operators from a finite-dimensional representation of $\Gamma$.
The representation of $\Gamma$ will be used later to build the ideal strategy of the full test.

%----------------------------------------
\subsection{The group $\Pg_0$}
\label{sec:g0}
%----------------------------------------

The group $\Pg_0$ is a conjugacy group (\cref{def:spe_conj_grp}), and is defined below.
The purpose of this step is to embed $\Pg$ into a group where all the
generators are of order-$2$, because
it is impossible to find a homomorphism from $\Pg$ to a solution group that maps $u$ or $o$, which can have arbitrary order, to a single binary generator. 
On the other hand, a homomorphism from $\Pg_0$ to a solution group can be easier to find because all the generators of $\Pg_0$ are binary.
\begin{definition}	
\label{def:g0}
	Let the set $S_0$ be 
	\begin{align*}
		S_0 = \{o_i\}_{i=1}^{r} \cup \{u_i\}_{i=1}^5,
	\end{align*}
	and the set $R_0$ be the union of 
	\begin{align*}
	&R_{base} = \{s^2 = e | s \in S\},\\
	&R_{u} = \{ u_3 = u_2o_1u_2 \quad u_4 = u_2o_2u_2 \quad 
	u_5 = u_1u_3u_1 \quad o_2 = u_1u_4u_1 \quad u_5 = o_1 o_r o_1\},
	\end{align*}
	and 
	\begin{align*}
	R_o = 
	\begin{cases} 
	\bigcup_{j=1}^{r/2-1}\{ o_{1+2j} = o_1o_{2j}o_1 \quad o_{2+2j} = o_2o_{1+2j}o_2\} \quad \text{ if $r$ is even },\\
	\{o_3 = o_2o_1o_2\} \cup \bigcup_{j=1}^{(r-3)/2}\{
	 o_{2+2j} =o_1o_{1+2j}o_1 \quad o_{3+2j} = o_2o_{2j+2}o_2 \}\quad\text{ if $r$ is odd }.
	\end{cases}
	\end{align*}
	The group $\Pg_0$ is a group with presentation $\ip{R_0:S_0}$.
\end{definition}
The special generators of $\Pg_0$ are $o_1, o_2, u_1$ and $u_2$ because the product $o_1o_2$ behaves similarly as $o$,
and the product $u_1u_2$ behaves similarly as $u$.
Moreover, the conjugacy relations in $R_0$ can be combined to recover the relation $uou^{-1} = o^r$,
as proved in the next proposition.
\begin{proposition}
    \label{prop:g0}
	The group $\Pg$ is embedded in $\Pg_0$.
	Furthermore, the generators of $\Pg_0$ satisfy the relations 
	\begin{align}
	    \label{eq:uo_comm}
	    &(u_1u_2) o_2 (u_2u_1) = o_2 \\ 
	    \label{eq:uo_rel}
		&(u_1u_2)^j o_1o_2 (u_2u_1)^j = (o_1o_2)^{r^j}, 
	\end{align}
	for any $ j \in \N$.
\end{proposition}
\begin{proof}
    We prove \cref{eq:uo_comm} first. It is easy to see that
    \begin{align}
        u_1 (u_2o_2u_2) u_1 = u_1u_4u_1 =o_2.
    \end{align}

    Next, The proof that $\Pg$ is embedded in $\Pg_0$ is based on
    the $j = 1$ case of \cref{eq:uo_rel}.
    Once it is proved, we can use the recipe given in
    the proof of Proposition $33$ of \cite{slofstra2017} 
    to construct a finite-dimensional representation of $\Pg_0$ from a 
    finite-dimensional representation of $\Pg$. 
    Then Lemma $15$ of \cite{slofstra2017}
    proves that a homomorphism $\phi: \Pg \rightarrow \Pg_0$ such that $\phi(u) = u_1u_2$ and 
    $\phi(o) = o_1o_2$ is injective.

	We prove \cref{eq:uo_rel} by induction.
	The base case follows the direct calculation given below
	\begin{align*} 
		u_1u_2 o_1o_2 u_2u_1 = &u_1 (u_2 o_1u_2) (u_2o_2 u_2) u_1 \\
	=& (u_1 u_3 u_1) (u_1 u_4 u_1)\\
	=& u_5 o_2\\
	=& o_1o_ro_1 o_2.
	\end{align*}
	Then we substitute the definitions of $o_k$ for $k = r \dots 3$ into the equation above, which 
	proves the $j=1$ case.

	Assume the statement is true for $j = n$, then for $j = n+1$
	\begin{align*}
		(u_1u_2)^{n+1} o_1o_2 (u_2u_1)^{n+1} =& u_1u_2 (o_1o_2)^{r^n} u_2u_1 \\ 
		 =& (u_1u_2 o_1o_2 u_2u_1)^{r^n} \\
		 =& (o_1o_2)^{r^{n+1}},
	\end{align*}
	where from the first line to the second line, we insert $(u_2u_1)(u_1u_2)$ between every pair of $o_1o_2$, and
	from the second line to the third line, we apply the $j=1$ case
	$r^n$ times. Hence, the statement is true.
\end{proof}
Note that our representation $\Psi_0$ of $\Pg_0$ is different from the representation
constructed in the proof of Proposition $33$ of \cite{slofstra2017}.
We give the formal definition of $\Psi_0$ (\cref{def:rep_g0}) in \cref{sec:rep_gamma}.

To better characterize all the conjugacy
relations of $\Pg_0$ and simplify the presentation of $\Pg_0$, 
we relabel $u_i$ and $o_i$'s as $a_i$'s in such a way that
\begin{align}
	a_j := 
	\begin{cases}
	 o_j &\text{ for } j = 1,2 \\
	 u_{j-2} &\text{ for } j = 3\dots 7 \\
	 o_{j-5} &\text{ for } j = 8 \dots r+5.
	\end{cases}
\end{align}
Then we define the set $C(r)$ of tuples $(i,j,k)$ such that
\begin{align}
\label{eq:cr}
	C(r) = \{ (i,j,k) |  (a_ia_ja_i = a_k) \in R_0 \}.
\end{align}
The size of $C(r)$ is $r+3$. 
Hence, another presentation of the group $\Pg_0$ is
\begin{align*}
    \Pg_0 = \ip{ \{a_i\}_{i=1}^{r+5} : \{a_i^2 =e \}_{i=1}^{r+5} \cup \{ a_ia_ja_i = a_k \vert (i,j,k) \in C(r)\}},
\end{align*}
and we will construct $\Pg_1$ based on this presentation of $\Pg_0$.

%---------------------------------
\subsection{The group $\Pg_1$}
\label{sec:g1}
%---------------------------------
The group $\Pg_1$ is a nice conjugacy group (\cref{def:nice_conj_grp}).
We embed $\Pg_0$ into $\Pg_1$ to
make the construction of a finite-dimension representation of $\Gamma$ from
a finite-dimension representation of $\Pg_1$ easier.
\begin{definition}
\label{def:g1}
Define the set $S_1$ to be
\begin{align*}
	S_1 = \{a_i, b_i, c_i, d_i\}_{i=1}^{r+5} \cup \{ f_0 \} \cup \{h_{jk}\}_{(i,j,k) \in C(r)},
\end{align*}
and the set $R_1$ to be
\begin{align*}
	R_1 = &\{ s^2 = e \vert s \in S_1\}\\ 
	&\cup 
	\{a_i = b_ic_i = f_0d_i, \quad f_0b_if_0 =c_i\}_{i= 1}^{r+5} \\ 
	&\cup
	\{h_{jk}b_j c_k = e, \quad d_ib_jd_i = c_k\}_{(i,j,k) \in C(r)}.
\end{align*}
The group $\Pg_1$ is a group with the presentation $\ip{S_1 : R_1}$.
\end{definition}
Note that the group $\Pg_1$ is the same as the group $K$ used in the proof of Lemma $29$ of \cite{slofstra2017}.
For each $1 \leq i \leq r+5$, the relation $a_i = b_ic_i$ guarantees that $b_i$ commutes with $c_i$.
For each $(i,j,k) \in C(r)$, the linear relation $h_{jk} b_j c_k = e$ guarantees that
$b_j$ commutes with $c_k$.

The relation between $\Pg_0$ and $\Pg_1$ is summarized in the propositions below.
\begin{proposition}
    \label{prop:g0g1}
	The group $\Pg_0$ is embedded in $\Pg_1$. 
\end{proposition}
\begin{proof}
    We prove this proposition by showing that generators $a_i$'s in $S_1$
    satisfy the relation $a_i a_j a_i = a_k$ for all $(i,j,k) \in C(r)$.
    This means the homomorphism $\phi: \Pg_0 \rightarrow \Pg_1$ that maps $a_i$
    to $a_i$ is homomorphism from $\Pg_0$ to $\Pg_1$.

	We first show that $d_i c_j d_i = b_k$, which is because
	\begin{align*}
		d_i c_j d_i = d_i (f_0 b_j f_0) d_i = f_0 (d_i b_j d_i) f_0 = f_0 c_k f_0 = b_k.
	\end{align*}
	Then, we can prove that 
	\begin{align*}
		a_i a_j a_i = f_0 d_i b_j c_j f_0 d_i 
		= (f_0 d_i b_j d_i f_0)(f_0 d_i c_j d_i f_0) 
		= (f_0 c_k f_0)(f_0 b_k f_0) 
		= b_k c_k = a_k,
	\end{align*}
	where we use the fact that $f_0$ commutes with $d_i$.

	A representation of $\Pg_1$ can also be built upon a representation of $\Pg_0$.
    Suppose $\Psi_0$ is a representation of $\Pg_0$ on some Hilbert space $\calH$, then 
    a representation of $\Pg_1$, $\Psi_1$, on the Hilbert space $\C^2 \x \calH$ can
    be constructed such that
    \begin{align*}
        &\Psi_1(a_i) = \1_{\C^2} \x \Psi_0(a_i)
    \end{align*}
    as proved in Lemma $29$ of \cite{slofstra2017}.
    Then, the injectivity of $\phi$ follows Lemma $15$ of \cite{slofstra2017}.
\end{proof}
The consequence of this proposition is that $\Pg$ is embedded in $\Pg_1$.
We give the formal definition of the representation $\Psi_1$ of $\Pg_1$ 
(\cref{def:rep_g1}) in \cref{sec:rep_gamma}.

%-------------------------------------
\subsection{The group $\Gamma$}
\label{sec:gamma}
%-------------------------------------
The group $\Gamma$ is a solution group,
so all the generators of $\Gamma$ are of order-$2$
and all the relations are linear relations.
Moreover, the linear relations can be combined
to recover the conjugacy relations in $R_1$.
\begin{definition}
\label{def:gamma}
Define the set $S$ to be 
\begin{align*}
	S = \{a_i, b_i, c_i, d_i,\{ p_{i,j} \}_{j=1}^5\}_{i=1}^{r+5} \cup \{f_i,g_i,m_i\}_{i=0}^2 \cup \{h_{jk}\}_{(i,j,k) \in C(r)} \cup 
	\{\{q_{\bc,j} \}_{j=1}^6\}_{\bc \in C(r)},
\end{align*}
and the set $R$ to be the union of the following
sets of linear relations:
\begin{itemize}
\item for each $i = 1 \dots r+5$:
\begin{align*}
	R_i =  \{ &a_i = b_ic_i = f_0d_i\quad
	%&g_2 a_i w_i = e \\
	f_0 f_1 f_2 = e \quad b_i f_2 p_{i,1} = e \quad p_{i,1} p_{i,2} p_{i,3} = e\\
	&f_0 p_{i,3} p_{i,4} = e \quad c_i p_{i,4} p_{i,5} =e \quad f_1 p_{i,2} p_{i,5} = e\},
\end{align*}
\item for each $ \bc = (i,j,k) \in C(r)$:
\begin{align*}
	R_{\bc} = \{&h_{jk}b_j c_k = e \quad
	d_i q_{\bc,1} f_2 = e \quad b_j f_2 q_{\bc,2} = e \quad q_{\bc,2} q_{\bc,3} q_{\bc,4} = e\\
	&d_i q_{\bc,4} q_{\bc,5} = e \quad c_k q_{\bc,5} q_{\bc,6} =e \quad q_{\bc,1} q_{\bc,3} q_{\bc,6} = e\},
\end{align*}
\item 
$
	R_{MS} = \{f_0 f_1 f_2 = e,\quad g_0 g_1 g_2 = e,\quad m_0 m_1 m_2 = e,\quad
	f_0 g_2 m_0 = e, \quad f_2 g_0 m_1 = e \quad f_1 g_1 m_2 = J\},
$
\item $R_J = \{ J s J s = e \vert s \in S\} \cup \{J^2 = e\}$,
\item $R_{base} = \{ s^2 = e \vert s \in S\}$.
\end{itemize}
The group $\Gamma$ is a group with the presentation
$\ip{S \cup \{J\} : R}$.
\end{definition}

In the next proposition, we prove that the group $\Pg_1$ is embedded in $\Gamma$.
\begin{proposition}
    \label{prop:g1gamma}
	The group $\Pg_1$ is embedded in $\Gamma$.
	The generators of $\Gamma$ also satisfy the relations: 
	\begin{align}
	    \label{eq:mg1} &f_0 g_0 f_0 g_0 = J,\\
	    \label{eq:mg2} &f_2 g_2 f_2 g_2 = J.
	\end{align}
\end{proposition}
\begin{proof}
    To prove the first claim, we show that the generators
    $b_i$, $c_i$, $d_i$ satisfy the relations
    $d_ib_jd_i = c_k$ for each $\bc = (i,j,k) \in C(r)$.
    \begin{align*}
        d_ib_jd_i =& (k_{\bc 1} f_2)(f_2 k_{\bc 2})(k_{\bc 4} k_{\bc 5})\\ 
        =& k_{\bc 1}(k_{\bc 2}k_{\bc 4})k_{\bc 5} \\
        =& (k_{\bc 1} k_{\bc 3}) k_{\bc 5} \\
        =& k_{\bc 6} k_{\bc 5} \\
        =& c_k.
    \end{align*}
    The derivation of $f_0 b_i f_0 = c_i$ is similar for $1 \leq i \leq r+5$, so we omit it here.
    Hence, a map $\phi:\Pg_1 \rightarrow \Gamma$ that maps $s$
    to $s$ for each $s \in S_1$ is a homomorphism between $Pg_1$ and 
    $Gamma$. 
    A representation of $\Gamma$ can be constructed from a representation 
    of $\Pg_1$ as shown in the proof of Proposition $27$ of \cite{slofstra2017},
    so Lemma $15$ of \cite{slofstra2017} proves that
    the homomorphism $\phi$ is injective.

    \Cref{eq:mg1} can be proved by
    \begin{align*}
        &f_0 g_0 f_0 g_0 \\
        = &(f_1 f_2) (f_2 m_1)  (f_1 f_2) (f_2 m_1)\\
        = &f_1 m_1 f_1 m_1 \\
        = &f_1 (g_0 f_2) f_1 (m_0 m_2) \\
        = &f_1 g_0 (f_2 f_1) (f_0 g_2) m_2 \\
        = &f_1 g_0 g_2 m_2 \\
        = &f_1 g_1 m_2 \\
        = &J.
    \end{align*}
    \Cref{eq:mg2} can be proved analogously.
\end{proof}
Since $\Pg$ is embedded in $\Pg_1$, we can conclude that
$\Pg$ is embedded in $\Gamma$ as well.

The formal definition of a representation of $\Gamma$ is given in \cref{def:rep_gamma}. 
We highlight a few important operators here.
The representation $\Psi$ of $\Gamma$ is defined on the Hilbert space
$W_2 \x W_2 \x W_{d-1}$,
where $W_k$ denotes a Hilbert space of
dimension $k$,
\begin{align}
    \label{eq:basis}
    W_k = \spn\{ \ket{x_j} \}_{j=1}^k.
\end{align}
We denote the standard basis of $W_k$ by $\{\ket{x_j}\}_{j=1}^{k}$.
We denote the identity operator for the subspace $W_k$ by $\1_{W_k}$.

We define an operator $\tO$ on $W_{d-1}$ as
\begin{align*}
    \tO = \sum_{j=1}^{d-1} \omega_d^j \ketbra{x_j}{x_j}.
\end{align*}
Note that another form of the basis of $W_{d-1}$ is $\{\ket{x_{r^j}}\}_{j=0}^{d-2}$, 
where the subscript $r^j$ is taken modulo $d$ implicitly,
because another form of $\tO$ is 
\begin{align}
    \label{eq:O}
    \tO = \sum_{j=0}^{d-2} \omega_d^{r^j} \ketbra{x_{r^j}}{x_{r^j}}.
\end{align}
We use the second form of the basis when it is convenient.
The other basis of $W_{d-1}$ that we use is $\{ \ket{u_j} \}_{j=0}^{d-2}$ defined by
\begin{align*}
	\ket{u_k} = \frac{1}{\sqrt{d-1}} \sum_{j=0}^{d-2} \omega_{d-1}^{jk} \ket{x_{r^j}}.
\end{align*}
The basis $\{\ket{u_j}\}$ is the eigen-basis of operator $\tU$ defined by
\begin{align*}
    \tU = \sum_{j=0}^{d-2} \omega_{d-1}^j \ketbra{u_j}{u_j}.
\end{align*}
Note that in the standard basis, $\tU$ is
\begin{align}
    \label{eq:U}
    \tU = \sum_{j=0}^{d-2} \ketbra{x_{r^{j-1}}}{x_{r^j}}.
\end{align}
The unitaries $\tU$ and $\tO$ are special because 
they satisfy the relation $\tU \tO \tU\ct = \tO^r$.

Here we only present what $a_1, a_2, a_3$ and $a_4$ are mapped to by the representation $\Psi$,
since they are the key generators of $\Gamma$.
\begin{align}
    \label{eq:o1} \Psi(a_1) &= \1_{W_2} \x \1_{W_2} \x \left(\sum_{j=1}^{(d-1)/2} \omega_d^j \ketbra{x_j}{x_{d-j}} + \omega_d^{-j} \ketbra{x_{d-j}}{x_{j}}\right) \\
    \label{eq:o2} \Psi(a_2) &= \1_{W_2} \x \1_{W_2} \x \left(\sum_{j=1}^{d-1} \ketbra{x_j}{x_{d-j}} \right).
\end{align}
\begin{equation}
    \begin{aligned}
    \label{eq:u1} \Psi(a_3) &= \1_{W_2} \x \1_{W_2} \x \Big[\ketbra{u_0}{u_0} +\omega_{d-1}^{(d-1)/2}\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}}\\ 
    &+ \sum_{k=1}^{(d-3)/2}\left( \omega_{d-1}^k\ketbra{u_k}{u_{d-1-k}} + \omega_{d-1}^{-k}\ketbra{u_{d-1-k}}{u_k}\right) \Big]
    \end{aligned}
\end{equation}
\begin{equation}
    \begin{aligned}
    \label{eq:u2}\Psi(a_4) &= \1_{W_2} \x \1_{W_2} \x \Big[\ketbra{u_0}{u_0} +\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}} \\
	& + \sum_{k=1}^{(d-3)/2}\left(\ketbra{u_{d-1-k}}{u_k} + \ketbra{u_k}{u_{d-1-k}}\right)\Big].
    \end{aligned}
\end{equation}
The tensor structure of the representations come from the construction of $\Psi$.
These operators are important because $\Psi(a_1a_2) = \1_{W_2} \x \1_{W_2} \x \tO$
and $\Psi(a_3a_4) = \1_{W_2} \x \1_{W_2} \x \tU$ where $\tO$ is defined in \cref{eq:O} and
$\tU$ is defined in \cref{eq:U}.
The unitaries $\Psi(a_1a_2)$ and $\Psi(a_3a_4)$ satisfy the following
equation
\begin{align*}
    \Psi(a_3a_4) \Psi(a_1a_2) \Psi(a_4a_3) = \Psi(a_1a_2)^r,
\end{align*}
as $\tU \tO \tU\ct = \tO^r$.

% A representation of $\Gamma$ can also be built from
% a representation $\Psi_1$ of $\Pg_1 \circlearrowleft \mathbb{C}^2 \otimes \mathcal{H}$
% that was constructed above.
% We denote the representation of $\Gamma$ by $\Psi$ and
% it is defined on $\C^2 \x \C^2 \x \calH$ such that
% \begin{align*}
%     \Psi(s) = \1_{\C^2} \x \Psi_1(s) \text{ for each }
%     s \in S_1.
% \end{align*}
% In particular, it means that for each $a_i \in S_0$
% \begin{align*}
%     \Psi(a_i) = \1_{\C^2} \x \1_{\C^2} \x \Psi_0(a_i).
% \end{align*}
% Hence, a homomorphism $\phi: \Pg_1 \rightarrow \Gamma$ that 
% maps $s$ to $s$ for each $s \in S_1$ is an embedding of $\Pg_1$ into $\Gamma$
% by Lemma $15$ of Ref.~\cite{slofstra2017}.

%---------------------------------------------------
\subsection{The game $\LS(r)$}
\label{sec:lsr_def}
%---------------------------------------------------
From the group $\Gamma$ we can construct the binary linear system game $\LS(r)$. Basically,
we associate a variable $x(s)$ to each generator $s \in S$.
Then each linear relation in $R$ can be translated to a linear equation.
For example, the linear relation $f_0f_1f_2 = e$ is translated 
to the equation $x(f_0) + x(f_1) + x(f_2) = 0$, and the relation 
$f_1g_1m_2 = J$ is translated to the equation
$x(f_1) + x(g_1) + x(m_2) = 1$. 
The complete list of the equations is given in \cref{def:lsg_eq}.
\begin{definition}
\label{def:lsg}
The linear system game $\LS(r)$ consists of
\begin{align*}
    & \calX^{\LS} = \{I_i\}_{i=1}^{\mr}, \\
    & \calY^{\LS} = \{ x(s) : s \in S \},
\end{align*}
where $I_i = \{ s : H(i,s) = 1\}$ is the set of all the nonzero coefficient of equation $i$
and $\mr := 14r + 62$ is the number of equations.  
Details of $I_i$ for $1 \leq i \leq \mr$ can be found in \cref{def:lsg_eq}.
Alice answer sets are 
\begin{align*}
    \calA^{\LS}_{I_i} = (\Z_2)^{I_i}
\end{align*}
for $1 \leq i \leq \mr$.
Bob's answer set is
\begin{align*}
    \calB^{\LS} = \Z_2.
\end{align*}
The probability distribution is uniform over all the pairs $(I_i, x(s))$ such that $s \in I_i$, so
\begin{align*}
    \pi^{\LS}(I_i, x(s)) = 
    \begin{cases}
        1/N(r) & \text{ if } s \in I_i \\
        0 & \text{otherwise,}
    \end{cases}
\end{align*}
where $N(r) := 157r + 685$ is the total number of valid question pairs.
We define 
\begin{align*}
    X_i := \{ \ba \in (\Z_2)^{I_i}: \sum_{s \in I_i} a(s) = c(i) \},
\end{align*}
which is the set of all the valid answers of equation $i$.
Given an answer $(\ba,b)$ for question $(I_i,x(s))$, 
the score function is
\begin{align*}
    V^{\LS}_{I_i, x(s)}(\ba, b) = 
    \begin{cases}
        1 \text{, if } \ba \in X_i \text{ and } a(s) = b\\
        0 \text{, otherwise}.
    \end{cases}
\end{align*}
\end{definition}

The construction of $\LS(r)$ proves the following theorem.
\begin{theorem}
	The group $\Pg$ can be embedded in the solution group of a binary linear system game which has
	$\nr := 16r+75$ variables and $\mr := 14r + 62$ equations, where each equation has $3$ variables.
\end{theorem}
As indicated in \cref{def:lsg}, the total
number of question pairs is $N(r) = 157r + 685 = \Theta(r)$.

The ideal strategy of $\LS(r)$ is based on
the representation $\Psi$ of $\Gamma$ defined in \cref{def:rep_gamma}.
We discuss the ideal strategy of $\LS(r)$ in \cref{sec:tests}
when we introduce the ideal strategy of the full test.

% \begin{proposition}
% 	In the group $\Gamma(\LS(r))$, the generator $f_2$ commutes with $f_0$ and $a_i$ for all $i = 1 \dots r+5$;
% 	the generator $g_2$ commutes with $g_0$ and $a_i$ for all $i = 1 \dots r+5$.
% \end{proposition}
% \begin{proof}
% 	We show the commutativity property of $f_2$ as an example. The proof for $g_2$ is very similar.
% 	The commutativity between $f_0$ and $f_2$ is immediate from the relations.
% 	From the relation, we can also deduce that $f_2$ commutes with $b_i$ for all $i$.
% 	Since $a_i =b_ic_i = b_i f_0 b_i f_0$, we know $f_2$ commutes with $a_i$.
% \end{proof}

%====================================================
\section{Extending the weighted CHSH inequality $\I_\alpha$ into a nonlocal test}
\label{sec:ext_chsh}
%====================================================
In this section, we introduce the extended weighted CHSH test,
which is going to be a component of the full test introduced in \cref{sec:full_test}.
The ideal correlation for this test contains the ideal correlation that 
achieves $\I_{-\cot(\pi/d)}^{\max}$. 
To make the study of the implication of a strategy that realize 
the ideal correlation of the extended weighted CHSH test easier,
we fill in the background of the robust self-testing property of the
general weighted CHSH inequality, $\I_\alpha$, first.

%----------------------------------------------------
\subsection{Robust self-testing of $\ket{EPR}$ \textit{via} $\I_\alpha$}
\label{sec:chsh}
%----------------------------------------------------
In this section, we discuss how the weighted CHSH inequality can be used to robustly self-test $\ket{EPR}$ and Pauli operators defined in \cref{def:ideal}.

% Intuitively, robust self-testing means that if some quantum strategy achieves a value $\I_\alpha$ that is close to $\I_\alpha^{\max}$,
% then the shared state must be close to $\ket{EPR}$ and their observables are close to the rotated Pauli operators 
% defined in \cref{def:ideal},
% up to some local isometry.
The formal statement is given in the theorem below. Since the proof uses the same techniques as in \cite{bamps2015},
we defer the proof till \cref{sec:w_chsh_proof}, and we only list the key relations that will be reused in \cref{sec:main} here.
\begin{theorem}
\label{thm:selftest}
    Define $\mu := \arctan(1/\alpha)$.
	If a quantum strategy $(\ket{\psi}, \{M_x:=M_x^0-M_x^1\}_{x=1,2}, \{N_y :=N_y^0-N_y^1\}_{y = 1,2} )$ achieves that
	$\I_\alpha \geq \I_\alpha^{\max} - \epsilon$
	for some $\epsilon > 0$, then
	there exists a local isometry $\Phi = \Phi_A \x \Phi_B$ and an auxiliary state $\ket{junk}$  such that
	\begin{align*}
		\| \Phi( M_x \x N_y \ket{\psi}) -\ket{junk} \x (\tM_x \x \tN_y) \ket{EPR}^{\x 2}  \| = O(\frac{1}{\cos^2(\mu)\sin^{1/2}(\mu)}+
		\frac{1}{\sin(\mu)^{3/2}})) \sqrt{\epsilon}
	\end{align*}
	for $x,y \in \{0, 1, 2\}$ where the subscript $0$ refers to the identity operator and $\tM_x, \tN_y$ are 
	defined in \cref{def:ideal}.  
\end{theorem}
We follow the convention used in \cite{bamps2015},
which defines
\begin{align*}
	&Z_A := M_1 && X_A := M_2\\
	&Z_B := \frac{N_1+N_2}{2\cos{\mu}} && X_B := \frac{N_1-N_2}{2\sin{\mu}}.
\end{align*}
The key relations that will be reused are
\begin{align}
% 	\label{eq:za-zb}& Z_A\ket{\psi} \appd{O( \sqrt{\mu} \se)} Z_B\ket{\psi},\\
% 	\label{eq:xa-xb}&X_A\ket{\psi} \appd{O(\frac{\se}{\sqrt{\mu}})} X_B \ket{\psi}, \\
% 	\label{eq:xazb}&X_A(\1+Z_B)\ket{\psi} \appd{O(\frac{\se}{\sqrt{\mu}})} X_B(\1-Z_A) \ket{\psi},\\
% 	\label{eq:zaxb}&Z_A(\1+X_B)\ket{\psi} \appd{O(\frac{\se}{\sqrt{\mu}})} Z_B(\1-X_A) \ket{\psi},\\
% 	\label{eq:zaxa}&Z_AX_A\ket{\psi} \appd{O(\frac{\se}{\sqrt{\mu}})} -X_AZ_A \ket{\psi},\\
% 	\label{eq:zaxaxbzb}&X_AZ_A \ket{\psi} \appd{O(\frac{\se}{\mu^{3/2}})} -X_BZ_B \ket{\psi},
\label{eq:za-zb}& Z_A\ket{\psi} \appd{\frac{\sqrt{s\epsilon}}{c}} Z_B\ket{\psi},\\
	\label{eq:xa-xb}&X_A\ket{\psi} \appd{\frac{\sqrt{\epsilon}}{\sqrt{s}}} X_B \ket{\psi}, \\
	\label{eq:xazb}&X_A(\1+Z_B)\ket{\psi} \appd{\frac{c+1}{c\sqrt{s}} \se} X_B(\1-Z_A) \ket{\psi},\\
	\label{eq:zaxb}&Z_A(\1+X_B)\ket{\psi} \appd{ \frac{s+1}{c\sqrt{s}} \se} Z_B(\1-X_A) \ket{\psi},\\
	\label{eq:zaxa}&Z_AX_A\ket{\psi} \appd{\frac{1+c+s}{c\sqrt{s}} \se} -X_AZ_A \ket{\psi},\\
	\label{eq:zaxaxbzb}&X_AZ_A \ket{\psi} \appd{\frac{2s^2+cs+c+s}{cs\sqrt{s}} \sqrt{\epsilon}} -X_BZ_B \ket{\psi}.
\end{align}
We keep the exact error terms so that some
of the substitution steps in \cref{sec:main} are 
easier to follow.

%-------------------------------------------
\subsection{The extended weighted CHSH test}
\label{sec:ext_w_chsh}
%-------------------------------------------
In this subsection, we first give the definition of the 
extended weighted CHSH test, then we discuss the ideal strategy of it.
We defer the ideal correlation of this test til \cref{sec:cor}
as it will be presented as a part of the ideal correlation of the full test.
\begin{definition}
    \label{def:ext_w_chsh}
    In the extended weighted CHSH test, the input alphabets and output alphabets for Alice and Bob are
    \begin{align*}
        &\calX^{\EXT} = \calY^{\EXT} = \{0, 1, 2, \nr+1, \nr+2\}, \\
        &\calA_z^{\EXT} = \calB_z^{\EXT} =
        \begin{cases}
            \{0,2\} &\text{ if } z = 0, \\
            \{0,1\} &\text{ if } z \in \{1, 2\} \\
            \{0,1,2\} &\text{ if } z \in \{\nr+1, \nr+2\}. 
        \end{cases}
    \end{align*}
    The probability distribution over all the question pairs 
    $(x,y) \in \calX^{\EXT} \times \calY^{\EXT}$ is uniform.
\end{definition}
The reason why this test is called the extended weighted CHSH test will be clear 
after we introduce the ideal strategy.

Recall the Hilbert space $W_{d-1}$ defined in \cref{eq:basis}. 
To make the shared state and the observables of the ideal strategy have
simpler forms, we introduce the third basis of $W_{d-1}$: $\{ \ket{j} \}_{j=1}^{d-1}$, which is defined by
\begin{align*}
    &\ket{j} = -\frac{1}{\sqrt{2}}(\ket{x_j} + e^{-ij\pi/d}\ket{x_{d-j}}) \\
    &\ket{d-j} = \frac{i}{\sqrt{2}}(\ket{x_j} - e^{-ij\pi/d}\ket{x_{d-j}})
	%\ket{x_j} = \frac{-1}{\sqrt{2}}(\ket{j} + i\ket{d-j}), \\
	%\ket{x_{d-j}} = \frac{-e^{ij\pi/d}}{\sqrt{2}}(\ket{j} - i\ket{d-j})
\end{align*}
for $1 \leq j \leq \frac{d-1}{2}$.
The shared state of the ideal strategy is 
\begin{equation}
    \label{eq:tpsi}
    \begin{aligned}
        \ket{\tpsi}^{\EXT}
         =& \frac{1}{\sqrt{d-1}} \sum_{j=1}^{(d-1)/2} e^{ij\pi/d}(\ket{j}\ket{j} + \ket{d-j}\ket{d-j}),
    \end{aligned}
\end{equation}
where the superscript $\EXT$ stresses the fact this state is for the extended weighted CHSH test.
Note that in the standard basis of $W_{d-1}$,
\begin{align*}
    \ket{\tpsi}^{\EXT} = \frac{1}{\sqrt{d-1}} \sum_{j=1}^{d-1} \ket{x_j}\ket{x_{d-j}}.
\end{align*}

To introduce the projectors, 
we first define two subspaces of $W_{d-1}$: $V = \spn\{\ket{1}, \ket{d-1}\}$ and $V^\perp = W_{d-1} \setminus\spn\{\ket{1}, \ket{d-1}\}$.
The corresponding projective measurements are $\Pi_V$ and $\Pi_{V^\perp}$. Note that $V$ is the subspace on which they should maximize $\I_{\cot(-\pi/d)}$.

First of all, the ideal strategy ensures that when 
Alice and Bob get the same question from the set $\{0,\nr+1,\nr+2\}$, they always give the same answer.
Alice's and Bob's measurements for questions $0, \nr+1$ and $\nr+2$ are
\begin{align*}
	&\tP_0^0 = \tQ_0^0 = \Pi_V && \tP_0^2 = \tQ_0^2 = \Pi_{V^\perp} \\
	&\tP_{\nr+1}^0 = \tQ_{\nr+1}^0 = \ketbra{1}{1} &&
	\tP_{\nr+1}^1 = \tQ_{\nr+1}^1 = \ketbra{d-1}{d-1} \\
	&\tP_{\nr+2}^0 = \tQ_{\nr+2}^0 = \ketbra{1_{\times}}{1_{\times}} && \tP_{\nr+2}^1 = \tQ_{\nr+2}^1 = \ketbra{(d-1)_{\times}}{(d-1)_{\times}} \\
	&\tP_{\nr+1}^2 = \tQ_{\nr+1}^2 = \Pi_{V^\perp} &&
	 \tP_{\nr+2}^2 = \tQ_{\nr+2}^2 = \Pi_{V^\perp},
\end{align*}
where
\begin{align*}
    \ket{1_{\times}} = \frac{1}{\sqrt{2}}\ket{1} + \frac{1}{\sqrt{2}}\ket{d-1}
	&&\ket{(d-1)_{\times}} = \frac{1}{\sqrt{2}}\ket{1} - \frac{1}{\sqrt{2}}\ket{d-1}.
\end{align*}
Intuitively, these questions force Alice and Bob to agree on the subspace $V$ and
on the basis of $V$ that they will measure.

The ideal strategy also ensures that 
when Alice gets $x \in \{1,2\}$ and Bob gets $y \in \{\nr+1, \nr+2\}$, 
their answers $a,b \in \{0,1\}$ should maximize $\I_{\cot(-\pi/d)}$ with Alice and Bob's roles flipped;
and when Alice gets $x \in \{\nr+1,\nr+2\}$ and Bob gets $y \in \{1, 2\}$,
their answers $a,b \in \{0,1\}$ should maximize $\I_{\cot(-\pi/d)}$.
This is the reason behind naming this test the extended weighted CHSH test.
It is also the reason behind having Alice and Bob agreeing on the subspace $V$.
Before introducing the measurements, 
we need to define some basis states of $V$
\begin{align*}
	&\ket{1_+} = \cos(-\frac{\pi}{2d})\ket{1} + \sin(-\frac{\pi}{2d})\ket{d-1}
	&&\ket{(d-1)_+} = \sin(-\frac{\pi}{2d})\ket{1} - \cos(-\frac{\pi}{2d})\ket{d-1}\\
	&\ket{1_-} = \cos(-\frac{\pi}{2d})\ket{1} - \sin(-\frac{\pi}{2d})\ket{d-1}
	&&\ket{(d-1)_-} = \sin(-\frac{\pi}{2d})\ket{1} + \cos(-\frac{\pi}{2d})\ket{d-1},
\end{align*}
so that $\{\ket{1_+}, \ket{(d-1)}_+\}$ and $\{\ket{1_-}, \ket{(d-1)_-}\}$ are two basis
of $V$.
When Alice and Bob gets question $1$ or $2$, what they measure on $V$ are
\begin{align*}
	&\tP_1^0 = \tQ_1^0 =  \ketbra{1_+}{1_+} && \tP_1^1 = \tQ_1^1 =  \ketbra{(d-1)_+}{(d-1)_+}\\
	&\tP_2^0 = \tQ_2^0 =  \ketbra{1_-}{1_-} && \tP_1^1 = \tQ_1^1 =  \ketbra{(d-1)_-}{(d-1)_-}.
\end{align*}
Notice that their measurements on $V$ follow the measurements to achieve $\I_{\cot(-\pi/d)}^{\max}$.
What they measure on $V^\perp$ is irrelevant for this test. 
%Later we will see that their measurements on $V^\perp$ are determined by their observables for variables $x(a_1)$ and $x(a_2)$ of $\LS(r)$.

% We will present the ideal correlation of the extended weighted CHSH test
% in \cref{sec:cor} as part of the full correlation.

%==========================
\section{The full test}
\label{sec:full_test}
%==========================
The full test consists of the linear system game $\LS(r)$,
the extended weighted CHSH test and a commutation game. We explain how the games and test are conducted and
introduce the ideal strategy for each of them in \cref{sec:tests}. 
Then we explain how they are combined in the full test and present the ideal strategy for the full test in \cref{sec:ideal_strat}. The full correlation, denoted by $\fC(r,d)$, is generated by the ideal strategy and presented in \cref{sec:cor}.

%----------------------------
\subsection{The components of the full test}
\label{sec:tests}
%----------------------------

The first component of the full test is $\LS(r)$ (\cref{def:lsg}).
The ideal strategy of $\LS(r)$ follows the representation $\Psi$ of $\Gamma$ (\cref{def:rep_gamma}).
Recall the two Hilbert spaces $W_{d-1}$ and $W_2$ defined in \cref{eq:basis}. 
In the ideal strategy Alice and Bob share the state 
\begin{align}
    \label{eq:ideal_st_ls}
        \ket{\tpsi}^{\LS} = \frac{1}{2} (\ket{x_1}\ket{x_1} + \ket{x_2}\ket{x_2})^{\x 2} \x \left(\frac{1}{\sqrt{d-1}} \sum_{j=1}^{d-1} \ket{x_j}\ket{x_{d-j}}\right).
\end{align}
Alice and Bob's observables for each variable $x(s)$ are
defined by
\begin{align*}
    \tM(s) = \tN(s) = \Psi(s) 
\end{align*}
for each $s \in S$.

The second component is the extended weighted CHSH test, introduced in \cref{sec:ext_w_chsh}.

% Note that when the referee wants to ask Alice $x(a_1)$ or $x(a_2)$, which also labels a variable of $\LS(r)$, 
% the referee gives Alice an equation label $i$ such that  
% the equation $i$ contains the variable $x(a_1)$ or $x(a_2)$. 
% For simplicity we omit the equation label in the alphabet $\calX$, and
% later we will show that the equation label can be made implicit.
% Only some of the question pairs of $\calX \times \calY$ are given to Alice and Bob.
% Alice and Bob's ideal behaviour to pass this test is summarized 
% below:
% \carl{[Does the phrase ``ideal behavior'' have a formal definition? This seems to be another case where are you are writing in a way that is sort of formal, and sort of just a sketch -- that's not a good way to write.  Do one or the other.]}
% \begin{enumerate}
% 	\item if $x = y = \ast$, Alice and Bob should answer with $a = b \in \{0, \perp\}$;
% 	\item if $x = y \in \{x_{\nr+1}, x_{\nr+2}\}$, Alice and Bob should answer with $a = b \in \{0, 1, \perp\}$;
% 	\item if $x \in \{x(a_1),x(a_2)\}$ and $y \in \{x_{\nr+1}, x_{\nr+2}\}$, 
% 		\begin{enumerate}
% 		\item if they answer with $a,b \in \{0,1\}$, then
% 	they should maximize $\I_{\cot(-\pi/d)}$ with Alice and Bob's roles flipped, 
% 		\item if Bob answers with $\perp$, then Alice's answer doesn't matter;
% 		\end{enumerate}
% 	\item if $x \in \{x_{\nr+1}, x_{\nr+2}\}$ and $y \in \{x(a_1),x(a_2)\}$,
% 	\begin{enumerate} 
% 		\item if they answer with $a,b \in \{0,1\}$, then
% 		they should maximize $\I_{\cot(-\pi/d)}$, 
% 		\item if Alice answers with $\perp$, then Bob's answer doesn't matter;
% 	\end{enumerate}
% 	\item $x \in \{x_{\nr+1},x_{\nr+2}\}, y = \ast$ 
% 		\begin{enumerate}
% 		\item if Bob answers with $0$, Alice should answer with $\{0.1\}$ but not $\perp$, 
% 		\item if Bob answers with $\perp$ Alice should answer with $\perp$ too;
% 		\end{enumerate}
% 	\item $x = \ast, y \in \{x_{\nr+1}, x_{\nr+2}\}$
% 	\begin{enumerate}
% 		\item if Alice answers with $0$, Bob should answer with $\{0.1\}$ but not $\perp$, 
% 		\item if Alice answers $\perp$, Bob should answer $\perp$ too.
% 	\end{enumerate}
% \end{enumerate}
% Only the question pairs mentioned in the list above are given to Alice and Bob.

The third component is a commutation game.
Our commutation game is based on the commutation test used in \cite{coladan2017verifier}.
The goal of the commutation game is to make sure that some
of Alice's binary observables for $\LS(r)$ 
commute with her observable for the extended weighted CHSH test,
when acting on the shared state.
So we reuse questions and answers from the two other components
of the full test.

Bob's input set consists of eight question pairs
\begin{align*}
    \calY^{\COMM} = \{(\nr+1, x(s)), (\nr+2, x(s)): s = f_0, f_2, g_0, g_2 \},
\end{align*}
and
Alice's input set is 
\begin{align*}
    \calX^{\COMM} = \{\nr+1, \nr+2, x(f_0), x(f_2), x(g_0), x(g_2)\}.
\end{align*}
The answer sets are 
\begin{align*}
    &\calA^{\COMM}_x = 
    \begin{cases}
    \Z_3 &\text{ if } x \in \{\nr+1, \nr+2\} \\
    \Z_2 &\text{otherwise.}
    \end{cases}
    \\
    &\calB^{\COMM} = \Z_3 \times \Z_2.
\end{align*}
The probability distribution is
\begin{align*}
    \pi^{\COMM}(x,(y_1,y_2)) =
    \begin{cases}
        1/16 &\text{ if } x = y_1 \text{ or } x = y_2 \\
        0   &\text{ otherwise}.
    \end{cases}
\end{align*}
The scoring function is
\begin{align*}
    V^{\COMM}(a, (b_1,b_2), x, (y_1,y_2)) = 
    \begin{cases}
    1 &\text{ if } x = y_1 \text{ and } a = b_1 \\
    1 &\text{ if } x = y_2 \text{ and } a = b_2 \\
    0 &\text{ otherwise}.
    \end{cases}
\end{align*}
The probability distribution makes sure that Alice gets one of
the two questions that Bob gets.
To win this game, Alice and Bob should agree on the answer to the
common question.
% We remark that if Alice's question is to assign a value to some variable of $\LS(r)$,
% the label of a equation that contains this variable is given 
% to Alice and the referee checks the bit corresponding to this
% variable in Alice's answer. This is equivalent to asking 
% Alice to measure an observable for the variable.

We defer the ideal strategy for this test till the next subsection because the ideal
strategy for the other two tests above can win this game as well.

% What we conclude is slightly different from the original claim
% in Ref.~\cite{coladan2017verifier}.
% Instead of just proving the binary observables commute,
% we prove that the projective measurements
% for question $\nr+1$ commute with the binary observables for question
% $x(s)$ with $s = f_0, f_2, g_0, g_2$ 
% by examining the probability distribution generated by the ideal strategy.  \carl{I don't think there's enough context for this paragraph.  It may be best to leave it out.}

% In the full test, the three tests are conducted with equal probability.
% The full correlation, denoted by $\fC(d,r)$, is generated by the ideal strategy that can pass all the three tests. We give the ideal strategy first and then $\fC(d,r)$.
%------------------------------------------------
\subsection{The full test and the ideal strategy}
\label{sec:ideal_strat}
%-----------------------------------------------
\begin{definition}[The full test]
\label{def:full_test}
Define 
\begin{align*}
    \calX^{\EXT'} = \calY^{\EXT'} = \{0, x(a_1), x(a_2), \nr+1, \nr+2\}.
\end{align*}
The question sets of the full test are
\begin{align*}
    \calX = \calX^{\LS} \cup \calX^{\EXT'} \cup \calX^{\COMM} \\
    \calY = \calY^{\LS} \cup \calY^{\EXT'} \cup \calY^{\COMM},
\end{align*}
Define answer sets $\calA_x^{\EXT'}$ for each $x \in \calX^{\EXT'}$
by 
\begin{align*}
    &\calA_{x(a_1)}^{\EXT'} =\calA_{1}^{\EXT} \\
    &\calA_{x(a_2)}^{\EXT'} =\calA_{2}^{\EXT} \\
    &\calA_{x}^{\EXT'} = \calA_{x}^{\EXT} 
    \quad \text{ for } x \neq x(a_1), x(a_2)
\end{align*}
The answer sets $\calB_y^{\EXT'}$ are defined analogously.
The answer sets of the full test are
\begin{align*}
    \calA_x = 
    \begin{cases}
    \calA_x^{\LS} &\text{ if } x \in \calX^{\LS} \\
    \calA_x^{\EXT'} &\text{ if } x \in \calX^{\EXT'} \\
    \calA_x^{\COMM} & \text{ if } x \in \calX^{\COMM},
    \end{cases}
    &&
    \calB_y = 
    \begin{cases}
    \Z_2 &\text{ if } y \in \calY^{\LS} \\
    \calB_y^{\EXT'} &\text{ if } y \in \calY^{\EXT'} \\
    \calB_y^{\COMM} & \text{ if } y \in \calY^{\COMM}.
    \end{cases}
\end{align*}
Define 
\begin{align*}
    N :=  N(r) + 25 + 16 = 157r + 726.
\end{align*}
The question distribution of the full test, $\pi: \calX \times \calY \rightarrow \R$ is defined by
\begin{align*}
    \pi(x,y) = \begin{cases}
    \frac{1}{N} & \text{ if } (x,y) \in \calX^{\LS} \times \calY^{\LS} 
    \text{ and } \pi^{\LS}(x,y) > 0 \\
    \frac{1}{N} & \text{ if } (x,y) \in \calX^{\EXT'} \times \calY^{\EXT'}\\
    \frac{1}{N} & \text{ if } (x,y) \in \calX^{\COMM} \times \calY^{\COMM} 
    \text{ and } \pi^{\COMM}(x,y) > 0 \\
    0 & \text{ otherwise}.
    \end{cases}
\end{align*}
\end{definition}
Note that $\calX^{\EXT'}$ is obtained from $\calX^{\EXT}$ by replacing 
$1$ by $x(a_1)$ and $2$ by $x(a_2)$, and similarly for $\calY^{\EXT'}$.
The answer sets $\calA_x^{\EXT'}$ are constructed from $\calA_x^{\EXT}$
in a similar way.

The ideal strategy for the full test is an extension of the ideal strategy of $\LS(r)$ to make it able to produce the ideal correlation of 
the extended weighted CHSH test.
Recall the subspace $W_{d-1}$ and $W_2$ defined in \cref{eq:basis}.
The shared state of the ideal strategy is
\begin{align*}
    \ket{\tpsi} = \ket{\tpsi}^{\LS} = \frac{1}{2}(\ket{x_1}\ket{x_1} + 
    \ket{x_2}\ket{x_2})^{\x 2} \x \ket{\tpsi}^{\EXT}.
\end{align*}
Alice and Bob's observables for each variable in the linear system of 
$\LS(r)$ are 
\begin{align*}
    \tM(s) = \tN(s) = \Psi(s) 
\end{align*}
for each $s \in S$ and $\Psi$ is defined in \cref{def:rep_gamma}.
When Alice get $I_i \in \calX^{\LS}$, she measures her half of the shared
state according to $\tM(s)$ for each $s \in I_i$.

Alice's and Bob's observables for question $x(a_1)$ and $x(a_2)$ of $\calX^{\EXT'}$ and $\calY^{\EXT'}$ are
\begin{align*}
    \tM(a_j) = \tN(a_j) = \Psi(a_j) 
\end{align*}
for $j= 1,2$.
They can reproduce the ideal correlation because
\begin{align*}
    &(\1_{W_2} \x \1_{W_2} \x \Pi_V) \tM(a_1) (\1_{W_2} \x \1_{W_2} \x \Pi_V) = \1_{W_2} \x \1_{W_2} \x (\tP_1^0 - \tP_1^1) \\
    &(\1_{W_2} \x \1_{W_2} \x \Pi_V) \tM(a_2) (\1_{W_2} \x \1_{W_2} \x \Pi_V) = \1_{W_2} \x \1_{W_2} \x (\tP_2^0 - \tP_2^1),
\end{align*}
and similarly on Bob's side.
In other words, the observables $\tM(a_j)$ and $\tN(a_j)$ are 
measuring the same state on the subspace $V$ as $\tP_j$ and $\tQ_j$
for $j = 1,2$.

Recall the ideal projectors $\tP_x^a$ and $\tQ_y^b$ for the extended 
weighted CHSH test defined in \cref{sec:ext_w_chsh}.
Alice's projectors for the question $x \in \{0, \nr+1, \nr+2\}$ of $\calX^{\EXT'}$ are
\begin{align*}
    \tM_x^a = \1_{W_2} \x \1_{W_2} \x \tP_x^a
\end{align*}
for each $a \in \calA^{\EXT'}_x$.
Bob's projectors for the question $y \in \{0, \nr+1, \nr+2\}$ of $\calY^{\EXT'}$ are 
\begin{align*}
    \tN_y^b = \1_{W_2} \x \1_{W_2} \x \tQ_y^b
\end{align*}
for each $b \in \calB^{\EXT'}_y$.

Alice's measurements for $x \in \calX^{\COMM}$ is the same as her
measurements for the other two tests.
When Bob gets question $(y_1, y_2) \in \calY^{\COMM}$, he measures 
according to $\tN_{y_1}^b$ first and then $\tN(y_2)$ because $\tN_{y_1}^b$
commutes with $\tN(y_2)$ for all $b$ by construction.
This ensures that the ideal strategy can win the commutation game.

An important property of the ideal strategy for it to win $\LS(r)$ is summarized in the 
following proposition.
\begin{proposition}
The observables $\tM(s)$, $\tN(s)$
and the shared state $\ket{\tpsi}$ defined above
satisfy the relations
\begin{align}
	\label{eq:mn_preserve} &\tM(s) \tN(s) \ket{\tpsi} = \ket{\tpsi},
\end{align}
for each $s \in S$.
\end{proposition}
\begin{proof}
    Define $S' = \{f_0, f_1, f_2, g_0, g_1, g_2, m_0, m_1, m_2\} \subset S$. Notice that for each $s \in S'$, both $\tM(s)$ and $\tN(s)$ act
    trivially on $W_{d-1}$, so we omit the verification about 
    $\tM(s)\tN(s) \ket{\tpsi} = \ket{\tpsi}$ for each $s \in S'$

	Next, we notice that each $s \in S$ can be expressed as the 
	product of a sequence of generators in $S'$ and $a_1, a_2, a_3, a_4$
	following the relations in $R$, so
	to prove \cref{eq:mn_preserve}, it suffices to prove that it 
	holds true for $s = a_1, a_2, a_3, a_4$. 
	Recall that $\tM(a_j) = \tN(a_j) = \1_{W_2} \x \1_{W_2} \x Psi_0(a_j)$
	in \cref{def:rep_gamma},
	where $\Psi_0$ is defined in \cref{def:rep_g0},
	and $\ket{\tpsi} = \frac{1}{2} (\ket{x_1}\ket{x_1}+\ket{x_2}\ket{x_2})^{\x 2} \x \ket{\tpsi}^{\EXT}$.
	Hence, the problem is further reduced to proving
	\begin{align*}
	    \Psi_0(a_j) \x \Psi_0(a_j) \ket{\tpsi}^{\EXT} = \ket{\tpsi}^{\EXT}.
	\end{align*}
	for $1 \leq j \leq 4$.

	Using \cref{eq:o1,eq:o2}, it is straightforward to see that $\Psi_0(a_j)\x \Psi_0(a_j) \ket{\tpsi}^{\EXT} = \ket{\tpsi}^{\EXT}$ for $j = 1,2$.
	Before checking the two cases $j=3$ and $j=4$, we can use \cref{eq:u1u2} to derive that
	\begin{align*}
		\Psi_0(a_3 a_4) \x\Psi_0(a_3 a_4) \ket{\tpsi}^{\EXT} = \ket{\tpsi}^{\EXT}.
	\end{align*}
	The eigenvectors of $\Psi_0(a_3 a_4)$ are $\ket{u_k}$,
	for $0 \leq k \leq d-2$,
	and $\ket{u_k}$ are defined by
    \begin{align*}
        \ket{u_k} = \frac{1}{\sqrt{d-1}} \sum_{j=0}^{d-2} \omega_{d-1}^{jk} \ket{x_{r^j}}.
    \end{align*}
    Hence we can express $\ket{x_{r^j}}$ in terms of $\ket{u_k}$ as 
    \begin{align*}
        \ket{x_{r^j}} = \frac{1}{\sqrt{d-1}} \sum_{k=0}^{d-2} \omega_{d-1}^{-jk} \ket{u_k}.
    \end{align*}
    Then using the expression of $\Psi_0(a_4)$ from \cref{def:rep_g0},
    we can deduce that
    \begin{align*}
        \Psi_0(a_4) \ket{x_{r^j}} =& \frac{1}{\sqrt{d-1}} \Psi_0(a_4) \left(\ket{u_0} + \omega_{d-1}^{-j(d-1)/2}\ket{u_{(d-1)/2}}  + 
        \sum_{k=1}^{(d-3)/2} (\omega_{d-1}^{-jk} \ket{u_k} + \omega_{d-1}^{jk} \ket{u_{d-1-k}})\right) \\
        =&\frac{1}{\sqrt{d-1}} \left(\ket{u_0} + (-1)^j\ket{u_{(d-1)/2}}  + 
        \sum_{k=1}^{(d-3)/2} (\omega_{d-1}^{-jk} \ket{u_{d-1-k}} + \omega_{d-1}^{jk} \ket{u_{k}})\right) \\
        =& \ket{x_{r^{d-1-j}}}.
    \end{align*}
    Note that the expression of $\Psi_0(a_4)$ can also be found in
    \cref{eq:u2}, as $\Psi(a_4) = \1_{W_2} \x \1_{W_2} \x \Psi_0(a_4)$.
    Consequently, we can show that $\Psi_0(a_4) \x \Psi_0(a_4) \ket{\tpsi}^{\EXT} = \ket{\tpsi}^{\EXT}$ 
    because
    \begin{align*}
        \Psi_0(a_4) \x \Psi_0(a_4) \ket{\tpsi}^{\EXT} &= \frac{1}{\sqrt{d-1}}  \sum_{j=0}^{d-2} \Psi_0(a_4)\x  \Psi_0(a_4) \ket{x_{r^j}} \ket{x_{r^{d-1-j}}} \\
        &= \frac{1}{\sqrt{d-1}} \sum_{j=0}^{d-2} \ket{x_{r^{d-1-j}}} \ket{x_{r^{j}}}\\ &=\ket{\tpsi}^{\EXT},
    \end{align*}
	Combining it with the fact that $\Psi_0(a_3 a_4) \x \Psi_0(a_3 a_4) \ket{\tpsi}^{\EXT} = \ket{\tpsi}^{\EXT}$, we can conclude that
	$\Psi_0(a_3) \x \Psi_0(a_3) \ket{\tpsi}^{\EXT} = \ket{\tpsi}^{\EXT}$.
\end{proof}

%---------------------------------------------------
\subsection{The correlation $\fC(d,r)$}
\label{sec:cor}
%---------------------------------------------------
The full correlation is generated by the ideal strategy presented in the previous subsection.
We denote the full correlation by $\fC(d,r) = \{( \tpr{ab}{xy} )_{a,b}\}_{x,y}$.
To help readers to understand the proofs in \cref{sec:main} better, 
parts of the full correlation are given in the tables below.

\begin{table}[H]
\begin{center}
\begin{tabular}{|c|c||c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{} &
\multicolumn{3}{|c|}{$x=\nr+1$}&
\multicolumn{3}{|c|}{$x=\nr+2$} \\
\cline{3-8}
\multicolumn{2}{|c|}{} &
$a = 0$ & $a=1$ & $a=2$ &
$a = 0$ & $a=1$ & $a=2$\\
\hline
\hline
\multirow{2}{*}{$y = 1$} & $b=0$ & $\frac{\cos^2(\pi/2d)}{d-1}$ & $\frac{\sin^2(\pi/2d)}{d-1}$ & \small $\pr{20}{00}$ 
& $\frac{1+\sin(\pi/d)}{2(d-1)}$ & $\frac{1-\sin(\pi/d)}{2(d-1)}$ & \small  $\pr{20}{10}$ \\
\cline{2-8}
&$b=1$ & $\frac{\sin^2(\pi/2d)}{d-1}$ & $\frac{\cos^2(\pi/2d)}{d-1}$ & $\frac{d-3}{d-1}-\pr{20}{00}$ 
&  $\frac{1-\sin(\pi/d)}{2(d-1)}$ & $\frac{1+\sin(\pi/d)}{2(d-1)}$ & \small $\frac{d-3}{d-1} - \pr{20}{10}$  \\
\hline
\multirow{2}{*}{$y = 2$} & $b=0$ & $\frac{\cos^2(\pi/2d)}{d-1}$ & $\frac{\sin^2(\pi/2d)}{d-1}$ & \small $\pr{20}{01}$ & 
$ \frac{1-\sin(\pi/d)}{2(d-1)}$ & $ \frac{1+\sin(\pi/d)}{2(d-1)}$ & \small $\pr{2 0}{11}$  \\
\cline{2-8}
&$b=1$ & $\frac{\sin^2(\pi/2d)}{d-1}$ & $\frac{\cos^2(\pi/2d)}{d-1}$ & \small $\frac{d-3}{d-1}-\pr{20}{01}$ &  
$ \frac{1+\sin(\pi/d)}{2(d-1)}$ & $ \frac{1-\sin(\pi/d)}{2(d-1)}$ & \small $\frac{d-3}{d-1}- \pr{20}{11}$ \\
\hline
\end{tabular}
\end{center}
\caption{The correlation for $x \in \{\nr+1, \nr+2\}$ and $y \in \{1,2\}$.}
\label{tb:chsh}
\end{table}
We don't explicitly calculate the conditional probabilities of the form $\pr{20}{xy}$ when $x \in \{\nr+1, \nr+2\}$ and $y \in \{1, 2\}$ 
because they are irrelevant.
Note that when $x \in \{1,2\}$ and $y \in \{\nr+1, \nr+2\}$, the correlation table is 
very similar to Table~\ref{tb:chsh}, so we omit it here.

\begin{table}[H]
\begin{center}
\begin{tabular}{|c|c||c|c|c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{} &
\multicolumn{3}{|c|}{$x=\nr+1$}&
\multicolumn{3}{|c|}{$x=\nr+2$}&
\multicolumn{2}{|c|}{$x=0$}\\
\cline{3-10}
\multicolumn{2}{|c|}{} &
$a = 0$ & $a=1$ & $a=2$ &
$a = 0$ & $a=1$ & $a=2$ &
$a = 0$ & $a =2 $\\
\hline
\hline
\multirow{3}{*}{$y = \nr+1$} & $b=0$ & $\frac{1}{d-1}$ & $0$ & 0 
& $\frac{1}{2d-2}$ & $\frac{1}{2d-2}$ & 0 & $\frac{1}{d-1}$ & 0 \\
\cline{2-10}
&$b=1$ & 0 & $\frac{1}{d-1}$ & $0$ 
&  $\frac{1}{2d-2}$ & $\frac{1}{2d-2}$ & 0 &$\frac{1}{d-1}$ & 0 \\
\cline{2-10}
&$b=2$ & 0 & 0 & $\frac{d-3}{d-1}$ 
&  0 & 0 &  $\frac{d-3}{d-1} $ &0 & $\frac{d-3}{d-1}$ \\
\hline
\multirow{3}{*}{$y = \nr+2$} & $b=0$ & $\frac{1}{2d-2}$ & $\frac{1}{2d-2}$ & 0 
& $\frac{1}{d-1}$ & $0$ & 0 & $\frac{1}{d-1}$ & 0 \\
\cline{2-10}
&$b=1$ & $\frac{1}{2d-2}$ & $\frac{1}{2d-2}$ & $0$ 
&  0 & $\frac{1}{d-1}$ & $0$ &$\frac{1}{d-1}$ & 0 \\
\cline{2-10}
&$b=2$ & 0 & 0 & \small $\frac{d-3}{d-1}$ 
&  0 & 0 & \small $\frac{d-3}{d-1} $ &0 &\small $\frac{d-3}{d-1}$ \\
\hline
\multirow{2}{*}{$y = 0$} & $b=0$ & $\frac{1}{d-1}$ & $\frac{1}{d-1}$ & 0 
& $\frac{1}{d-1}$ & $\frac{1}{d-1}$ & 0 & $\frac{2}{d-1}$ & 0 \\
\cline{2-10}
&$b=2$ & 0 & 0 & $\frac{d-3}{d-1}$ 
&  0 & 0 & \small $\frac{d-3}{d-1} $ &0 & \small $\frac{d-3}{d-1}$ \\
\hline
\end{tabular}
\end{center}
\caption{The correlation for $x ,y\in\{0, \nr+1,\nr+2\} $.}
\end{table}

\begin{table}[H]
\begin{center}
\begin{tabular}{|c|c||c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{} &
\multicolumn{6}{|c|}{$y=(\nr+1, f_0)$}\\
\cline{3-8}
\multicolumn{2}{|c|}{} &
$b = (0,0)$ & $b=(0,1)$ & 
$b = (1,0)$ & $b=(1,1)$ &
$b = (2,0)$ & $b=(2,1)$   \\
\hline
\hline
\multirow{3}{*}{$x = \nr+1$} & $a=0$ & $\frac{1}{2d-2}$ & $\frac{1}{2d-2}$ &  $0$
& $0$ & $0$ & $0$  \\
\cline{2-8}
&$a=1$ & $0$ & $0 $ & $\frac{1}{2d-2}$ 
&  $\frac{1}{2d-2}$ & $0$ & $0$  \\
\cline{2-8}
&$a=2$ & 0 & 0 & $0$ 
&  0 & $\frac{d-3}{2d-2}$ & $\frac{d-3}{2d-2} $  \\
\hline
\multirow{2}{*}{$x = f_0$} & $a=0$ & $\frac{1}{2d-2}$ & $0$ & $\frac{1}{2d-2}$ 
& $0$ & $\frac{d-3}{2d-2}$ & 0  \\
\cline{2-8}
&$a=1$ & $0$ & $\frac{1}{2d-2}$ & $0$ 
&  $\frac{1}{2d-2}$ & $0$ & $\frac{d-3}{2d-2}$  \\
\hline
\end{tabular}
\end{center}
\caption{The correlation for the commutation test for questions $x(f_0)$ and $\nr+1$.}
\label{tbl:comm}
\end{table}
The correlation tables for the other seven pairs
of questions are similar.

%=====================================
\section{Robust self-testing with $\fC(d,r)$}
\label{sec:main}
%=====================================
In this section, we examine the properties of a strategy $\bS$ that can reproduce the
correlation $\fC(d,r)$ approximately and 
prove that the shared state of $\bS$ must be close to the state $\ket{\tpsi}$
up to some local isometry.

We first give the definition of an approximate strategy of a fixed correlation.
\begin{definition}
    Let $\calX, \calY$ be input alphabets,
    $\{\calA_x\}_{x \in \calX}, \{\calB_y\}_{y \in \calY}$
    be families of output alphabets.
	For a correlation $\tilde{C} = (\tpr{ab}{xy})$
	with $(a,b) \in \calA_x \times \calB_y$ and $(x,y) \in \calX \times \calY$,
	and a distribution $\pi$ over all the $(x,y) \in \calX \times \calY$,
	we say a quantum strategy $\bS = (\ket{\psi}, \{M_x^a\}, \{N_y^b\})$
	is an $\ep$-approximate quantum strategy of $\tilde{C}$
	with respect to the distribution $\pi$, 
	if it produces the correlation
	$( \pr{ab}{xy} = \bra{\psi} M_x^a N_y^b \ket{\psi} )$ such that
	\begin{align}
	\label{eq:expect}
	\E_{(x,y) \sim \pi} \left[\sum_{ab} \abs{\pr{ab}{xy} - \tpr{ab}{xy}}\right] \leq \ep.
	\end{align}
\end{definition}
In our full test, the probability distribution over question pairs is $\pi$ given in 
\cref{def:full_test}.
In the rest of this section, when we say some strategy can approximate
$\fC(d,r)$, we mean that the approximation is with respect to 
this distribution $\pi$.

The number of all the possible question pairs is 
$N = \Theta(r)$.
Considering that
the number of possible output pairs is constant for each question pair $(x,y)$, 
the approximate condition in \cref{eq:expect} directly implies that, if $\bS$ is an $\ep$-approximate strategy
of $\fC(d,r)$, then
\begin{align*}
    \pr{ab}{xy} \appd{O(r\ep)} \tpr{ab}{xy},
\end{align*}
for all valid $(a,b,x,y)$ of the full test.

In the next three subsections, we analyze the implication of an approximate strategy 
for the correlations of the 
three components introduced in \cref{sec:tests}.

%------------------------------------------------------------------------------------------
\subsection{Implication of approximately winning $\LS(r)$}
\label{sec:imp_lct}
%------------------------------------------------------------------------------------------
In this subsection, we study the implications of a strategy that can win $\LS(r)$ with probability close to $1$. We prove a series of 
propositions, which lead us to \cref{prop:lct_adv}.
\Cref{prop:lct_adv} is a key step of the self-testing proof in \cref{sec:self-test}.

Recall that in $\LS(r)$ the question pair is denoted by $(I_i,x(s))$ where $I_i$ 
labels an equation and $x(s)$ is a variable for some $s \in I_i$. 
We denote 
Alice's observable for variable $x(s)$ in equation $i$ by $M_i(s)$ and
Bob's observable for variable $x(s)$ by $N(s)$.

The first implication of an approximate winning strategy is given in the proposition below.
\begin{proposition}
	\label{prop:lct_base}
	Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	Define $M(s) := M_i(s)$ where 
	$i$ is the smallest among all $i$'s
	such that $x(s) \in I_i$.
	Then, 
	\begin{align}
		&M(s) N(s) \ket{\psi} \appd{O(\sr\se)} \ket{\psi} \\
		&\Pi_{s \in I_i} M(s) \ket{\psi} \appd{O(\sr\se)}(-1)^{c(i)} \ket{\psi},
	\end{align}
	for all $s \in S$.
\end{proposition}

\begin{proof}
The condition that $(\tpr{\ba b}{I_i s})$ of $\fC(d,r)$ wins the linear system game optimally is equivalent to
\begin{align*}
\sum_{\ba,b: \ba \in X_i, a(j) = b} \tpr{\ba b}{I_i s} = 1
\end{align*}
for all $i = 1 \dots \mr$.
We also know that the shared state $\ket{\psi}$ and observables $M_i(s)$
and $N(s)$ of $\bS$ satisfy the relation
\begin{align*}
	\bra{\psi} M_i(s) N(s) \ket{\psi} &= 2 \sum_{\ba,b: \ba \in X_i, a(s) = b} \pr{\ba b}{I_i s} -1,
\end{align*}
as proved in \cite{slofstra2017}.
Using the fact that $\bS$ approximates the winning correlation of $\LS(r)$, 
we can further see that
\begin{align*}
	\bra{\psi} M_i(s) N(s) \ket{\psi} \appd{O(r \ep)} 2\sum_{\ba,b: \ba \in X_i, a(j) = b} \tpr{\ba b}{is} - 1 = 1.
\end{align*}
Since $M_i(s)$ and $N(s)$ are binary observables, the equation above
implies that
\begin{align*}
    \norm{M_i(s)\ket{\psi} - N(s)\ket{\psi}}^2 
    = 2 - 2 \bra{\psi}M_i(s)N(s) \ket{\psi} 
    \leq 2 - 2(1 - O(r\ep)) = O(r\ep),
\end{align*}
or equivalently,
\begin{align*}
	M_{i}(s) \ket{\psi} \appd{O(\sr\se)} N(s) \ket{\psi},
\end{align*}
for each $1 \leq i \leq \mr$ and generator $s$ of
$\Gamma$.
Recall that $M(s) := M_{i}(s)$ for the smallest $i$ such that $ s \in I_i$.
Therefore, 
\begin{align*}
	M(s) N(s) \ket{\psi} & = M_{i}(s)N(s)\ket{\psi} \\
	    & = N(s) M_{i}(s) \ket{\psi} \\
	    &\appd{O(\sr\se)} N(s)^2 \ket{\psi} \\
	    &= \ket{\psi}
\end{align*}
for all $s \in S$.

The fact that $\bS$ is an $\ep$-approximate strategy also tells us that
Alice gives satisfying assignment for each equation with probability $1 - O(r\ep)$,
or equivalently,
\begin{align*}
	\Pi_{s \in I_i} M_{i}(s) \ket{\psi} \appd{O(\sr\se)} (-1)^{c(i)} \ket{\psi}
\end{align*}
for $1 \leq i \leq \mr$.
If we replace each $M_{i}(s)$ with $M(s)$ in the equation above, we demonstrate the effect on a
general case that $I_i = \{s_1, s_2, s_3\}$,
\begin{align*}
	&\norm{ M_{i}(s_1)M_{i}(s_2)M_{i}(s_3) \ket{\psi} - M(s_1)M(s_2)M(s_3) \ket{\psi}}\\
	\leq & \norm{M_{i}(s_1)M_{i}(s_2)M_{i}(s_3)\ket{\psi} - M(s_1)M_{i}(s_2)M_{i}(s_3) \ket{\psi}}\\
	&+ \norm{M(s_1)M_{i}(s_2)M_{i}(s_3) \ket{\psi} - M(s_1)M(s_2)M_{i}(s_3) \ket{\psi}} \\
	&+\norm{M(s_1)M(s_2)M_{i}(s_3) \ket{\psi} - M(s_1)M(s_2)M(s_3) \ket{\psi}}\\
	\leq &3O(\sr\se) = O(\sr\se),
\end{align*}
where we bound each term in line $2,3$ and $4$ by \cref{lm:sub}.
Hence we can conclude that 
\begin{align*}
	\Pi_{s \in I_i} M(s) \ket{\psi} \appd{O(\sr\se)} (-1)^{c(i)}\ket{\psi}
\end{align*}
for $1 \leq i \leq \mr$.
Note that with the same technique, we can prove that 
\begin{align*}
	\Pi_{s \in I_i} N(s) \ket{\psi} \appd{O(\sr\se)} (-1)^{c(i)}\ket{\psi}  
\end{align*}
for $1 \leq i \leq \mr$.
\end{proof}

To make the following derivations easier to understand, 
we introduce a new terminology.
\begin{definition}
    Let $S$ be a set of generators, and $\{s_i\}_{i=1}^n$ and $\{s'_j\}_{j=1}^m$
    be two sequences of $S$.
    Define a relation $r$, whose equation form is $\Pi_{i=1}^n s_i = \Pi_{j=1}^m s'_j$.
    We say operators $\{M(s) : s \in S\}$ $\ep$-approximately satisfy the relation
    $r$ on the state $\ket{\psi}$ if 
    \begin{align*}
        \Pi_{i=1}^n M(s_i)\ket{\psi} \appd{\ep} \Pi_{j=1}^m M(s'_j) \ket{\psi}. 
    \end{align*}
\end{definition}
Using the new terminology, \cref{prop:lct_base} can be rephrased 
as the following proposition.
\begin{proposition}
    \label{prop:lct_group}
    Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	Then the observables $\{ M(s) \}_{s \in S}$ and
	$\{N(s)\}_{s \in S}$ defined in \cref{prop:lct_base}, 
	$O(\sr \se)$-approximately satisfy
	each relation $t \in R$ on the shared state $\ket{\psi}$, 
	where $S$ is the generator set of
	$\Gamma$ and
	and $R$ is the relation set of $\Gamma$.	
\end{proposition}
Recall that the group $\Pg_1$ (\cref{def:g1}) is embedded in $\Gamma$,
as proved in \cref{prop:g1gamma}. 
Following the embedding procedure, we can prove the next proposition.
\begin{proposition}
    \label{prop:lct_pg1}
    Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	Then the observables $\{ M(s) \}_{s \in S_1}$ and
	$\{N(s)\}_{s \in S_1}$, defined in 
\cref{prop:lct_group}, $O(\sr\se)$-approximately satisfy
	each relation $t \in R_1$ on $\ket{\psi}$, 
	where $S_1$ is the generator set of $\Pg_1$
    and $R_1$ is the relation set of $\Pg_1$.
\end{proposition}

\begin{proof}
    We prove 
    \begin{align}
        \label{eq:f0b1}
        M(f_0)M(b_1)M(f_0) \ket{\psi} \appd{O(\sr\se)} M(c_1)\ket{\psi}
    \end{align}
    to demonstrate the techniques.
    From the fact that Alice's observables $O(\sr\se)$-approximately
    satisfy the linear relations of $\Gamma$, we can get that
    \begin{align*}
        &M(f_0) \ket{\psi} \appd{O(\sr \se)} M(f_1)M(f_2)\ket{\psi} \\
        &M(b_1) \ket{\psi} \appd{O(\sr \se)} M(f_2)M(p_{1,1}) \ket{\psi}\\
        &M(f_0) \ket{\psi} \appd{O(\sr \se)} M(p_{1,3})M(p_{1,4}) \ket{\psi}
    \end{align*}
    Then we expand $M(f_0)M(b_1)M(f_0)\ket{\psi}$
    using the three equations above and \cref{lm:sub},
    \begin{align*}
        M(f_0)M(b_1)M(f_0) \ket{\psi} 
        &\appd{O(\sr \se)}M(f_1)M(f_2)M(b_1)M(f_0) \ket{\psi} \\
        &\appd{O(\sr \se)} M(f_1)M(f_2)M(f_2)M(p_{1,1})M(f_0) \ket{\psi} \\
        & = M(f_1)M(p_{1,1})M(f_0) \ket{\psi} \\
        & \appd{O(\sr \se)} M(f_1)M(p_{1,1})M(p_{1,3})M(p_{1,4}) \ket{\psi}.
    \end{align*}
    In the rest of the proof, we skip steps of applying \cref{lm:sub}.
    Next, we can simplify $M(f_0)M(b_1)M(f_0) \ket{\psi}$ as
    \begin{align*}
        M(f_0)M(b_1)M(f_0) \ket{\psi}
        &\appd{O(\sr \se)} M(f_1)M(p_{1,1})M(p_{1,3})M(p_{1,4}) \ket{\psi}\\
        &\appd{O(\sr \se)} M(f_1)M(p_{1,2})M(p_{1,4})\ket{\psi} \\
        &\appd{O(\sr \se)} M(p_{1,5})M(p_{1,4})\ket{\psi} \\
        & \appd{O(\sr \se)} M(c_1) \ket{\psi},
    \end{align*}
    where from the second line to the third line, we used
    the relation $M(p_{1,1})M(p_{1,2})M(p_{1,3})\ket{\psi} \appd{O(\sr \se)} \ket{\psi}$, and from the third line to the
    fourth line, we used the relation 
    $M(c_1)M(p_{1,4})M(p_{1,5})\ket{\psi} \appd{O(\sr \se)} \ket{\psi}$.

    The derivation of other relations in $R_1$ from relations in $R$
    involves the same number of substitutions and uses similar techniques, so we omit it here.
\end{proof}
Essentially the proof above follows the proof of 
\cref{prop:g1gamma},
so we can follow the proof of \cref{prop:g0g1}
to prove the next proposition.
\begin{proposition}
    \label{prop:lct_pg0}
    Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	Then the observables $\{ M(s) \}_{s \in S_0}$ and 
	$\{N(s)\}_{s \in S_0}$, defined in 
\cref{prop:lct_group}, $O(\sr \se)$-approximately satisfy
	each relation $t \in R_0$ on the shared state $\ket{\psi}$, 
	where $S_0$ is the generator set of $\Pg_0$
    and $R_0$ is the relation set of $\Pg_0$.
\end{proposition}
Following the steps of the embedding of $(a_3a_4)^l a_1a_2(a_4a_3)^l = (a_1a_2)^{r^l}$ into $\Gamma$ presented in
the proof of \cref{prop:g0},
we can prove the next lemma.
\begin{lemma}
\label{prop:lct_adv}
Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$ and $\{M(s), N(s)\}_{s \in S}$ be the observables defined in 
\cref{prop:lct_group}. 
Then,
\begin{align}
	&M(a_1)M(a_2) (M(a_4)M(a_3))^l \ket{\psi}\appd{O(r^{l+1/2} \se)} (M(a_4)M(a_3))^l (M(a_1)M(a_2))^{r^l} \ket{\psi}, \\
	&N(a_1)N(a_2) (N(a_4)N(a_3))^l \ket{\psi}\appd{O(r^{l+1/2} \se)} (N(a_4)N(a_3))^l (N(a_1)N(a_2))^{r^l} \ket{\psi}, 
\end{align}
for all $l \geq 1$.
\end{lemma}
\begin{proof}
Proving the $l=1$ case is essentially proving that observables 
$\{M(a_i)\}_{i=1}^4$ and $\{N(a_i)\}_{i=1}^4$
$O(r^{3/2} \se)$-approximately satisfy the relation 
\begin{align}
    \label{eq:a1a2r}
    (a_3a_4)(a_1a_2)(a_4a_3) = (a_1a_2)^r
\end{align}
on the shared state $\ket{\psi}$.
The derivation follows the proof of \cref{prop:g0}.
The $O(r^{3/2}\se)$ approximation error comes from the fact that 
the derivation of \cref{eq:a1a2r} has $\Theta(r)$ substitution steps, where each substitutes uses
a different relation of $R_0$. 
So we have
\begin{align}
	\label{eq:ls_comm_rel}
	M(a_1)M(a_2) (M(a_4)M(a_3)) \ket{\psi} \appd{O(r^{3/2}\se)} (M(a_4)M(a_3))(M(a_1)M(a_2))^r \ket{\psi}.
\end{align}
Assume the statement is true for $l = n$ and consider the case that $l = n+1$,
\begin{align*}
	M(a_1)M(a_2) (M(a_4)M(a_3))^{n+1} \ket{\psi}  &\appd{O(\sr \se)} (N(a_3)N(a_4)) M(a_1)M(a_2)(M(a_4)M(a_3))^n \ket{\psi} \\
	&\appd{O(r^{n+1/2}\se)} (N(a_3)N(a_4)) (M(a_4)M(a_3))^n (M(a_1)M(a_2))^{r^n} \ket{\psi}\\
	&\appd{O(\sr \se)} (M(a_4)M(a_3))^n (M(a_1)M(a_2))^{r^n} (M(a_4)M(a_3)) \ket{\psi} \\
	& \appd{O(r^{n+3/2} \se)} (M(a_4)M(a_3))^{n+1} (M(a_1)M(a_2))^{r^{n+1}} \ket{\psi},
\end{align*}
where from the second last to the last line we repeat the process of applying \cref{eq:ls_comm_rel} 
and the relation $(M(a_1)M(a_2))^r \ket{\psi} \appd{O(r^{3/2}\se)} (N(a_1)N(a_2))^r \ket{\psi}$, $r^n$ times.
In short, we have 
\begin{align}
	M(a_1)M(a_2)(M(a_4)M(a_3))^{n+1} \ket{\psi} \appd{O(r^{n+3/2} \se)} (M(a_4)M(a_3)^{n+1} (M(a_1)M(a_2))^{r^{n+1}}\ket{\psi}.
\end{align}
By the principal of induction, the proof is complete.
The proof for Bob's observables $\{ N(s) \}_{s \in S}$ is analogous.
\end{proof}
%An informal justification of the corollary is the following calculation
%\begin{align}
%	M_1M_2(M_4M_3)^k \ket{\psi} &\appd{O((k-1)r\se)} (N_3N_4)^{k-1} M_1M_2(M_4M_3) \ket{\psi} \\
%	&\appd{O(r\se)} (N_3N_4)^{k-1}(M_4M_3) (M_1M_2)^r \ket{\psi} \\
%	&\appd{O(r\se)} (N_4N_4)^{k-2} (M_4M_3) (M_1M_2)^{r} (M_4M_3) \ket{\psi}\\
%	&\appd{O(r^2 \se)} (N_4N_3)^{k-2} (M_4M_3)^2 (M_1M_2)^{r^2} \ket{\psi} \\
%	& \ldots (k-2  \text{ more such swaps} ) \ldots \\
%	&\appd{O(r^{k-1}\se)} (M_4M_3)^{k} (M_1M_2)^{r^{k}}\ket{\psi}.
%\end{align}
%The total distance introduced by the whole procedure above is
%\begin{align}
%	\sum_{j=0}^{k-1} O(r^j \se) = O(r^k \se).
%\end{align}

We define 
\begin{align}
    \label{eq:uaoa} & O_A := M(a_1)M(a_2) && U_A := M(a_3)M(a_4) \\
    \label{eq:ubob} & O_B := N(a_1)N(a_2) && U_B := N(a_3)N(a_4).
\end{align}
The conclusion of \cref{prop:lct_adv} can be rewritten as 
\begin{align*}
	 O_A (U_A\ct)^n \ket{\psi} \appd{O(r^{n+1/2} \se)} (U_A\ct)^n O_A^{r^n} \ket{\psi}.
\end{align*}
On Bob's side, the similar relation holds.

The other key relations that are embedded in $\Gamma$ are 
$f_0g_0 = Jg_0f_0$ and $f_2g_2 = Jg_2 f_2$.
We summarize the implications in the following proposition.
\begin{proposition}
	\label{prop:lct_comm}
	Let $\bS$, $\{M(s), N(s)\}_{s \in S}$ be the strategy and observables defined in \cref{prop:lct_group}.
	Then,
	\begin{align}
		&M(f_0)M(g_0) \ket{\psi} \appd{O(\sr \se)} -M(g_0)M(f_0) \ket{\psi} \\
		&M(f_2)M(g_2) \ket{\psi} \appd{O(\sr \se)} -M(g_2)M(f_2) \ket{\psi}.
	\end{align}
\end{proposition}
Following the substitution steps to prove \cref{eq:mg1} and \cref{eq:mg2} in the proof of \cref{prop:g1gamma},
this proposition can be proved with the same techniques used in the proof of \cref{prop:lct_group}.

%------------------------------------------------------------------------------------------
\subsection{Implication of the extended weighted CHSH test}
\label{sec:imp_chsh}
%------------------------------------------------------------------------------------------
In this subsection, we prove a series of propositions and lemmas 
derived from the fact that some strategy can approximate the ideal correlation of the extended weighted CHSH test,
which lead to a key result that will be
used in the self-testing proof in \cref{sec:self-test}.

We will work with a strategy $\bS$ that can approximate
$\fC(d,r)$, which contains
the ideal correlation of the extended weighted CHSH test.
About the notation,
for $(x,y) \in \calX^{\EXT} \times \calY^{\EXT}$ and $(a,b) \in \calA_x^{\EXT} times \calB_y^{\EXT}$, 
Alice and Bob's projectors are denoted 
by $M_x^a$ and $N_y^b$,
as illustrated in \cref{sec:prelim}.
Since the two answers for question $1$ and $2$ are $0$ and $1$, 
we denote Alice and Bob's binary observables for these two questions
by $M_1, M_2$ and $N_1, N_2$ respectively.
Note that the observables $M_1 \equiv M(a_1)$ and $M_2 \equiv M(a_2)$,
where $M(a_1)$ and $M(a_2)$ are used in $\LS(r)$,
because questions $1$ and $2$ are equivalent to questions $x(a_1)$ and $x(a_2)$ in the full test.  
Similarly, on Bob's side, we have $N_1 \equiv N(a_1)$ and $N_2 \equiv N(a_2)$.
For simplicity, in this subsection, we
use $M_1, M_2$ and $N_1, N_2$ only.

\begin{proposition}
\label{prop:chsh_base}
Let $\bS$ be an $\ep$-approximate quantum strategy of $\fC(d,r)$.
Then the strategy $\bS$ satisfies the following relations
\begin{align*}
	  M_0^0 N_0^0 \ket{\psi}
	  &\appd{O( \sr \se)}M_0^0 \ket{\psi}\\
	  &\appd{O( \sr \se)} (M_{\nr+1}^0+M_{\nr+1}^1)\ket{\psi}\\
	  &\appd{O( \sr \se)} (M_{\nr+2}^0+M_{\nr+2}^1)\ket{\psi} \\
	  &\appd{O( \sr \se)} N_0^0 \ket{\psi}\\
	  &\appd{O( \sr \se)} (N_{\nr+1}^0+N_{\nr+1}^1)\ket{\psi}\\
	  &\appd{O( \sr \se)} (N_{\nr+2}^0+N_{\nr+2}^1)\ket{\psi}\\
	M_0^2 N_0^2\ket{\psi}&\appd{O( \sr \se)} M_0^2 \ket{\psi}\\
	&\appd{O( \sr \se)} M_{\nr+1}^2 \ket{\psi}\\
	&\appd{O( \sr \se)} M_{\nr+2}^2 \ket{\psi}\\
	&\appd{O( \sr \se)} N_0^2 \ket{\psi}\\
	&\appd{O( \sr \se)} N_{\nr+1}^2 \ket{\psi}\\
	&\appd{O( \sr \se)} N_{\nr+2}^2 \ket{\psi},
\end{align*}
and
\begin{align*}
&M_{\nr+1}^0 \ket{\psi} \appd{O( \sr \se)} N_{\nr+1}^0 \ket{\psi} && M_{\nr+1}^1\ket{\psi} \appd{O( \sr \se)} N_{\nr+1}^1 \ket{\psi}\\
&M_{\nr+2}^0 \ket{\psi} \appd{O( \sr \se)} N_{\nr+2}^0 \ket{\psi} && M_{\nr+2}^1\ket{\psi} \appd{O( \sr \se)} N_{\nr+2}^1 \ket{\psi}.
\end{align*}
\end{proposition}

\begin{proof} 
We prove $(M_{\nr+1}^0+M_{\nr+1}^1)\ket{\psi} \appd{O(\sr \se)} N_0^0 \ket{\psi}$ as an example.
The rest of the relations follow the same line of argument.

Recall that the implication of $\bS$ being an $\ep$-approximate strategy is that 
$\pr{ab}{xy} \appd{O(r \ep)} \tpr{ab}{xy}$ for each question 
pair $(x,y)$ and its associating answer pair $(a,b)$.
From the correlation, we can observe that
\begin{align*}
 &\bra{\psi} (M_{\nr+1}^0+M_{\nr+1}^1)N_0^0 \ket{\psi} \\
 = &\pr{0,0}{\nr+1,0} + \pr{1,0}{\nr+1,0}\\
 \geq &\frac{2}{d-1} - O(r \ep).
\end{align*}
On the other hand, the correlation also tells us that 
\begin{align*}
	&\bra{\psi} N_0^0 \ket{\psi} = \pb{0}{0} \appd{O(r\ep)} \frac{2}{d-1}\\
 &\bra{\psi}(M_{\nr+1}^0+M_{\nr+1}^1)\ket{\psi} 
 = \pa{0}{\nr+1} + \pa{1}{\nr+1}
 \appd{O(r\ep)} \frac{2}{d-1},
\end{align*}
The equations above imply that 
\begin{align*}
	&\norm{(M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi} - N_0^0 \ket{\psi}}^2 \\
	= &\bra{\psi} N_0^0 \ket{\psi}+\bra{\psi}(M_{\nr+1}^0+M_{\nr+1}^1)\ket{\psi} - 2\bra{\psi}N_0^0(M_{\nr+1}^0 + M_{\nr+1}^1)N_0^0 \ket{\psi}\\
	= &O(r \ep),
\end{align*}
where we used the fact that $(M_{\nr+1}^0 + M_{\nr+1}^1)^2 = M_{\nr+1}^0 + M_{\nr+1}^1$.
Hence we can conclude that $(M_{\nr+1}^0 + M_{\nr+1}^1)\ket{\psi} \appd{O(\sr \se)} N_0^0\ket{\psi}$.
%From the correlation, we know that 
%\begin{align}
%	\bra{\psi} M_\ast^\perp N_\ast^\perp \ket{\psi} = \frac{d-3}{d-1} - O(r^2 \ep).
%\end{align}
%In the mean time, we also have 
%$\norm{M_\ast^\perp \ket{\psi}}^2 \appd{O(r^2\ep)} \norm{N_\ast^\perp\ket{\psi}}^2 \appd{O(r^2\ep)} (d-3)/(d-1)$.
%The two facts above imply that 
%\begin{align}
%	\norm{M_\ast^\perp \ket{\psi} - N_\ast^\perp \ket{\psi}}^2 = \norm{M_\ast^\perp \ket{\psi}}^2 + \norm{N_\ast^\perp \ket{\psi}}^2
%	- 2 \bra{\psi} M_\ast^\perp N_\ast^\perp \ket{\psi}  = O(r^2 \ep),
%\end{align}
%or equivalently,
%\begin{align}
%	M_\ast^\perp \ket{\psi} \appd{O(r\se)} N_\ast^\perp \ket{\psi}.
%\end{align}
\end{proof}

\begin{proposition}
	\label{prop:s1s2}
	Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$. We define
	\begin{align}
		&M_{\nr+1} = M_{\nr+1}^0 - M_{\nr+1}^1  && 
		M_{\nr+2} = M_{\nr+2}^0 - M_{\nr+2}^1,\\
		&N_{\nr+1} = N_{\nr+1}^0 - N_{\nr+1}^1  && 
		N_{\nr+2} = N_{\nr+2}^0 - N_{\nr+2}^1.
	\end{align}
	Then,
	\begin{align}
		\bS_1 = \left(M_{\nr+1}, M_{\nr+2}, N_1, N_2, \frac{M_0^0\ket{\psi}}{\norm{M_0^0\ket{\psi}}}\right),\\
		\bS_2 = \left(N_{\nr+1}, N_{\nr+2}, M_1, M_2, \frac{M_0^0\ket{\psi}}{\norm{M_0^0\ket{\psi}}}\right).
	\end{align}
	are two $O(d \sr \se)$-approximate strategies of the ideal correlation that achieves $\I_{\cot(-\pi/d)}^{\max}$.
\end{proposition}
Note that here we only consider $\ep$ small enough such that 
$\norm{M_0^0\ket{\psi}} > 0$.
\begin{proof}
We first prove that $\bS_1$ is an 
$O(d \sr \se)$-approximate strategies of $\I_{\cot(-\pi/d)}^{\max}$.
This step requires examining the terms of the form
$\bra{\psi} M_0^0 M_{x}^a N_y^b M_0^0\ket{\psi}$ for $x = \nr+1,\nr+2$, $y=1,2$ and $a,b = 0,1$.
We find that these terms relate to $\bra{\psi} M_x^a N_y^b \ket{\psi}$ by
\begin{align*}
	   \bra{\psi} M_{x}^aN_y^b \ket{\psi} 
	&= \bra{\psi}(M_0^0 + M_0^2) M_x^a N_y^b (M_0^0 + M_0^2)\ket{\psi} \\
	&=  \bra{\psi}M_0^0 M_x^aN_y^b M_0^0\ket{\psi} + \bra{\psi}M_0^0 M_x^aN_y^b M_0^2\ket{\psi}\\
	&\quad\quad+\bra{\psi}M_0^2 M_x^aN_y^b M_0^0\ket{\psi} + \bra{\psi}M_0^2 M_x^aN_y^b M_0^2\ket{\psi}\\
	&\appd{O(\sr\se)}  \bra{\psi}M_0^0 M_x^aN_y^b M_0^0\ket{\psi} + \bra{\psi}M_0^0 M_x^aN_y^b M_x^2\ket{\psi} \\
	&\quad \quad +\bra{\psi}M_x^2 M_x^aN_y^b M_0^0\ket{\psi} + \bra{\psi}M_x^2 M_x^aN_y^b M_x^2\ket{\psi}\\
	&=\bra{\psi}M_0^0 M_x^aN_y^b M_0^0\ket{\psi},
\end{align*}
where we use the facts that $M_0^2 \ket{\psi} \appd{O(\sr \se)} M_x^2 \ket{\psi}$ and that 
$M_x^0 M_x^2 = 0$ for the relevant values of $(x,y,a,b)$.

To demonstrate the effect of this relation, we use an example.
If Alice and Bob share state $M_0^0 \ket{\psi}/\norm{M_0^0 \ket{\psi}}$ and apply $M_{\nr+1}^0N_1^0$,
then
\begin{align*}
	\frac{\bra{\psi}M_0^0 M_{\nr+1}^0N_1^0 M_0^0\ket{\psi}}{\bra{\psi} M_0^0 \ket{\psi}}  
	\appd{O(d \sr \se)} \frac{\bra{\psi} M_{\nr+1}^0N_1^0 \ket{\psi}}{\bra{\psi} M_0^0 \ket{\psi}}
	\appd{O(d r \ep)}
	\frac{\cos^2(-\pi/2d)}{2},
\end{align*} 
where the first error term is derived from
\begin{align*}
	\frac{ \norm{\bra{\psi}M_0^0 M_{\nr+1}^0N_1^0 M_0^0\ket{\psi} -  \bra{\psi} M_{\nr+1}^0N_1^0 \ket{\psi}}}
	{\bra{\psi} M_0^0 \ket{\psi}} \leq \frac{O(\sr\se)}{\frac{2}{d-1} - O(r \ep)} = O(d \sr \se).
\end{align*}
The rest of the conditional probabilities produced by $\bS_1$ can be derived in the same way.
Using the fact that $O(d \sr \se)$ dominates $O(d r \ep)$,
we can prove that $\bS_1$ is an $O(d \sr\se)$-approximate strategy of the ideal correlation that achieves $\I_{\cot(-\pi/d)}^{\max}$.
The proof for $\bS_2$ is analogous so we omit it here.
\end{proof}

Based on \cref{prop:chsh_base,prop:s1s2}, we identify a special sub-normalized state $\ket{\psi_1}$, 
\begin{align}
    \label{eq:psi1}
	\ket{\psi_1} =1/2(M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 +M_{\nr+1}^1) \ket{\psi}.
\end{align}
The special properties of $\ket{\psi_1}$ are summarized
in the propositions below. 
Based on the properties of $\ket{\psi_1}$, we will prove that $\ket{\psi_1}$ can be used to approximate $\ket{\psi}$
in \cref{lm:decomp_psi}.
\begin{proposition}
    \label{prop:psi1_norm}
	Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$ and $\ket{\psi_1}$ be as defined in
	\cref{eq:psi1}.
	Then, 
	\begin{align}
	 %\text{ and } \ket{\psi_2} \appd{O(\sd \sr\qe)} \ket{\psi_4}, \\
	 %\label{eq:psi1_sum} &M_\ast^0 \ket{\psi} \appd{O( r\se)} \ket{\psi_1}+\ket{\psi_2}, \\
	 %\label{eq:psi1_orth} &\braket{\psi_2}{\psi_1} \appd{O( r\se)} 0,\\
	 \label{eq:psi1_norm} &\norm{\ket{\psi_1}}^2 \appd{O(\sr\se)} 1/(d-1).
	\end{align}
\end{proposition}
\begin{proof}
	We start by expanding the inner product
	\begin{align*}
		&\braket{\psi_1}{\psi_1} \\
		=&1/4\bra{\psi}(M_{\nr+1}^0 - iM_{\nr+1}^1M_{\nr+2} + iM_{\nr+1}^0M_{\nr+2} +M_{\nr+1}^1)\\
		&(M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 -iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1)\ket{\psi}\\
		=&1/2\bra{\psi}(M_{\nr+1}^0+M_{\nr+1}^1-iM_{\nr+1}^1M_{\nr+2}M_{\nr+1}^0+iM_{\nr+1}^0M_{\nr+2}M_{\nr+1}^1\ket{\psi}
	\end{align*}
	Here we apply the trick to flip $M_{\nr+1}^1$ and $M_{\nr+1}^0$ to Bob's side and get 
	\begin{align*}
		\bra{\psi}M_{\nr+1}^1M_{\nr+2}M_{\nr+1}^0\ket{\psi} 
		&\appd{O(\sr\se)} \bra{\psi}N_{\nr+1}^1M_{\nr+2}N_{\nr+1}^0\ket{\psi} \\
		&= \bra{\psi}N_{\nr+1}^1N_{\nr+1}^0M_{\nr+2}\ket{\psi} = 0.
	\end{align*}
	Similarly, we have $\bra{\psi}M_{\nr+1}^0M_{\nr+2}M_{\nr+1}^1\ket{\psi} \appd{ O(\sr \se)} 0$.
	Hence, from the ideal correlation, we get 
	\begin{align}
		\braket{\psi_1}{\psi_1} \appd{O(\sr\se)} 1/2\bra{\psi}(M_{\nr+1}^0+M_{\nr+1}^1\ket{\psi} \appd{O(r\ep)} 
		 \frac{1}{d-1}.
	\end{align}
\end{proof}
\begin{proposition}
    \label{prop:psi1_eig1}
	Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$ and $\ket{\psi_1}$ be as defined in
	\cref{eq:psi1}.
	Then,
	\begin{align}
	    \label{eq:psi1_eig1} 
	    &N_1N_2 \ket{\psi_1} \appd{O(\qr \qe/\sd)} \omega_d\ket{\psi_1}.
	    %&N_2N_1 \ket{\psi_1} \appd{O(\sr \qe/\sd)} \omega_d^{-1} \ket{\psi_1}.
	\end{align}
\end{proposition}
The idea of this proof is to apply the general results in \cref{sec:chsh}
to the strategy $\bS_1$,
so in the first half of the proof, we can derive how $N_1$ and $N_2$
act on the state $M_{\nr+1}^0 \ket{\psi}$ and $M_{\nr+1}^1 \ket{\psi}$.
In the second half, we combine $M_{\nr+1}^0 \ket{\psi}$ and $M_{\nr+1}^1 \ket{\psi}$ in a way and show that it is an eigenvector of $N_1N_2$.
It turns out that this combination of $M_{\nr+1}^0 \ket{\psi}$ and $M_{\nr+1}^1 \ket{\psi}$
is closely related to $\ket{\psi_1}$.
We show all the calculations to make the proof easier to read. 
\begin{proof}
	Following the convention in \cref{sec:chsh}, we define
	\begin{align*}
		& c := \cos(\frac{\pi}{d}) && s := \sin(\frac{\pi}{d}) \\
		&Z_A  := M_{\nr+1} && X_A := M_{\nr+2} \\
		&Z_B := \frac{N_1+N_2}{2c} && X_B := \frac{N_1-N_2}{-2s}.
	\end{align*}
%	Note that $Z_A$ should be defined by $M_{\nr+1}^0 - M_{\nr+1}^1 + M_{\nr+1}^\perp$,
%	but $M_{\nr+1}^\perp$ is almost orthogonal to $M_\ast^0$, so we omit the $M_{\nr+1}^\perp$ term here.
%	In the definition of $X_A$, we omit $M_{\nr+2}^\perp$ too.
	In the first step, we will show how $(N_1 + N_2)/2c$, or $Z_B$,
	acts on $M_{\nr+1}^0\ket{\psi}$ and 
	$M_{\nr+1}^1 \ket{\psi}$.
	Recall that \cref{thm:selftest} applies
	to all approximating strategies pf $\I_{-\cot(\pi/d)}^{\max}$, so
	we can substitute appropriate operators
	and the shared state into
    \cref{eq:za-zb} to see that
	\begin{align*}
		\norm{Z_B M_0^0 \ket{\psi} - Z_AM_0^0 \ket{\psi}} \leq 
		\sqrt{s}\norm{M_0^0 \ket{\psi}} O(\sd \qr \qe)
		= O(\qr \qe/ \sd ).
	\end{align*}
	Since $M_0^0 \ket{\psi} \appd{O(\sr \se)} (M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi}$,
	the equation above can be rewritten as 
	\begin{align*}
		Z_B (M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi} \appd{O(\qr \qe / \sd )} (M_{\nr+1}^0 - M_{\nr+1}^1) \ket{\psi},
	\end{align*}
	where we use the fact that $Z_A = M_{\nr+1}^0 - M_{\nr+1}^1$.
	Using the fact $M_{\nr+1}^0 M_{\nr+1}^1 = 0$, we can 
	further deduce that
	\begin{align*}
		O(\sr \se /d) \geq &\norm{Z_B (M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi} - (M_{\nr+1}^0 - M_{\nr+1}^1) \ket{\psi}}^2\\
		=&\norm{ Z_BM_{\nr+1}^0 \ket{\psi} - M_{\nr+1}^0\ket{\psi}}^2 + \norm{ Z_BM_{\nr+1}^1 \ket{\psi} + M_{\nr+1}^1\ket{\psi}}^2,
	\end{align*}
	which immediately gives us that 
	\begin{align}
	    \label{eq:n1+n20}
		&\frac{N_1+N_2}{2c} M_{\nr+1}^0\ket{\psi} \appd{O( \qr \qe /\sd)} M_{\nr+1}^0 \ket{\psi},\\
		\label{eq:n1+n21}
		&\frac{N_1+N_2}{2c} M_{\nr+1}^1\ket{\psi} \appd{O( \qr \qe /\sd)} -M_{\nr+1}^1 \ket{\psi}.
	\end{align}

	Next, we study how $X_B = (N_1-N_2)/(-2s)$ acts on 
	$M_{\nr+1}^0\ket{\psi}$ and $M_{\nr+1}^1 \ket{\psi}$.
	Substituting the appropriate operators and the shared state into \cref{eq:zaxb,eq:xazb}, we get
	\begin{align*}
		\frac{N_1-N_2}{-2s} M_{\nr+1}^0M_0^0 \ket{\psi} 
		\appd{(\norm{M_0^0\ket{\psi}}/\sqrt{s})\;O(\sd \qr \qe)}M_{\nr+2} M_{\nr+1}^1M_0^0\ket{\psi}\\
		\frac{N_1 -N_2}{-2s} M_{\nr+1}^1M_0^0 \ket{\psi}  
		\appd{(\norm{M_0^0 \ket{\psi}}/\sqrt{s})\;O(\sd \qr \qe)}M_{\nr+2} M_{\nr+1}^0 M_0^0 \ket{\psi}.
	\end{align*}
	Using the facts that $M_0^0 \ket{\psi} \appd{O(\sr \se)} (M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi}$
	and $O(\sd \qr \qe)$ dominates $O(\sr \se)$, we can
	simplify the equations above to get
	\begin{align}
		\label{eq:n1-n20}
		&(N_1-N_2) M_{\nr+1}^0 \ket{\psi} \appd{O(\qr\qe/\sd )} -2sM_{\nr+2} M_{\nr+1}^1 \ket{\psi},\\
		\label{eq:n1-n21}
		&(N_1-N_2) M_{\nr+1}^1 \ket{\psi}\appd{O(\qr\qe/\sd )} -2sM_{\nr+2} M_{\nr+1}^0 \ket{\psi}.
	\end{align}

	In summary, we can rearrange terms in \cref{eq:n1+n20,eq:n1+n21,eq:n1-n20,eq:n1-n21} to get
        \begin{align*}
        		&N_1M_{\nr+1}^0 \ket{\psi} \appd{O(\qr \qe/ \sd)} cM_{\nr+1}^0\ket{\psi} - sM_{\nr+2}M_{\nr+1}^1\ket{\psi},\\
        		&N_1M_{\nr+1}^1 \ket{\psi} \appd{O(\qr \qe/ \sd)} -cM_{\nr+1}^1\ket{\psi} - sM_{\nr+2}M_{\nr+1}^0\ket{\psi},\\
        		&N_2M_{\nr+1}^0 \ket{\psi} \appd{O( \qr \qe /\sd)} cM_{\nr+1}^0\ket{\psi} + sM_{\nr+2}M_{\nr+1}^1\ket{\psi},\\
        		&N_2M_{\nr+1}^1 \ket{\psi} \appd{O(\qr \qe/ \sd )} -cM_{\nr+1}^1\ket{\psi} + sM_{\nr+2}M_{\nr+1}^0\ket{\psi}.
        \end{align*}
        Then we can expand $N_1N_2 M_{\nr+1}^0\ket{\psi}$ by first substitute in
        the approximate expression of $N_2 M_{\nr+1}^0\ket{\psi}$ and repeatedly
        substitute in the four equations above
        to get that
        \begin{align*}
        		&\quad N_1N_2 M_{\nr+1}^0 \ket{\psi} \\
		        &\appd{O(\qr \qe/\sd)} (c N_1M_{\nr+1}^0 +sM_{\nr+2}N_1M_{\nr+1}^1)\ket{\psi}\\
        		&\appd{O( \qr \qe/\sd)}(c^2M_{\nr+1}^0- csM_{\nr+2}M_{\nr+1}^1 -csM_{\nr+2}M_{\nr+1}^1 -s^2 M_{\nr+2}^2 M_{\nr+1}^0)\ket{\psi}\\
        		&= (\cos(\frac{2\pi}{d}) M_{\nr+1}^0 - \sin(\frac{2\pi}{d}) M_{\nr+2}M_{\nr+1}^1)\ket{\psi}.
        \end{align*}
        Following similar steps, we can get that
        \begin{align*}
            &\quad N_1N_2 M_{\nr+1}^1 \ket{\psi} \\
		    &\appd{O( \qr \qe/\sd)} (-c N_1M_{\nr+1}^1 +sM_{\nr+2}N_1M_{\nr+1}^0)\ket{\psi}\\
        		&\appd{O(\qr \qe/\sd)}(c^2M_{\nr+1}^1+ csM_{\nr+2}M_{\nr+1}^0 +csM_{\nr+2}M_{\nr+1}^0 -s^2 M_{\nr+2}^2 M_{\nr+1}^1)\ket{\psi}\\
        		&= (\cos(\frac{2\pi}{d}) M_{\nr+1}^1 + \sin(\frac{2\pi}{d}) M_{\nr+2}M_{\nr+1}^0)\ket{\psi}.
        \end{align*}
        Multiplying $M_{\nr+2}$ to 
        the approximate expression of $N_1N_2 M_{\nr+1}^1 \ket{\psi}$,
        we can get that
        \begin{align*}
        N_1N_2 M_{\nr+2}M_{\nr+1}^1 \ket{\psi}\appd{O(\qr \qe/\sd)} 
		\sin(\frac{2\pi}{d}) M_{\nr+1}^0\ket{\psi} + \cos(\frac{2\pi}{d})M_{\nr+2}M_{\nr+1}^1\ket{\psi},
        \end{align*}
        where we use the fact that 
        \begin{align*}
            (M_{\nr+2})^2 M_{\nr+1}^0 \ket{\psi} 
            &= (M_{\nr+2}^0 + M_{\nr+2}^1)M_{\nr+1}^0 \ket{\psi}\\
            &\appd{O(\sr \se)}N_{\nr+1}^0 M_0^0 \ket{\psi}\\
            &\appd{O(\sr \se)}N_{\nr+1}^0 (M_{\nr+1}^0+M_{\nr+1}^1)\ket{\psi} \\
            &\appd{O(\sr \se)}(M_{\nr+1}^0+M_{\nr+1}^1)M_{\nr+1}^0\ket{\psi} \\
            &= M_{\nr+1}^0 \ket{\psi}.
        \end{align*}
    Combining the approximate
    expressions of $N_1N_2M_{\nr+1}^0\ket{\psi}$
    and $N_1N_2 M_{\nr+2} M_{\nr+1}^0\ket{\psi}$ gives us that
	\begin{align}
		\label{eq:omegad}
		&N_1N_2 (M_{\nr+1}^0 + i M_{\nr+2}M_{\nr+1}^1) \ket{\psi} 
		\appd{O(\qr \qe/\sd)}\omega_d(M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1) \ket{\psi}.
	\end{align}

	In the last step, we prove $\ket{\psi_1}$ can be 
	approximated by
	\begin{align*}
	\ket{\psi_1} \appd{O(\sr \se)} 1/2(\1 - iM_{\nr+2})(M_{\nr+1}^0 + i M_{\nr+2}M_{\nr+1}^1) \ket{\psi}.
	\end{align*}
	The central step of proving the approximation above is
	to prove $(M_{\nr+2})^2M_{\nr+1}^0 \ket{\psi} 
	\appd{O(\sr \se)} M_{\nr+1}^0 \ket{\psi}$,
	which is proved in the previous paragraph.
	So we multiply \cref{eq:omegad} on both sides by $1/2(\1 - iM_{\nr+2})$ and get
	\begin{align}
		N_1N_2 \ket{\psi_1} \appd{O(\qr \qe/\sd)} \omega_d\ket{\psi_1}.
	\end{align}

% 	With similar calculations, we can get that 
% 	\begin{align}
% 	    N_2N_1 \ket{\psi_1} \appd{O(\sr \qe/\sd)} \omega_d^{-1}\ket{\psi_1}.
% 	\end{align}
\end{proof}
\begin{proposition}
    \label{prop:psi1_eig2}
	Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$ and $\ket{\psi_1}$ be as defined in
	\cref{eq:psi1}.
	Then,
	\begin{align}
    \label{eq:psi1_eig2} 
    &M_1M_2 \ket{\psi_1} \appd{O(\sd\qr \qe)} \omega_d^{-1} \ket{\psi_1}.
	 %&& M_2M_1 \ket{\psi_1} \appd{O(\sd\qr \qe)} \omega_d \ket{\psi_1}.
	\end{align}
\end{proposition}
This proof is similar to the proof of \cref{prop:psi1_eig1} but with one extra step. 
We first prove that there exists a state $\ket{\psi_2}$ that is
close to $\ket{\psi_1}$, 
then we show $\ket{\psi_2}$ is close to an eigenstate of $M_1M_2$. 
In the last step, we show that
$\ket{\psi_1}$ is also close to an eigenstate of $M_1M_2$.
\begin{proof}
%	Using the fact that for $0 < x < 1$, $\sqrt{1+x} = 1 + 1/2 x - O(x^2)$ and $\sqrt{1-x} = 1- 1/2x - O(x^2)$,
%	we have
%	\begin{align}
%		1-\frac{(d-1)r^2 \ep}{2} \leq \sqrt{d-1} \norm{\ket{\psi_1}} \leq 1 + \frac{(d-1)r^2 \ep}{2},
%	\end{align}
%	and we conclude that 
%	\begin{align}
%		\norm{\ket{\psi_1}} \appd{O(\sd r^2 \ep)} \frac{1}{\sqrt{d-1}}.
%	\end{align}	
    In this proof, we still use the notation $c := \cos(\pi/d)$ and $s := \sin(\pi/d)$. 
	We first introduce an approximation of $\ket{\psi_1}$,  
	\begin{align}
	    \label{eq:psi2}
	    &\ket{\psi_2} =1/2 (N_{\nr+1}^0 - iN_{\nr+2}N_{\nr+1}^1+iN_{\nr+2}N_{\nr+1}^0 + N_{\nr+1}^1)\ket{\psi},
	\end{align}
	and we will show that 
	\begin{align}
	    \label{eq:psi1_eq} &\ket{\psi_1} \appd{O(\sd \qr\qe)} \ket{\psi_2}.
	\end{align}

	To prove \cref{eq:psi1_eq}, we compare each term of $\ket{\psi_1}$
	with $\ket{\psi_2}$.
	Recall that $M_{\nr+1}^0 \ket{\psi} \appd{O(\sr \se)} N_{\nr+1}^0 \ket{\psi}$ 
	and $M_{\nr+1}^1 \ket{\psi} \appd{O(\sr \se)} N_{\nr+1}^1 \ket{\psi}$ from \cref{prop:chsh_base},
	then 
	\begin{align*}
	    \ket{\psi_2} - &\ket{\psi_1} \\
	    &\appd{O(\sr \se)}
	    (- N_{\nr+2}N_{\nr+1}^1+N_{\nr+2}N_{\nr+1}^0)\ket{\psi}
	    - (M_{\nr+2}M_{\nr+1}^1 - M_{\nr+2}M_{\nr+1}^0) \ket{\psi}.
	\end{align*}
	Therefore, proving \cref{eq:psi1_eq} 
	is reduced to proving
	$(M_{\nr+2}M_{\nr+1}^1 - M_{\nr+2}M_{\nr+1}^0) \ket{\psi}$ is close to $(- N_{\nr+2}N_{\nr+1}^1+N_{\nr+2}N_{\nr+1}^0)\ket{\psi}$,
%	or equivalently,
%	$M_{\nr+2}M_{\nr+1} \ket{\psi}$ is close to $-N_{\nr+2}N_{\nr+1} \ket{\psi}$.
	At this point, we can 
	substitute appropriate operators and the shared 
	state of $\bS_1$ into 
	the anti-commutation relation (\cref{eq:zaxa})
	and see that 
	\begin{align*}
		M_{\nr+1}M_{\nr+2} M_0^0 \ket{\psi} \appd{\frac{\norm{M_0^0\ket{\psi}}}{c\sqrt{s}}O(\sd\qr\qe)} -M_{\nr+1}M_{\nr+2} M_0^0 \ket{\psi}.
	\end{align*}
	Using the fact $c = 1 - \pi^2/(2d^2) + O(1/d^4)$ and $s = \pi/d + O(1/d^3)$, the error term above is of order $O(\sd \qr \qe)$.
	Then we can show that 
	\begin{align*}
		M_{\nr+2}M_{\nr+1}\ket{\psi} &= M_{\nr+2}M_{\nr+1}(M_{\nr+1}^0 + M_{\nr+1}^1+M_{\nr+1}^2)\ket{\psi} \\
			&\appd{O(\sr\se)} M_{\nr+2}M_{\nr+1} M_0^0 \ket{\psi} \\
			&\appd{O(\sd \qr \qe)} -M_{\nr+1}M_{\nr+2} M_0^0 \ket{\psi}\\
			&\appd{O(\sr \se)} -M_{\nr+1}M_{\nr+2} \ket{\psi} \\
			&\appd{O(\sr \se)} -M_{\nr+1} N_{\nr+2} \ket{\psi} \\
			&\appd{O(\sr \se)} -N_{\nr+2} N_{\nr+1} \ket{\psi},
	\end{align*}
	where we use the fact that $M_0^0 \ket{\psi} \appd{O(\sr \se)} (M_{\nr+1}^0+M_{\nr+1}^1)\ket{\psi} \appd{O(\sr \se)}
	(M_{\nr+2}^0+M_{\nr+2}^1) \ket{\psi}$.

	Next, we show that $M_1M_2 \ket{\psi_2}\appd{O(\qr \qe/\sd)} \omega_d^{-1} \ket{\psi_2}$, which gives
	us \cref{eq:psi1_eig2} after substituting in \cref{eq:psi1_eq}.
	With similar derivation as in the proof of \cref{prop:psi1_eig1}, we can apply results from \cref{sec:chsh}
	to the strategy $\bS_2$ to get that
	\begin{align}
		&(M_1+M_2) N_{\nr+1}^0 \ket{\psi} \appd{O(\qr \qe/\sd)} 2c N_{\nr+1}^0 \ket{\psi}.\\
		&(M_1+M_2) N_{\nr+1}^1 \ket{\psi}  \appd{O(\qr \qe/\sd)} -2cN_{\nr+1}^1 \ket{\psi},\\
		&(M_1-M_2) N_{\nr+1}^0\ket{\psi} \appd{O(\qr\qe/\sd)}-2sN_{\nr+2} N_{\nr+1}^1\ket{\psi}\\
		&(M_1-M_2)N_{\nr+1}^1 \ket{\psi} \appd{O(\qr\qe/\sd)} -2sN_{\nr+2} N_{\nr+1}^0 \ket{\psi}.
	\end{align}
	So we can further deduce that 
	\begin{align}
	\label{eq:psi2_base}
		M_1M_2 (N_{\nr+1}^0 - i N_{\nr+2}N_{\nr+1}^1) \ket{\psi} \appd{O(\qr \qe/\sd)} 
		\omega_d^{-1} (N_{\nr+1}^0 - i N_{\nr+2}N_{\nr+1}^1)\ket{\psi} .
	\end{align}
	It can be shown that $\ket{\psi_2} \appd{O(\sr \se)} 1/2(\1+iN_{\nr+2})(N_{\nr+1}^0 - i N_{\nr+2}N_{\nr+1}^1) \ket{\psi}$. 
	Proving this approximation of $\ket{\psi_1}$
	uses the same techniques as in the proof of \cref{prop:psi1_eig1}, so we omit it here.

	We multiply both sides of \cref{eq:psi2_base} by $1/2(\1+iN_{\nr+2})$ and get that
	\begin{align}
		M_1M_2 \ket{\psi_2} \appd{O(\qr \qe/\sd)} \omega_d^{-1}\ket{\psi_2}.
	\end{align}
	In the end, the relation $\ket{\psi_1}  \appd{O(\sd \qr\qe)} \ket{\psi_2}$ implies that 
	\begin{align}
		M_1M_2 \ket{\psi_1} \appd{O(\sd \qr \qe)} \omega_d^{-1}\ket{\psi_1}.
	\end{align}
% 	With similar derivation we can get that 
% 	\begin{align}
% 	    M_2M_1 \ket{\psi_1} \appd{O(\sd \qr \qe)} \omega_d\ket{\psi_1}.
% 	\end{align}
\end{proof}
%Note that it can be shown that 
%\begin{align}
%	&\ket{\psi_1} =  (\1 - iA_{n+2})(\1 + i \frac{B_1 - B_2}{-2\sin(\pi/d)})A_{n+1}^0 \ket{\psi}\\
%	&\ket{\psi_2} =  (\1 - iA_{n+2})(\1 + i \frac{B_1 - B_2}{-2\sin(\pi/d)})A_{n+1}^1 \ket{\psi}.
%\end{align}
Next lemma is the major result of this subsection, which gives us a decomposition of $\ket{\psi}$.
Before proving the next lemma, we make one important observation.
Recall that \cref{prop:lct_base} establishes that there exist observables $\{M(s), N(s)\}$ such that 
$M(s)N(s) \ket{\psi} \appd{O(\sr\se)} \ket{\psi}$ for $s \in S$.
Then we can deduce that 
\begin{align*}
	M(a_4)M(a_3) \x N(a_4)N(a_3) \ket{\psi} \appd{O(\sr\se)} M(a_4) \x N(a_4) \ket{\psi} \appd{O(\sr\se)} \ket{\psi},
\end{align*}
or equivalently,
\begin{align*}
	U_A\ct \x U_B\ct \ket{\psi} \appd{O(\sr\se)} \ket{\psi}.
\end{align*}
Intuitively, it means that $\ket{\psi}$ is invariant under the permutation induced by $U_A\ct \x U_B\ct$.
We use this observation in the proof of the following lemma.
\begin{lemma}
\label{lm:decomp_psi}
Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$ and 
define 
$\ket{\psi'} := \sum_{j=1}^{(d-1)} (U_A\ct U_B\ct)^{\log_r j} \ket{\psi_1}$
with $\log_r j$ being the discrete log and
$\ket{\psi_1}$ as defined in \cref{eq:psi1}.
Then,
\begin{align}
	&\ket{\psi} \appd{O(d r^{d/2} \ep^{1/8})} \ket{\psi'}.
\end{align}
\end{lemma}
\begin{proof}
% 	We first prove that $\norm{\ket{\psi'}}^2 \appd{O(d r^d \qe)} 1$.
	The proof takes two steps. In the first step, 
	we show that the norm of $\ket{\psi'}$ is close to $1$.
	In the second step, 
	we show that $\braket{\psi}{\psi'}$ is close to $1$.

	Expanding $\braket{\psi'}{\psi'}$, we get
	\begin{align*}
	    \braket{\psi'}{\psi'} 
	    = \sum_{j=1}^{d-1} \bra{\psi_1} (U_AU_B)^{\log_r j} (U_A\ct U_B\ct)^{\log_r j}  \ket{\psi_1}
	    + \sum_{j\neq j'} \bra{\psi_1} (U_AU_B)^{\log_r j} (U_A\ct U_B\ct)^{\log_r j'}  \ket{\psi_1}.
	\end{align*}
	Observe that 
	\begin{align*}
		\bra{\psi_1} (U_AU_B)^{\log_r j} (U_A\ct U_B\ct)^{\log_r j}  \ket{\psi_1} =
		\braket{\psi_1}{\psi_1} \appd{O(\sr\se)} \frac{1}{d-1},
	\end{align*}
	so we need to show when $j \neq j'$, the inner product 
	$\bra{\psi_1} (U_AU_B)^{\log_r j} (U_A\ct U_B\ct)^{\log_r j'}  \ket{\psi_1}$
	is close to $0$.
	Using the relation $O_B(U_B\ct)^j \ket{\psi} \appd{O(r^{j+1/2} \se)} (U_B\ct)^j O_B^{r^j}\ket{\psi}$ and \cref{eq:psi1_eig1}, we can deduce that 
	\begin{align*}
	    O_B (U_B\ct)^j \ket{\psi_1} &= \frac{1}{2} (\1 - iM_{\nr+2})(M_{\nr+1}^0 + i M_{\nr+2}M_{\nr+1}^1)  O_B (U_B\ct)^j \ket{\psi}\\
	   & \appd{O(r^{j+1/2} \se)} \frac{1}{2} (\1 - iM_{\nr+2})(M_{\nr+1}^0 + i M_{\nr+2}M_{\nr+1}^1) (U_B\ct)^j 
	   O_B^{r^j} \ket{\psi} \\
	   & = (U_B\ct)^j O_B^{r^j} \ket{\psi_1} \\
	   & \appd{O(r^{j+1/4} \qe/ \sd)} \omega_d^{r^j} (U_B\ct)^j\ket{\psi_1}
	\end{align*}
	In short, we have shown that 
    \begin{align}
        \label{eq:B_omega}
        O_B (U_B\ct)^j \ket{\psi_1} \appd{O(r^{j} \qe/ \sd )} \omega_d^{r^j}  (U_B\ct)^j \ket{\psi_1},
    \end{align}
    where we drop the insignificant error term $r^{1/4}$.
	%Using the relation $O_A(U_A\ct)^j \ket{\psi} \appd{O(r^j \se)} (U_A\ct)^j O_A^{r^j}\ket{\psi}$ and $\ket{\psi_1} \appd{O(\sd \sr \qe)} \ket{\psi_2}$ defined in the proof of \cref{prop:psi1},
    %we can see that 
    %\begin{align*}
            % 	O_A (U_A\ct)^j \ket{\psi_1} &\appd{O(\sd \sr \qe)} O_A(U_A\ct)^j \ket{\psi_2} \\
            % 	& = 1/2 (\1 + i N_{\nr+2})(N_{\nr+1}^0 - iN_{\nr+2}N_{\nr+1}^1) O_A (U_A\ct)^j \ket{\psi} \\
            % 	& \appd{O(r^j \se)} 1/2  (\1 + i N_{\nr+2})(N_{\nr+1}^0 - iN_{\nr+2}N_{\nr+1}^1)  (U_A\ct)^j O_A^{r^j} \ket{\psi} \\
            % 	& = (U_A\ct)^j O_A^{r^j} \ket{\psi_2} \\
            % 	& \appd{O(r^j \sr \qe/\sd)} \omega_d^{-r^j}(U_A\ct)^j \ket{\psi_2} \\
            % 	& \appd{O(\sd \sr\qe)} \omega_d^{-r^j}  (U_A\ct)^j \ket{\psi_1}.
    %\end{align*}
    %In short, we have shown that 
    With similar derivation, we can also get
    \begin{align}
    \label{eq:A_omega}
        O_A (U_A\ct)^j \ket{\psi_1} \appd{O(\sd r^{j} \qe)}\omega_d^{-r^j}  (U_A\ct)^j \ket{\psi_1}.
    \end{align}
	Hence for $j \neq j'$, $(U_A\ct U_B\ct)^{j} \ket{\psi_1}$ and $(U_A\ct U_B\ct)^{ j'} \ket{\psi_1}$ are close to eigenvectors
	of the unitary $O_B$ with different eigenvalues.
	Applying \cref{prop:orthog} to \cref{eq:B_omega} gives us that  
	\begin{align*}
		\bra{\psi_1} (U_AU_B)^{j} (U_A\ct U_B\ct)^{ j'} \ket{\psi_1} \appd{O( (r^j + r^{j'}) \qe)} 0,
	\end{align*}
	where we use the fact that $\norm{1 - \omega_d^{r^{j'}-r^j}} = O(1/d)$.
	Therefore, we can calculate the norm of $\ket{\psi'}$ 
	\begin{align*}
		\braket{\psi'}{\psi'} &=\sum_{j,j'=1}^{d-1} \bra{\psi_1} (U_AU_B)^{\log_r j} (U_A\ct U_B\ct)^{\log_r j'} \ket{\psi_1}\\
		&\appd{O(dr^{d} \qe)} \sum_{j=1}^{d-1} \bra{\psi_1} (U_AU_B)^{\log_r j} (U_A\ct U_B\ct)^{\log_r j} \ket{\psi_1}\\
		&\appd{O(dr\se)} (d-1) \frac{1}{d-1} = 1.
	\end{align*}

	Next we calculate $\braket{\psi}{\psi'}$ which is
	\begin{align*}
		\braket{\psi}{\psi'} = \sum_{j=1}^{d-1} \bra{\psi} (U_A\ct U_B\ct)^{\log_r j} \ket{\psi_1}
		 \appd{O(d^2\sr\se)} (d-1) \braket{\psi}{\psi_1}.
	\end{align*}
	The problem is reduced to calculate $\braket{\psi}{\psi_1}$, which is
	\begin{align*}
		\braket{\psi}{\psi_1} &= \frac{1}{2} \bra{\psi} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 +M_{\nr+1}^1) \ket{\psi}\\
		& = \frac{1}{2} \left(\bra{\psi}(M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi} - i \bra{\psi}M_{\nr+2}M_{\nr+1} \ket{\psi} \right)\\
		& \appd{O(\sr\se)} \frac{1}{2} \left(\bra{\psi}(M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi}  - i \bra{\psi}N_{\nr+2}M_{\nr+1} \ket{\psi} \right)\\
		& \appd{O(r \ep)} \frac{1}{2} \frac{2}{d-1} = \frac{1}{d-1},
	\end{align*}
	where the approximation of $\bra{\psi}N_{\nr+2}M_{\nr+1} \ket{\psi}$ comes from the correlation.
	Hence, we know $\braket{\psi}{\psi'} \appd{O(d^2 \sr \se)} 1$.

	In the end, we combine $\braket{\psi}{\psi'} \appd{O(d^2 \sr \se)} 1$ and $\norm{\ket{\psi'}}^2 \appd{O(d r^d \qe)} 1$,
	and apply \cref{prop:close_vec} to conclude
	\begin{align*}
		\ket{\psi} \appd{O(d r^{d/2} \ep^{1/8})} \ket{\psi'}.
	\end{align*}
\end{proof}

%------------------------------------------------------------------------------------------
\subsection{Implications of passing the commutation games}
\label{sec:imp_comm}
%-------------------------------------------------------------------------------------------
The commutation game is introduced to certify some observables used in $\LS(r)$ commute with 
the operators for the extended weighted CHSH test. For the details of the general commutation test, 
we refer to Appendix~A.$2$ of \cite{coladan2017verifier}. 
We use similar techniques in our proof but our proof also bases on the correlation given in \cref{tbl:comm}.

The implication of winning the commutation game is summarized in the following proposition.
\begin{proposition}
	\label{prop:rel_comm}
	Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	Then, 
	\begin{align}
		&M_{\nr+1}^0 M(s) \ket{\psi} \appd{O(\sr \se)} M(s) M_{\nr+1}^0 \ket{\psi},\\ 
		&M_{\nr+1}^1 M(s) \ket{\psi} \appd{O(\sr \se)} M(s) M_{\nr+1}^1 \ket{\psi}, \\
		\label{eq:com_nr2} &M_{\nr+2} M(s) \ket{\psi} \appd{O(\sr \se)} M(s) M_{\nr+2} \ket{\psi},
	\end{align}
	for $s = f_0, f_2,g_0,g_2$.
\end{proposition}
Note that the $M(s)$'s for $s = f_0, f_2,g_0,g_2$ tested in the commutation game are measured for the Magic Square equations in $\LS(r)$.
They will be used to construct the isometry in the next subsection.
\begin{proof}
	Proving \cref{eq:com_nr2} is an application of Lemma~$28$ of \cite{coladan2017verifier} with $\se$ replaced by $O(\sr \se)$.

	For the other relations, we prove $M_{\nr+1}^0M(f_0) \ket{\psi} \appd{O(\sr \se)} M(f_0)M_{\nr+1}^0 \ket{\psi}$ to demonstrate 
	the techniques involved.

	We denote Bob's projectors by $\{ N_{(\nr+1, f_0)}^{b_1 b_2} \}$ where $b_1 \in \Z_3$ and
	$b_2 \in \Z_2$. From the correlation in \cref{tbl:comm}, 
	we can get that 
	\begin{align*}
		&M_{\nr+1}^0 \ket{\psi} \appd{O(\sr \se)} 
		(N_{(\nr+1, f_0)}^{00} + N_{(\nr+1, f_0)}^{01}) \ket{\psi} \\
		&M(f_0) \ket{\psi} \appd{O(\sr \se)} 
		(\sum_{b_1, b_2} (-1)^{b_2} N_{(\nr+1, f_0)}^{b_1b_2})\ket{\psi}.
	\end{align*}
	Then we can show that
	\begin{align*}
		M_{\nr+1}^0 M(f_0) \ket{\psi} &\appd{O(\sr \se)} M_{\nr+1}^0 (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1,f_0)}^{b_1b_2})\ket{\psi}\\
		&\appd{O(\sr\se)}(\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, f_0)}^{b_1b_2}) 
		(N_{(\nr+1, f_0)}^{00} + N_{(\nr+1, f_0)}^{01}) \ket{\psi} \\
		&= (N_{(\nr+1, f_0)}^{00} + N_{(\nr+1, f_0)}^{01}) (\sum_{b_1, b_2} (-1)^{b_2}N_{(\nr+1, f_0)}^{b_1b_2})  \ket{\psi} \\
		& \appd{O(\sr \se)} (N_{(\nr+1, f_0)}^{00} + N_{(\nr+1, f_0)}^{01}) M_{r+6} \ket{\psi} \\
		&= M(f_0)  (N_{(\nr+1, f_0)}^{00} + N_{(\nr+1, f_0)}^{01}) \ket{\psi} \\
		& \appd{O(\sr\se)} M(f_0)M_{\nr+1}^0 \ket{\psi}.
	\end{align*}
\end{proof}
%-----------------------------------------------
\subsection{Robust self-testing}
\label{sec:self-test}
%-----------------------------------------------
Based on the propositions and lemmas proved so far, we can prove the self-testing theorem.
\begin{theorem}
	\label{thm:self-test}
	Let $\bS$ be an $\ep$-approximate strategy of $\fC(d,r)$.
	Then there exists an isometry   
	$\Phi_A \x \Phi_B = (\Phi_{A,2} \x \Phi_{B,2})(\Phi_{A,1}\x\Phi_{B,1})$ and a state $\ket{junk}$
	such that $\norm{\ket{junk}} \appd{O(\sd \qr \qe)} 1$ and 
	\begin{align}
		&\Phi_A \x \Phi_B (\ket{\psi}) \appd{O(d r^{d} \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x \ket{\EPR{d-1}}\\	
		&\Phi_A \x \Phi_B (O_A\ket{\psi}) \appd{O(d r^{d}\ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \omega_d^{d-j}\ket{x_{d-j}}_{A'}\ket{x_j}_{B'} \\
		&\Phi_A \x \Phi_B (O_B\ket{\psi}) \appd{O(d r^{d} \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \omega_d^{j}\ket{x_{d-j}}_{A'}\ket{x_j}_{B'}\\
		&\Phi_A \x \Phi_B (U_A\ket{\psi}) \appd{O(d r^{d}  \ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{(d-j)r^{-1}}}_{A'}\ket{x_j}_{B'} \\
		&\Phi_A \x \Phi_B (U_B\ket{\psi}) \appd{O(d r^{d}\ep^{1/8}))} \ket{junk} \x \ket{\EPR{2}}^{\x 2} \x
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_{j r^{-1}}}_{B'}.
	\end{align}
\end{theorem}
Before proving \cref{thm:self-test}, we state the consequence of it.
It implies that,
for any odd prime number $d$ whose primitive root is $r$, 
there exists a correlation of size $\Theta(r)$ that can robustly self-test the maximally
entangled state of local dimension $4(d-1)$.
Since there are infinitely many prime numbers
whose smallest primitive roots are in the set $\{2,3,5\}$ \cite{murty1988},
we can apply \cref{thm:self-test} to the set $D$
of all such odd prime
numbers and prove the following theorem.
\begin{theorem}
	\label{thm:infty}
	There exists an infinity-sized set $D$ of prime numbers such that 
	for each $d \in D$, there exists
	a constant-sized correlation that can self-test the maximally entangled state
	of local dimension $4(d-1)$.
\end{theorem}

Before proving \cref{thm:self-test}, we introduce the isometry 
$\Phi_A \x \Phi_B$, which has two steps. 
We denote the isometry used in the first step and the second step by $\Phi_{A,1} \x \Phi_{B,1}$ and 
$\Phi_{A,2} \x \Phi_{B,2}$ respectively, which are illustrated 
in the two figures below.

To construct the isometry $\Phi_{A,1} \x \Phi_{B,1}$, 
we use $U_A, O_A$ and $U_B$, $O_B$ defined in \cref{eq:uaoa,eq:ubob} 
to produce a copy of $\ket{\EPR{d-1}}$.

To construct the isometry $\Phi_{A,2} \x \Phi_{B,2}$. we use the observables $M(f_0),M(f_2), M(g_0), M(g_2)$ and 
$N(f_0), N(f_2), N(g_0), N(g_2)$ to produce two copies of $\ket{EPR}$.
Intuitively, $M(f_0)$ acts as $\1 \x \paulix$; $M(f_2)$ acts as $\paulix \x \1 $;
$M(g_0)$ acts as $\1 \x \pauliz$ and $M(g_2)$ acts as $\pauliz \x \1$; and similarly on Bob's side.

\begin{figure}[H]
\center
        \begin{tikzpicture}[thick]
        %
        % `operator' will only be used by Hadamard (H) gates here.
        % `phase' is used for controlled phase gates (dots).
        % `surround' is used for the background box.
        \tikzstyle{operator} = [draw,fill=white,minimum size=1.5em] 
        \tikzstyle{phase} = [fill,shape=circle,minimum size=5pt,inner sep=0pt]
        \tikzstyle{surround} = [fill=blue!10,thick,draw=black,rounded corners=2mm]
        %
        % Bracket
        \draw[decorate,decoration={brace,mirror},thick] (0,-1) to
    	node[midway,left] (bracket1) {$\ket{\psi}$}
    	(0,-2);
        % Qubits
        \node at (0,0) (q1) {$\ket{x_0}$};
        \node at (0,-1) (q2) {};
        \node at (0,-2) (q3) {};
        \node at (0,-3) (q4) {$\ket{x_0}$};
        %
        % Column 1
        \node[operator] (op11) at (1,0) {$QFT_d$} edge [-] (q1);
        \node[operator] (op14) at (1,-3) {$QFT_d$} edge [-] (q4);
        %
        % Column 3
        \node[phase] (phase11) at (2,0) {} edge [-] (op11);
	\node[operator] (op22) at (2,-1) {$O_A$} edge [-] (q2);
	\node[operator] (op23) at (2, -2) {$O_B$} edge[-] (q3);
        \node[phase] (phase14) at (2,-3) {} edge [-] (op14);
        \draw[-] (phase11) -- (op22);
        \draw[-] (phase14) -- (op23);
        %
        % Column 4
        \node[operator] (op31) at (3,0) {$QFT_d^{-1}$} edge [-] (phase11);
        \node[operator] (op34) at (3,-3) {$QFT_d^{-1}$} edge [-] (phase14);
        %
        % Column 5
        \node[phase] (phase21) at (4,0) {} edge [-] (op31);
	\node[operator] (op42) at (4,-1) {$U_A$} edge [-] (op22);
	\node[operator] (op43) at (4, -2) {$U_B$} edge[-] (op23);
        \node[phase] (phase24) at (4,-3) {} edge [-] (op34);
        \draw[-] (phase21) -- (op42);
        \draw[-] (phase24) -- (op43);
        %
        % Column 6
        \node (end2) at (5,-1) {} edge [-] (op42);
        \node (end3) at (5,-2) {} edge [-] (op43);
        %
        % Bracket
        \draw[decorate,decoration={brace},thick] (5,-1) to
    	node[midway,right] (bracket) {$\ket{reuse}$}
    	(5,-2);
        %
        % Column 7
        \node (end1) at (6.5,-0.5) {} edge [-] (phase21);
        \node (end4) at (6.5,-2.5) {} edge [-] (phase24);
        % Dashed line
        \draw[loosely dotted] (0,-1.5) -- (5,-1.5);
        % Bracket
        \draw[decorate,decoration={brace},thick,] (6.5,-0.5) to
    	node[midway,right] (bracket2) {$\sum_{j=1}^{d-1}\ket{x_j}\ket{x_{d-j}}$}
    	(6.5,-2.5);
        %
        % Background Box
        \begin{pgfonlayer}{background} 
        \node[surround] (background) [fit = (q1) (op14) (bracket1)(bracket2)] {};
        \end{pgfonlayer}
        %
        \end{tikzpicture}
	\caption{The isometries $\Phi_{A,1} \x \Phi_{B,1}$.}
\end{figure}

The isometry $\Phi_{A,1} \x \Phi_{B,1}$ has the following steps:
\begin{enumerate}
	\item Append control register $\ket{x_0}_{A'}$ on Alice's side and $\ket{x_0}_{B'}$ on Bob's side;
	\item Apply Quantum Fourier Transform ($QFT_d$) to Alice and Bob's control registers;
	\item Apply Controlled-$O_{A/B}$ operations (i.e. if the control register is in state $\ket{x_k}_{A'/ B'}$, apply
	$O_{A/B}^k$.);
	\item Apply inverse Quantum Fourier Transform ($QFT_d^{-1}$) to the control registers;
	\item Apply Controlled-$U_{A/B}$ operations (i.e. If Alice's control register is in state $\ket{x_j}$, she applies
	$U_A^{\log_r (d-j)}$. If Bob's control register is in state $\ket{x_j}$, he applies $(U_B)^{\log_r j}$).
\end{enumerate}
Note that $\ket{x_0}$ is orthogonal to $W_{d-1}$, so
$\spn\{ x_j \}_{j=0}^{d-1}$ has dimension $d$.
\begin{figure}[H]
\center
        \begin{tikzpicture}[thick]
        %
        % `operator' will only be used by Hadamard (H) gates here.
        % `phase' is used for controlled phase gates (dots).
        % `surround' is used for the background box.
        \tikzstyle{operator} = [draw,fill=white,minimum size=1.5em] 
        \tikzstyle{phase} = [fill,shape=circle,minimum size=5pt,inner sep=0pt]
        \tikzstyle{surround} = [fill=blue!10,thick,draw=black,rounded corners=2mm]
        %
        % Bracket
        \draw[decorate,decoration={brace,mirror},thick] (0,-2) to
    	node[midway,left] (bracket1) {$\ket{reuse}$}
    	(0,-3);
        % Qubits
        \node at (0,0) (q1) {$\ket{0}$};
        \node at (0,-1) (q2) {$\ket{0}$};
        \node at (0,-2) (q3) {};
        \node at (0,-3) (q4) {};
        \node at (0,-4) (q5) {$\ket{0}$};
        \node at (0,-5) (q6) {$\ket{0}$};
        %
        % Column 1
        \node[operator] (op11) at (1,0) {$H$} edge [-] (q1);
        \node[operator] (op12) at (1,-1) {$H$} edge [-] (q2);
         \node[operator] (op15) at (1,-4) {$H$} edge [-] (q5);
        \node[operator] (op16) at (1,-5) {$H$} edge [-] (q6);
        %
        % Column 3
        \node[phase] (phase11) at (2,0) {} edge [-] (op11);
	\node[operator] (op23) at (2,-2) {$M(g_2)$} edge [-] (q3);
	\node[operator] (op24) at (2, -3) {$N(g_2)$} edge[-] (q4);
        \node[phase] (phase16) at (2,-5) {} edge [-] (op16);
        \draw[-] (phase11) -- (op23);
        \draw[-] (phase16) -- (op24);
        %
        % Column 4
        \node[phase] (phase12) at (3.5,-1) {} edge [-] (op12);
	\node[operator] (op33) at (3.5,-2) {$M(g_0)$} edge [-] (op23);
	\node[operator] (op34) at (3.5, -3) {$N(g_0)$} edge[-] (op24);
        \node[phase] (phase15) at (3.5,-4) {} edge [-] (op15);
        \draw[-] (phase12) -- (op33);
        \draw[-] (phase15) -- (op34);
        %
         % Column 5
        \node[operator] (op41) at (4.5,0) {$H$} edge [-] (phase11);
        \node[operator] (op42) at (4.5,-1) {$H$} edge [-] (phase12);
         \node[operator] (op45) at (4.5,-4) {$H$} edge [-] (phase15);
        \node[operator] (op46) at (4.5,-5) {$H$} edge [-] (phase16);
        %
        % Column 6
        \node[phase] (phase51) at (5.5,0) {} edge [-] (op41);
	\node[operator] (op53) at (5.5,-2) {$M(f_2)$} edge [-] (op33);
	\node[operator] (op54) at (5.5, -3) {$N(f_2)$} edge[-] (op34);
        \node[phase] (phase56) at (5.5,-5) {} edge [-] (op46);
        \draw[-] (phase51) -- (op53);
        \draw[-] (phase56) -- (op54);
        %
        % Column 7
        \node[phase] (phase62) at (7,-1) {} edge [-] (op42);
	\node[operator] (op63) at (7,-2) {$M(f_0)$} edge [-] (op53);
	\node[operator] (op64) at (7, -3) {$N(f_0)$} edge[-] (op54);
        \node[phase] (phase65) at (7,-4) {} edge [-] (op45);
        \draw[-] (phase62) -- (op63);
        \draw[-] (phase65) -- (op64);
        %
        % Bracket
        \node(end3) at (8, -2) {} edge [-] (op63);
        \node(end4) at (8, -3) {} edge [-] (op64);
        \draw[decorate,decoration={brace},thick] (8.2,-2) to
    	node[midway,right] (bracket) {$\ket{junk}$}
    	(8.2,-3);
        %
        % Column 7
        \node (end1) at (9.5,-1) {} edge [-] (phase51);
        \node(end2) at (9.5, -1.5) {} edge [-] (phase62);
       \node(end5) at (9.5, -3.5) {} edge [-] (phase65);
        \node (end6) at (9.5,-4) {} edge [-] (phase56);
        % Dashed line
        \draw[loosely dotted] (0,-2.5) -- (7.8,-2.5);
        \draw[solid] (end1) -- (9.3, -1);
        % Bracket
        \draw[decorate,decoration={brace},thick,] (9.5,-1.1) to
    	node[midway,right] (bracket2) {$\ket{EPR}^{\x 2}$}
    	(9.5,-3.9);
        %
        % Background Box
        \begin{pgfonlayer}{background} 
        \node[surround] (background) [fit = (q1) (op16) (bracket1)(bracket2)] {};
        \end{pgfonlayer}
        %
        \end{tikzpicture}
	\caption{The isometries $\Phi_{A,2} \x \Phi_{B,2}$.}
\end{figure}
The second isometry is the standard isometry used in the self-testing result of the Magic Square game \cite{wu2016}.
The only thing to note is that the state $\ket{reuse}$, which is on the same Hilbert space as $\ket{\psi}$ and produced by $\Phi_{A,1} \x \Phi_{B,1}$, is the input state to $\Phi_{A,2} \x \Phi_{B,2}$. In this sense, our isometry is a $2$-step sequential procedure.

%We first list the key relations needed for $\Phi_{A,1} \x \Phi_{B,1}$ to work.
%\begin{align}
%	&\ket{\psi} \appd{O(d^{5/4} r^{d/2} r^{1/4} \ep^{1/8})}\ket{\psi'} = \sum_{j=1}^{d-1} (U_A\ct U_B\ct)^{\log_r j} \ket{\psi_1},\\
%	&O_A(U_A\ct)^{\log_r j} \ket{\psi_1} \appd{O(\sd r^j \sr \qe)}\omega_d^{-r^j}  (U_A\ct)^j \ket{\psi_1}\\
%	&O_B(U_A\ct)^{\log_r j} \ket{\psi_1} \appd{O(\sd r^j \sr \qe)}\omega_d^{r^j}  (U_A\ct)^j \ket{\psi_1}.
%\end{align}
%The second isometry just requires properties of the Magic Square game, especially, the commutation relations in \cref{prop:lct_comm}.

\begin{proof}[Proof of \cref{thm:self-test}]
The proof takes two steps. 
We first show that 
\begin{align}
	\Phi_{A,1}\x\Phi_{B,1}(\ket{\psi}) \appd{O(d r^{d} \ep^{1/8})} \sqrt{d-1}\ket{\psi_1} \x \frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'}.
\end{align}
Then we show that there exists a state $\ket{junk}$ such that  
\begin{align}
\label{eq:phi2_result}
\Phi_{A,2} \x \Phi_{B,2} (\sqrt{d-1}\ket{\psi_1}) \appd{O(\sr\se)} \ket{junk} \x \ket{\EPR{2}}^{\x 2}.
\end{align}
Combining the two equations above we get that 
\begin{align*}
	&\quad\Phi_{A,2} \x \Phi_{B,2}( \Phi_{A,1} \x \Phi_{B,1} (\ket{\psi})) \\
	&\appd{O(d r^d \ep^{1/8})} \left[\Phi_{A,2} \x \Phi_{B,2}
	(\sqrt{d-1} \ket{\psi_1})\right] \x \frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'} \\
	&\appd{O(\sr\se)} \ket{junk} \x \ket{EPR}^{\x 2} \x\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'},
\end{align*}
where we use the fact that $\Phi_{A,2}\x \Phi_{B,2}$ only 
acts nontrivially on the state $\ket{\psi_1}$.
The state $\sqrt{d-1} \ket{\psi_1}$ is the state
$\ket{reuse}$ in the two figures.

\Cref{lm:decomp_psi} implies that $ \Phi_{A,1} \x \Phi_{B,1} (\ket{\psi}) \appd{O(d r^{d/2 +1/4} \ep^{1/8})}  \Phi_{A,1} \x \Phi_{B,1} (\ket{\psi'})$,  
so we focus on how the isometry evolves $\ket{\psi'}$.
The evolution is summarized below.
	\begin{align*}
		& \sum_{j=1}^{d-1} (U_A\ct U_B\ct)^{\log_r j} \ket{\psi_1} \ket{x_0}_{A'}\ket{x_0}_{B'}\\
		\xrightarrow[]{QFT_d}& \frac{1}{d}\sum_{k_1,k_2 = 0}^{d-1} \sum_{j=1}^{d-1} (U_A\ct U_B\ct)^{\log_r j}  \ket{\psi_1}\ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\xrightarrow[]{\text{Controlled-}O_{A/B}}& \frac{1}{d}\sum_{k_1,k_2 = 0}^{d-1} \sum_{j=1}^{d-1} O_A^{k_1}(U_A\ct)^{\log_r j} O_B^{k_2}(U_B\ct)^{\log_r j}
		\ket{\psi_1} \ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\appd{O(\sd r^{d} \qe)}&\frac{1}{d} \sum_{k_1,k_2 = 0}^{d-1} \sum_{j=1}^{d-1} (U_A\ct U_B\ct)^{\log_r j}\omega_d^{(k_2-k_1)j}\ket{\psi_1} \ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\xrightarrow[]{QFT_d^{-1}} &\frac{1}{d^2}\sum_{l_1,l_2 = 0}^{d-1}\sum_{k_1,k_2 = 0}^{d-1}\sum_{j=1}^{d-1} (U_A\ct U_B\ct)^{\log_r j} 
		\omega_d^{k_1(d-j-l_1)}\omega_d^{k_2(j-l_2)}\ket{\psi_1} \ket{x_{l_1}}_{A'}\ket{x_{l_2}}_{B'}\\
		= &\sum_{j=1}^{d-1}(U_A\ct U_B\ct)^{\log_r j} \ket{\psi_1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'} \\
		\xrightarrow[]{\text{Controlled-}U_{A/B}}& \sum_{j=1}^{d-1} U_A^{\log_r j} (U_A\ct)^{\log_r j} U_B^{\log_r j} (U_B\ct)^{\log_r j} \ket{\psi_1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'}\\
		=&\ket{\psi_1} \x \sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'}.
	\end{align*}
When we analyze the effect of the controlled-$O_{A/B}$, we applied \cref{eq:A_omega,eq:B_omega} repeatedly.
In summary, we have shown that
\begin{align}
	\Phi_{A,1}\x\Phi_{B,1}(\ket{\psi}) \appd{O(d r^{d} \ep^{1/8})} \left(\sqrt{d-1} \ket{\psi_1}\right) \x \left(\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'}\right),
\end{align}
where we use the fact that $O(d r^{d} \ep^{1/8})$ dominates both $O(d r^{d/2+1/4} \ep^{1/8})$ and 
$O(\sd r^{d} \qe)$.
The relation $ \sqrt{d-1} \norm{\ket{\psi_1}} \appd{O(\sd \qr\qe)} 1$ is derived from the fact that
$\norm{\ket{\psi_1}}^2 \appd{O(\sr \se)} 1/(d-1)$.

To prove \cref{eq:phi2_result}, we first
recall that
\begin{align*}
    \ket{\psi_1} =\frac{1}{2} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi}.
\end{align*}
The other observation that we need is that 
\begin{align*}
	&\Phi_{A,2} \x \Phi_{B,2} (\ket{\psi_1})  \\
	= &\frac{1}{16} \sum_{\ba, \bb \in\{0,1\}^4}
	M(f_0)^{a(4)}N(f_0)^{b(4)}M(f_2)^{a(3)}N(f_2)^{b(3)}M(g_0)^{a(2)}N(g_0)^{b(2)}M(g_2)^{a(1)}N(g_2)^{b(1)}
	\ket{\psi_1} \ket{\ba}_{A''} \ket{\bb}_{B''}
\end{align*}
For simplicity we define 
\begin{align*}
    &M^{\ba} := M(f_0)^{a(4)}M(f_2)^{a(3)}M(g_0)^{a(2)}M(g_2)^{a(1)} \\
    &N^{\bb} := N(f_0)^{b(4)}N(f_2)^{b(3)}N(g_0)^{b(2)}N(g_2)^{b(1)}.
\end{align*}
Our goal is to use \cref{prop:rel_comm} to show that 
\begin{align}
\label{eq:mua_comm}
M^{\ba} \ket{\psi_1} \appd{O(\sr\se)} \frac{1}{2} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1)
M^{\ba} \ket{\psi},
\end{align}
for all $\ba \in \{0,1\}^4$,

We show how to apply \cref{prop:rel_comm} to commute $M(g_2)^{a(1)}$ through 
$\frac{1}{2} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 \\- iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1)$ and then similar process can be repeated for $M(g_0)^{a(2)}, M(f_2)^{a(3)}$ and $M(f_0)^{a(4)}$ using
\cref{lm:sub},
\begin{align*}
	&\quad M(g_2)^{a(1)} \frac{1}{2} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1) \ket{\psi} \\
	&\appd{O(\sr\se)} \frac{1}{2}[ (M_{\nr+1}^0+M_{\nr+1}^1) M(g_2)^{a(1)} -i M(g_2)^{a(1)} M_{\nr+2}N_{\nr+1}] \ket{\psi} \\
	&\appd{O(\sr\se)} \frac{1}{2} [(M_{\nr+1}^0+M_{\nr+1}^1) M(g_2)^{a(1)} - iN_{\nr+1} M_{\nr+2}M(g_2)^{a(1)} ] \ket{\psi} \\
	&\appd{O(\sr\se)}\frac{1}{2} [(M_{\nr+1}^0+M_{\nr+1}^1) M(g_2)^{a(1)} - iM_{\nr+2}M(g_2)^{a(1)} M_{\nr+1}] \ket{\psi} \\
	&\appd{O(\sr\se)}\frac{1}{2} [(M_{\nr+1}^0+M_{\nr+1}^1) M(g_2)^{a(1)} - iM_{\nr+2}M_{\nr+1}M(g_2)^{a(1)} ] \ket{\psi}\\
	&= \frac{1}{2} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1)M(g_2)^{a(1)}\ket{\psi}.
\end{align*}
In summary, we have 
\begin{align}
	M^{\ba} N^{\bb} \ket{\psi_1} \appd{O(\sr\se)} \frac{1}{2} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1)M^{\ba}N^{\bb} \ket{\psi}
\end{align}
for all $\ba, \bb \in \{0,1\}^4$, which implies that 
\begin{align*}
	\Phi_{A,2} \x& \Phi_{B,2}(\ket{\psi_1}) \appd{O(\sr\se)}\\
	 &\frac{1}{2} (M_{\nr+1}^0 + iM_{\nr+2}M_{\nr+1}^1 - iM_{\nr+2}M_{\nr+1}^0 + M_{\nr+1}^1)(\Phi_{A,2}\x\Phi_{B,2} (\ket{\psi}).
\end{align*}
At this point we can apply Lemma C.$1$ of \cite{wu2016} to $\Phi_{A,2}\x\Phi_{B,2} (\ket{\psi})$ and conclude that 
\begin{align}
	\Phi_{A,2} \x \Phi_{B,2}(\sqrt{d-1} \ket{\psi_1}) \appd{O(\sd \sr \se)} \ket{junk} \x \ket{EPR}^{\x 2},
\end{align} 
for some state $\ket{junk}$ whose norm can be deduced from the norm of $\ket{\psi_1}$.

In the rest of the proof, we only show how $\Phi_{A,1} \x \Phi_{B,1}$ acts on $O_A\ket{\psi}$ and $U_A\ket{\psi}$.

If the initial state is $O_A\ket{\psi}$, we first use the fact that 
$ \Phi_{A,1} \x \Phi_{B,1} (O_A\ket{\psi}) \appd{O(d r^{d/2+1/4} \ep^{1/8})}  \Phi_{A,1} \x \Phi_{B,1} (O_A\ket{\psi'})$, 
and then we calculate $\Phi_{A,1} \x \Phi_{B,1} (O_A\ket{\psi'})$ as
\begin{align*}
	O_A \ket{\psi'} \ket{x_0}_{A'}\ket{x_0}_{B'} =&  
		\sum_{j=1}^{d-1} O_A(U_A\ct U_B\ct)^{\log_r j}\ket{\psi_1}\ket{x_0}_{A'}\ket{x_0}_{B'}\\
		\appd{O(\sd r^d \ep^{1/4})}&\sum_{j=1}^{d-1}(U_A\ct U_B\ct)^{\log_r j} \omega_d^{-j} \ket{\psi_1} \ket{x_0}_{A'}\ket{x_0}_{B'}\\
		\xrightarrow[]{QFT_d} &\frac{1}{d}\sum_{j=1}^{d-1} \sum_{k_1,k_2 = 0}^{d-1}(U_A\ct U_B\ct)^{\log_r j} \omega_d^{-j} 
		\ket{\psi_1}\ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\xrightarrow[]{\text{Controlled-}O_{A/B}}&\frac{1}{d}\sum_{j=1}^{d-1}\sum_{k_1,k_2 = 0}^{d-1} 
		 O_A^{k_1}(U_A\ct)^{\log_r j} O_B^{k_2}(U_B\ct)^{\log_r j} \omega_d^{-j} \ket{\psi_1} 
		 \ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\appd{O(d r^{d}  \ep^{1/4})}& \frac{1}{d}\sum_{k_1,k_2 = 0}^{d-1} \sum_{j=1}^{d-1} (U_A\ct U_B\ct)^{\log_r j}
		\omega_d^{-j}\omega_d^{(k_2-k_1)j}\ket{\psi_1}
		 \ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\xrightarrow[]{QFT_d^{-1}}& \sum_{j=1}^{d-1}  (U_A\ct U_B\ct)^{\log_r j}  
		\omega_d^{d-j}\ket{\psi_1} \ket{x_{d-j}}_{A'}\ket{x_j}_{B'}\\
		\xrightarrow[]{\text{Controlled-}U_{A/B}}&   \sqrt{d-1} \ket{\psi_1} \x  
		\frac{1}{\sqrt{d-1}}\sum_{j=1}^{d-1} \omega_d^{d-j}\ket{x_{d-j}}_{A'}\ket{x_j}_{B'}.
\end{align*}
The analysis for $\Phi_{A,1} \x\Phi_{B,1} (O_B \ket{\psi})$ is very similar.

If the initial state is $U_A\ket{\psi}$, we first use the fact that 
$ \Phi_{A,1} \x \Phi_{B,1} (U_A\ket{\psi}) \appd{O(d r^{d/2+1/4} \ep^{1/8})}  \Phi_{A,1} \x \Phi_{B,1} (U_A\ket{\psi'})$, 
and then we calculate $\Phi_{A,1} \x \Phi_{B,1} (U_A\ket{\psi'})$.
\begin{align*}
	U_A \ket{\psi'} \ket{x_0}_{A'}\ket{x_0}_{B'} =&  
		\sum_{j=1}^{d-1} U_A(U_A\ct U_B\ct)^{\log_r j}\ket{\psi_1} \ket{x_0}_{A'}\ket{x_0}_{B'}\\
		=&\sum_{j=1}^{d-1}(U_A\ct)^{\log_r j-1}  (U_B\ct)^{\log_r j} \ket{\psi_1} \ket{x_0}_{A'}\ket{x_0}_{B'}\\
		\xrightarrow[]{QFT_d} &\frac{1}{d}\sum_{j=1}^{d-1} \sum_{k_1,k_2 = 0}^{d-1}(U_A\ct)^{\log_r j-1} (U_B\ct)^{\log_r j} \ket{\psi_1}  \ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\xrightarrow[]{\text{Controlled-}O_{A/B}}&\frac{1}{d}\sum_{j=1}^{d-1}\sum_{k_1,k_2 = 0}^{d-1} 
		 O_A^{k_1}(U_A\ct)^{\log_r j-1} O_B^{k_2}(U_B\ct)^{\log_r j} \ket{\psi_1} 
		 \ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\appd{O(d r^{d} \qe)}& \frac{1}{d}\sum_{k_1,k_2 = 0}^{d-1} \sum_{j=1}^{d-1} (U_A\ct)^{\log_r j-1} (U_B\ct)^{\log_r j}
		\omega_d^{k_2j-k_1jr^{-1}}\ket{\psi_1}
		 \ket{x_{k_1}}_{A'}\ket{x_{k_2}}_{B'}\\
		\xrightarrow[]{QFT_d^{-1}}& \sum_{j=1}^{d-1}  (U_A\ct)^{\log_r j-1} (U_B\ct)^{\log_r j}  
		\ket{\psi_1} \ket{x_{(d-j)r^{-1}}}_{A'}\ket{x_j}_{B'}\\
		\xrightarrow[]{\text{Controlled-}U_{A/B}}& \sqrt{d-1} \ket{\psi_1} \x  
		\frac{1}{\sqrt{d-1}} \sum_{j=1}^{d-1} \ket{x_{(d-j)r^{-1}}}_{A'}\ket{x_j}_{B'},
\end{align*}
where we use the fact that $(d-j)r^{-1} \equiv d -jr^{-1} \pmod{d}$.
The analysis for $\Phi_{A,1} \x\Phi_{B,1} (U_B \ket{\psi})$ is very similar.
\end{proof}

%%-----------------------------------------------------------------
%\subsection{Self-test}
%%-----------------------------------------------------------------
%\hl{This subsection needs to be formalized.}
%We give a formal version of \cref{thm:pr_2} first and then prove it.
%\begin{theorem}
%\label{thm:selftest}
%	If a quantum strategy using the shared state $\ket{\psi}$ achieves the ideal 2-party correlation $C(\dr{r})$ where $d$ is an odd
%	prime number with primitive root $r \in \{2,3,5\}$, then there exist local isometries $\Phi_A$ and $\Phi_B$, and a state $\ket{junk}$ such 
%	that $\Phi_A\x\Phi_B \ket{\psi} = \ket{\EPR{d-1}} \x \ket{junk}$.
%\end{theorem}
%\begin{proof}
%Suppose Alice and Bob achieve the optimal correlation with the quantum strategy $(\ket{\psi}, \{A_x\}, \{B_y\})$
%for all $x,y \in \calX \times \calY$ .
%The observables $A_x$ and $B_y$ and the shared state $\ket{\psi}$ shall not be confused with the ones used in the optimal strategy.
%By Lemma~$4.3$ of Ref.~\cite{coladan2017}, we can extract an operator solution from the perfect winning strategy 
%of the linear system game $\LS_r$. 
%For each variable $\{ x_i \}_{i=1}^{n_r}$, Alice and Bob has operators $A_i$ and $B_i$ respectively.
%The condition that they agree with assignment to variables means that 
%\begin{align}
%	\bra{\psi} A_i \otimes \overline{B_i} \ket{\psi} = 1 \Rightarrow A_i \otimes \overline{B_i} \ket{\psi} = \ket{\psi}
%	\text{ for } 1 \leq i \leq n_r
%\end{align}
%and the condition that Alice's assignments satisfy the constraint means that 
%\begin{align}
%	\Tr(\rho_A \Pi_{j: H(i,j) \neq 0} A_j) = \Tr(\rho_A) \text{ for all } 1 \leq i \leq m_r
%\end{align}
%where $\rho_A =  \Tr_B(\ketbra{\psi}{\psi})$. 
%Similarly we define $\rho_B = \Tr_A(\ketbra{\psi}{\psi})$.
%For any $\ket{v} \in \supp(\rho_A)$,
%we have 
%\begin{align}
%\Pi_{j:H(i,j) \neq 0} A_j \ket{v} = \ket{v} \text{ for all } 1 \leq i \leq m_r.
%\end{align}
%Since the relation $uxu^{-1} = x^r$, where $r$ is the primitive root of $d$, is embedded in this linear system game, we know
%\begin{align}
%	A_3A_4 A_1A_2 (A_3A_4)^\dagger \ket{v}= (A_1A_2)^r \ket{v} \text{ for all } \ket{v} \in \supp(\rho_A).
%\end{align}
%For simplicity, we define $X_A = A_1A_2$ and $U_A=A_3A_4$ such that
%the condition is equivalent to
%\begin{align}
%	\label{eq:ux_relation}
%	U_AX_AU_A^\dagger \ket{v} = X_A^r \ket{v} \text{ for all } \ket{v} \in \supp(\rho_A).
%\end{align}
%We will come back to the implication of this condition later.
%
%Suppose $A_{m_r+1} = A_\ast^0- A_\ast^\perp = \Pi_{V_A} - \Pi_{V_A^\perp}$m, where $V_A$ is a
%$2m$-dimensional vector space and $\Pi_{V_A}$ is the projector onto it. The reason why it has dimension $2m$
%will be clear shortly. On Bob's side, we also have $B_{n_r+1} = B_\ast^0- B_\ast^\perp = \Pi_{V_B} - \Pi_{V_B^\perp}$.
%Recall that from the extended weighted CHSH test, we know
%\begin{align}
%	A_\ast^0 \ket{\psi} = B_\ast^0 \ket{\psi} = A_\ast^0 \x B_\ast^0 \ket{\psi},
%\end{align}
%which implies that with an appropriate change of basis, we can get $V_A = V_B$, so in the rest of the proof
%we drop the subscript of $V$.
%By \cref{lm:chsh_comp} and \cref{prop:2d-subspace}, we know $V$ consists of $\omega_d$-eigenvectors and $\omega_d^{-1}$ eigenvectors of 
%$X_A$, so we can write 
%\begin{align}
%	\Pi_{V} = \Pi_{V_1} + \Pi_{V_{d-1}},
%\end{align}
%where $\Pi_{V_{1}}$ is the projector onto the $\omega_d$-eigenspace of $X_A$ and $\Pi_{V_{d-1}}$ is the projector 
%onto the $\omega_d^{-1}$-eigenspace of $X_A$.
%\hl{Here what we need is that $\Pi_{V_1,A} \ket{\psi} = \Pi_{V_{d-1},B} \ket{\psi} = \Pi_{V_1,A}\Pi_{V_{d-1},B}\ket{\psi}$
%and $\Pi_{V_1,B} \ket{\psi} = \Pi_{V_{d-1},A} \ket{\psi} = \Pi_{V_1,B}\Pi_{V_{d-1},A}\ket{\psi}$.
%This is not proved because $\Pi_{V_1}$ and $\Pi_{V_{d-1}}$ are not in the strategy.}
%Suppose $\ket{x_{1}} \in V_{1}$, then $X_A \ket{x_{1}} = \omega_d \ket{x_{1}}$.
%By \cref{eq:ux_relation} we can calculate that
%\begin{align}
%\label{eq:ladder}
% X_AU_A^\dagger \ket{x_{1}} = U_A^\dagger X_A^r \ket{x_{1}} = \omega_d^r U_A^\dagger \ket{x_{1}},
%\end{align}
%so $U_A^\dagger \ket{x_{1}}$ is an eigenvector of $X_A$ with eigenvalue $\omega_d^r$.
%By induction, we know $X_A (U_A^\dagger)^i \ket{x_{1}} = \omega_d^{r^i} (U_A^\dagger)^i\ket{x_{1}}$. 
%From the set $\{(U_A^\dagger)^i \Pi_{V_1} (U_A)^i \}_{i=0}^{d-2}$, we can identify $\Pi_{V_i}$ for $i = 1 \dots  d-1$
%such that $\Pi_{V_i}$ is the projector onto the $\omega_d^i$-eigenspace of $X_A$,
%\hl{Actually, we are not sure about the containment between $\supp((U_A\ct)^{\log_r(d-1)} \Pi_{V_1} U_A^{\log_r(d-1)})$ and 
%$\supp(\Pi_{V_{d-1}})$, so we cannot say that $\Pi_{V_{d-1}} = (U_A\ct)^{\log_r(d-1)} \Pi_{V_1} U_A^{\log_r(d-1)}$.}
%and 
%\begin{align}
% \cup_{i \in [d-1]+1} V_i \subset \supp(\rho_A).
%\end{align}
%Since unitary transformation does not change the rank of a matrix, we know $\rank(\Pi_{V_1}) = \rank(\Pi_{V_{i}}) =N$
%for $ i =1 \dots d-1$.
%We pick a basis for $V_1$ such that 
%\begin{align}
%	\Pi_{V_1} =  \sum_{j=1}^m \ketbra{x_{1,j}}{x_{1,j}},
%\end{align}
%then we can construct 
%\begin{align}
% \Pi_{V_i} = \sum_{j=1}^m \ketbra{x_{i,j}}{x_{i,j}},
%\end{align}
%where $\ket{x_{i,j}} = (U\ct)^{k_i} \ket{x_{1,j}}$ for $r^{k_i} \equiv i \pmod{d}$ and $1 \leq j \leq m$.
%By \cref{eq:ux_relation} we also know that $U_A \ket{x_{i,j}} = \ket{x_{i/2,j}}$ for $i = 1,2 \dots d-1$.
%In order to apply \cref{lm:ux_independ}, we construct $m$ subspaces $\{W_j\}_{j=1}^m$ where 
%\begin{align}
%	W_j = \spn( \{ \ket{x_{i,j}} \}_{i=1}^{d-1} )
%\end{align}
%The subspace $W_j$ is orthogonal to $W_{j'}$ for $j \neq j'$, and
%$U_A$ and $X_A$ satisfy the condition of \cref{lm:ux_independ} when their actions are 
%restricted to each $W_j$.
%Similar argument also applies to operator $X_B$ and $U_B$ on Bob's side.

%With an appropriate change of basis, we 
%assume that $\ket{\psi} = \vc(\tau)$ for some $\tau \in L(\supp(\rho_A))$
%The consistency condition is equivalent to
%\begin{align}
%\label{eq:con_tau}
%	A_i \tau B_i^\dagger = \tau.
%\end{align}
%Substituting $i=1,2$ into \cref{eq:con_tau}, we get 
%\begin{align}
%	X_A \tau X_A^\dagger = A_1A_2\tau B_2^\dagger B_1^\dagger = A_1\tau B_1^\dagger = \tau.
%\end{align}
%Similar argument gives us that 
%\begin{align}
%	U_A \tau U_A^\dagger = \tau.
%\end{align}
%Then we can conclude that for any $k \in \{0,1 \dots d-2\}$ and $l \in \{1,2\dots d-1\}$
%\begin{align}
%	U_A^kX_A^l \tau (U_A^kX_A^l)^\dagger = \tau.
%\end{align}
%Let $\Pi_{W_j}$ be the projector onto $W_j$. 
%By \cref{lm:ux_independ}, $\Pi_{W_j} \tau \Pi_{W_j}$ commutes with all the $(d-1) \times (d-1)$ matrices, which means that 
%\begin{align}
%	\label{eq:d-1}
%	\Pi_{W_j} \tau \Pi_{W_j}  = c_j \1_{W_j} \text{ for } j = 1\dots m.
%\end{align}
%In the vector form, we know
%\begin{align}
%	\sum_{j =1}^N \Pi_{W_j} \ket{\psi} =\sum_{i=1}^{d-1} \sum_{j=1}^m c_j \ket{x_{i,j}}\ket{x_{i,j}}. 
%\end{align}
%Recalling the fact that 
%\begin{align}
%\norm{ \Pi_{V_1} + \Pi_{V_{d-1}} \ket{\psi}}^2 = \norm{\Pi_{V_1} \ket{\psi}}^2 + \norm{\Pi_{V_{d-1}} \ket{\psi}}^2 = \frac{2}{d-1},
% \end{align}
% it means that 
% \begin{align}
% 	2 \sum_{j=1}^m \norm{c_j}^2 = \frac{2}{d-1}.
% \end{align}
% Then we can calculate the norm of $\sum_{j =1}^m \Pi_{W_j} \ket{\psi}$, which is
% \begin{align}
% \norm{\sum_{j =1}^m \Pi_{W_j} \ket{\psi}}^2 = \sum_{i=1}^{d-1} \sum_{j=1}^m \norm{c_j}^2 = 1 = \norm{\ket{\psi}}^2.
% \end{align}
% We can conclude that $\supp{\rho_A} = \cup_{i=1}^{d-1} V_i$ and 
% \begin{align}
% 	\ket{\psi} = \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \ket{x_{i,j}}\ket{x_{i,j}}.
% \end{align}
% 
%The last step is to construct the local isometries $\Phi_A$ and $\Phi_B$ that can produce $\ket{\EPR{d-1}}$ from $\ket{\psi}$.
%To help understanding, we draw a diagram to demonstrate the isometry first and then show
%how this isometry works.
%\begin{figure}[H]
%\center
%        \begin{tikzpicture}[thick]
%        %
%        % `operator' will only be used by Hadamard (H) gates here.
%        % `phase' is used for controlled phase gates (dots).
%        % `surround' is used for the background box.
%        \tikzstyle{operator} = [draw,fill=white,minimum size=1.5em] 
%        \tikzstyle{phase} = [fill,shape=circle,minimum size=5pt,inner sep=0pt]
%        \tikzstyle{surround} = [fill=blue!10,thick,draw=black,rounded corners=2mm]
%        %
%        % Bracket
%        \draw[decorate,decoration={brace,mirror},thick] (0,-1) to
%    	node[midway,left] (bracket1) {$\ket{\psi}$}
%    	(0,-2);
%        % Qubits
%        \node at (0,0) (q1) {$\ket{0}$};
%        \node at (0,-1) (q2) {};
%        \node at (0,-2) (q3) {};
%        \node at (0,-3) (q4) {$\ket{0}$};
%        %
%        % Column 1
%        \node[operator] (op11) at (1,0) {$QFT_d$} edge [-] (q1);
%        \node[operator] (op14) at (1,-3) {$QFT_d$} edge [-] (q4);
%        %
%        % Column 3
%        \node[phase] (phase11) at (2,0) {} edge [-] (op11);
%	\node[operator] (op22) at (2,-1) {$X_A$} edge [-] (q2);
%	\node[operator] (op23) at (2, -2) {$X_B$} edge[-] (q3);
%        \node[phase] (phase14) at (2,-3) {} edge [-] (op14);
%        \draw[-] (phase11) -- (op22);
%        \draw[-] (phase14) -- (op23);
%        %
%        % Column 4
%        \node[operator] (op31) at (3,0) {$QFT_d^{-1}$} edge [-] (phase11);
%        \node[operator] (op34) at (3,-3) {$QFT_d^{-1}$} edge [-] (phase14);
%        %
%        % Column 5
%        \node[phase] (phase21) at (4,0) {} edge [-] (op31);
%	\node[operator] (op42) at (4,-1) {$U_A$} edge [-] (op22);
%	\node[operator] (op43) at (4, -2) {$U_B$} edge[-] (op23);
%        \node[phase] (phase24) at (4,-3) {} edge [-] (op34);
%        \draw[-] (phase21) -- (op42);
%        \draw[-] (phase24) -- (op43);
%        %
%        % Column 6
%        \node (end2) at (5,-1) {} edge [-] (op42);
%        \node (end3) at (5,-2) {} edge [-] (op43);
%        %
%        % Bracket
%        \draw[decorate,decoration={brace},thick] (5,-1) to
%    	node[midway,right] (bracket) {$\ket{junk}$}
%    	(5,-2);
%        %
%        % Column 7
%        \node (end1) at (6.5,-0.5) {} edge [-] (phase21);
%        \node (end4) at (6.5,-2.5) {} edge [-] (phase24);
%        % Dashed line
%        \draw[loosely dotted] (0,-1.5) -- (5,-1.5);
%        % Bracket
%        \draw[decorate,decoration={brace},thick,] (6.5,-0.5) to
%    	node[midway,right] (bracket2) {$\ket{\EPR{d-1}}$}
%    	(6.5,-2.5);
%        %
%        % Background Box
%        \begin{pgfonlayer}{background} 
%        \node[surround] (background) [fit = (q1) (op14) (bracket1)(bracket2)] {};
%        \end{pgfonlayer}
%        %
%        \end{tikzpicture}
%	\caption{The isometries $\Phi_A$ and $\Phi_B$.}
%\end{figure}
%The isometry maps the state $\ket{\psi}$ in the following steps:
%\begin{enumerate}
%	\item Append $\ket{0}_A$ on Alice's side and $\ket{0}_B$ on Bob's side as control registers, and
%	the state becomes $\ket{\psi} \ket{0}_A \ket{0}_B$;
%	\item Apply $QFT_d$ to the control registers, and the state becomes
%	\begin{align}
%		\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} \ket{\psi} \ket{k_1}_A \ket{k_2}_B;
%	\end{align}
%	\item Apply $X_A^{k_1}$ to Alice's share of $\ket{\psi}$ and $X_B^{k_2}$ to Bob's share controlled
%	by the control registers, and the state becomes
%	\begin{align}
%		&\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1}X_A^{k_1} X_B^{k_2}\ket{\psi} \ket{k_1}_A \ket{k_2}_B\\
%		=&\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j X_A^{k_1}\ket{x_{i,j}}_A X_B^{k_2}\ket{x_{i,j}}_B
%		\ket{k_1}_A \ket{k_2}_B\\
%		=&\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^{k_1i+k_2i} \ket{x_{i,j}}_A\ket{x_{i,j}}_B
%		\ket{k_1}_A \ket{k_2}_B;
%	\end{align}
%	\item Apply $QFT_d^{-1}$ to the control registers, and the state becomes
%	\begin{align}
%		&\frac{1}{d^2} \sum_{l_1,l_2 =0}^{d-1}\sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^{k_1(i-l_1)+k_2(i-l_2)} \ket{x_{i,j}}_A\ket{x_{i,j}}_B
%		\ket{l_1}_A \ket{l_2}_B\\
%		=& \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \ket{x_{i,j}}_A\ket{x_{i,j}}_B \ket{i}_A \ket{i}_B;
%	\end{align}
%	\item Let $n_i$ satisfy the condition $r^{n_i} \equiv i \pmod{d}$, apply $U_A^{(i)} = U_A^{n_i}$ to Alice's share of $\ket{\psi}$ and $U_B^{(i)} = U_B^{n_i}$ to Bob's share controlled
%	by the control registers, and the state becomes
%	\begin{align}
%		&\sum_{i=1}^{d-1}\sum_{j=1}^m c_j U_A^{(i)}\ket{x_{i,j}}_A U_B^{(i)}\ket{x_{i,j}}_B \ket{i}_A\ket{i}_B \\
%		=&\sum_{i=1}^{d-1} \sum_{j=1}^m c_j \ket{x_{i r^{-n_i} ,j}}_A \ket{x_{i r^{-n_i},j}}_B \ket{i}_A\ket{i}_B\\
%		=& \sum_{i=1}^{d-1} \sum_{j=1}^m c_j \ket{x_{i i^{-1},j}}_A \ket{x_{i i^{-1},j}}_B \ket{i}_A\ket{i}_B\\
%		= &\left(\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B\right) \x \sum_{i=1}^{d-1} \ket{i}_A\ket{i}_B\\
%		=&\sqrt{d-1} \left(\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B\right) \x 
%		\frac{1}{\sqrt{d-1}}\sum_{i=1}^{d-1}\ket{i}_A\ket{i}_B\\
%		=&\sqrt{d-1} \left(\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B\right) \x \ket{\EPR{d-1}},
%	\end{align}
%	where in the last line we used the fact that $\norm{\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B} = 1/\sqrt{d-1}$.
%\end{enumerate}
%
%If the initial state is $X_A\ket{\psi}$, the isometry maps the state as following
%\begin{align}
%	X_A\ket{\psi} \to &X_A\ket{\psi}\ket{0}_A\ket{0}_B\\
%	\to &\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} X_A\ket{\psi} \ket{k_1}_A \ket{k_2}_B \\
%	\to &\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} X_A^{k_1+1} X_B^{k_2}\ket{\psi} \ket{k_1}_A \ket{k_2}_B \\
%	=&\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^i\omega_d^{(k_1+k_2)i} \ket{x_{i,j}}_A\ket{x_{i,j}}_B
%		\ket{k_1}_A \ket{k_2}_B\\
%	\to &\frac{1}{d^2}\sum_{k_1,k_2 =0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^i\omega_d^{(k_1-l_1)i+(k_2-l_2)i} \ket{x_{i,j}}_A\ket{x_{i,j}}_B
%		\ket{l_1}_A \ket{l_2}_B\\
%	=&\sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^i \ket{x_{i,j}}_A\ket{x_{i,j}}_B \ket{i}_A \ket{i}_B\\
%	\to& \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^i U_A^{(i)}\ket{x_{i,j}}_A U_B^{(i)}\ket{x_{i,j}}_B \ket{i}_A\ket{i}_B\\
%	=&\sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^i \ket{x_{1,j}}_A \ket{x_{1,j}}_B \ket{i}_A\ket{i}_B\\
%	=&\sqrt{d-1} \left(\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B\right) \x 
%		\frac{1}{\sqrt{d-1}} \sum_{i=1}^{d-1}\omega_d^i \ket{i}_A\ket{i}_B\\
%	=& \sqrt{d-1} \left(\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B\right) \x \tX_A \ket{\EPR{d-1}},
%\end{align}
%where $\tX_A$ is the ideal $X$ operator.
%The case of $X_B\ket{\psi}$ is similar, so we omit it here. 
%
%If the initial state is $U_A\ket{\psi}$, the isometry maps the state as following
%\begin{align}
%	U_A\ket{\psi} \to &U_A\ket{\psi}\ket{0}_A\ket{0}_B\\
%	\to &\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} U_A\ket{\psi} \ket{k_1}_A \ket{k_2}_B \\
%	= &\frac{1}{d} \sum_{k_1,k_2=0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j U_A\ket{x_{i,j}}_A\ket{x_{i,j}}_B
%		\ket{k_1}_A \ket{k_2}_B\\
%	= & \frac{1}{d} \sum_{k_1,k_2=0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \ket{x_{i r^{-1} ,j}}_A\ket{x_{i,j}}_B
%		\ket{k_1}_A \ket{k_2}_B\\
%	\to &\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j X_A^{k_1} \ket{x_{i r^{-1} ,j}}_A
%	X_B^{k_2}\ket{x_{i,j}}_B  \ket{k_1}_A \ket{k_2}_B \\
%	=&\frac{1}{d} \sum_{k_1,k_2 =0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^{k_1 i/r}\omega_d^{k_2i} \ket{x_{ir^{-1},j}}_A\ket{x_{i,j}}_B
%		\ket{k_1}_A \ket{k_2}_B\\
%	\to &\frac{1}{d^2}\sum_{k_1,k_2 =0}^{d-1} \sum_{i=1}^{d-1}\sum_{j=1}^m c_j \omega_d^{k_1(i/r-l_1)}\omega_d^{(k_2(i-l_2)} \ket{x_{ir^{-1},j}}_A\ket{x_{i,j}}_B
%		\ket{l_1}_A \ket{l_2}_B\\
%	=&\sum_{i=1}^{d-1}\sum_{j=1}^m c_j  \ket{x_{ir^{-1},j}}_A\ket{x_{i,j}}_B \ket{ir^{-1}}_A \ket{i}_B\\
%	\to& \sum_{i=1}^{d-1}\sum_{j=1}^m c_j U_A^{(ir^{-1})}\ket{x_{ir^{-1},j}}_A U_B^{(i)}\ket{x_{i,j}}_B \ket{i}_A\ket{i}_B\\
%	=&\sum_{i=1}^{d-1}\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B \ket{ir^{-1}}_A\ket{i}_B\\
%	=&\sqrt{d-1} \left(\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B\right) \x 
%		\frac{1}{\sqrt{d-1}} \sum_{i=1}^{d-1}\ket{ir^{-1}}_A\ket{i}_B\\
%	=& \sqrt{d-1} \left(\sum_{j=1}^m c_j \ket{x_{1,j}}_A \ket{x_{1,j}}_B\right) \x \tU_A \ket{\EPR{d-1}},
%\end{align}
%where $\tU_A$ is the ideal $U$ operator on Alice's side.
%The derivation is similar when $U_B$ is applied, so we omit it here.
%\end{proof}
%The effect of $\CHSH_X$ and $\SVT_X$ is that we have a set $\{ \ket{x_i} \}_{i \in [d]} \in \supp(\rho_A)$ such that
%\begin{align}
%	X \ket{x_i} = \omega_d^i \ket{x_i}.
%\end{align}
%The other effect of $\SVT_X$ is that we know 
%\begin{align}
%	\tA_\triangle^0 \rho_A \tA_\triangle^0 = \frac{1}{d} \ketbra{x_0}{x_0} = \frac{1}{d} \tA_\triangle^0.
%\end{align}
%\hl{\textbf{Question}: Can we show $\Tr(\ketbra{x_2}{x_2} \rho_A) = 1/d$?}

%We remark that our proof works for any odd prime number with primitive root $r$.
%However, since there is no upper-bound of $r$ for a general prime number $d$, we 
%cannot claim the corresponding correlation $C(d)$ is constant-sized. On the other
%hand, since $r$ is usually much smaller than $d$, our result implies a more efficient
%way to test EPR pairs of prime local dimension than the one proposed in Ref.~\cite{cgs2017}.
%The size of our correlation grows linearly in $r$ whereas the Coladangelo \textit{et. al.}'s
%method uses correlation grows linearly in $d$.

\section*{Acknowledgment}
The author is deeply grateful to his PhD advisor, Carl Miller, for the enlightening discussions with him, his patient guidance and useful critiques through this
project. His original idea about 
combining a linear system game with a Bell inequality and characterizing prime numbers
by their primitive roots are invaluable to
this work. He also helped the author with the design
of the extended weighted CHSH test
and the editing of the draft.

\bibliographystyle{alphaurl}
\bibliography{quantum_correlation}
\appendix

%===========================================
% \section{Embedding $\Pg$ into $\Gamma$}
% \label{sec:embedding}
%===========================================
% This section supports \cref{sec:embed_idea} by giving the definitions
% of $\Pg_0$ and $\Pg_1$.
% We also prove some properties of the generators of $\Pg_0$, $\Pg_1$ and 
% $\Gamma$ along the way.

% Recall that the goal of introducing $\Pg_0$ is to embed
% $\Pg$ in a group with only order-$2$ generators,
% so that it is easier to embed $\Pg$ into a solution group.

% Next we define the group $\Pg_1$.

% Recall the definition of $\Gamma$ in \cref{def:gamma}. We prove a property of it.

%========================================
\section{The linear system of $\LS(r)$}
\label{sec:lsg_eq}
%========================================
\begin{definition}
    \label{def:lsg_eq}
    The linear system of $\LS(r)$ has 
    variables $x(s)$ for each $s \in S$, and equations:
    \begin{itemize}
        \item for each $1 \leq i \leq r+5$:
        \begin{align*}
            & x(a_i) + x(b_i) + x(c_i) = 0
            && x(a_i) + x(f_0) + x(d_i) = 0 \\
            & x(b_i) + x(f_2) + x(p_{i,1}) = 0 
            && x(p_{i,2}) + x(p_{i,3}) + x(p_{i,4}) = 0 
            && x(f_0) + x(p_{i,3}) + x(p_{i,4}) = 0\\
            & x(c_i) + x(p_{i,4})+x(p_{i,5}) = 0
            && x(f_1) + x(p_{i,2}) + x(p_{i,3}) = 0; 
        \end{align*}
        \item for each $\bc = (i,j,k) \in C(r)$:
        \begin{align*}
            & x(h_{jk}) + x(b_j) + x(c_k) = 0 \\
            & x(d_i) + x(q_{\bc, 1}) + x(f_2) = 0 
            && x(b_j) + x(f_2) + x(q_{\bc, 2}) = 0 
            && x(q_{\bc, 2}) + x(q_{\bc, 3}) + x(q_{\bc, 4}) = 0 \\
            & x(d_i) + x(q_{\bc, 4}) + x(q_{\bc, 5}) = 0 
            && x(c_k) + x(q_{\bc, 5}) + x(q_{\bc, 6}) = 0 
            && x(q_{\bc, 1}) + x(q_{\bc, 3}) + x(q_{\bc, 6}) = 0;
        \end{align*}
        \item Magic Square equations:
        \begin{align*}
            &x(f_0) + x(f_1) + x(f_2) = 0 &&
            x(g_0) + x(g_1) + x(g_2) = 0 &&
            x(m_0) + x(m_1) + x(m_2) = 0 \\
            &x(f_0) + x(g_2) + x(m_0) = 0 &&
            x(f_2) + x(g_0) + x(m_1) = 0 &&
            x(f_1) + x(g_1) + x(m_2) = 1.
        \end{align*} 
    \end{itemize}
\end{definition}
%================================================
\section{A representation of $\Gamma$}
\label{sec:rep_gamma}
%================================================
In this section we give the full description of a representation of $\Gamma$,
$\Psi$,
which is built upon representations 
of $\Pg_0$ and $\Pg_1$. We need the representation $\Psi$ of $\Gamma$ 
to construct the ideal strategy of $\LS(r)$.

Recall the two Hilbert spaces $W_{d-1}$ and $W_2$ defined in \cref{eq:basis}, 
which are the carrier spaces
of the representations given below,
On the Hilbert space $W_2$, we define
\begin{align*}
	X_{W_2} = \ketbra{x_1}{x_2} + \ketbra{x_2}{x_1} &&
	Y_{W_2} = i\ketbra{x_1}{x_2} - i \ketbra{x_2}{x_1} &&
	Z_{W_2} = \ketbra{x_1}{x_1} - \ketbra{x_2}{x_2},
\end{align*}
\begin{definition}
\label{def:rep_g0}
The representation, $\Psi_0$, of $\Pg_0$ on $W_{d-1}$ is defined by
\begin{align*}
	\Psi_0(a_1) =&\sum_{j=1}^{(d-1)/2} \omega_d^j \ketbra{x_j}{x_{d-j}} + \omega_d^{-j} \ketbra{x_{d-j}}{x_{j}} \\
	\Psi_0(a_2) = &\sum_{j=1}^{d-1} \ketbra{x_j}{x_{d-j}}\\
	\Psi_0(a_3) = &\ketbra{u_0}{u_0} +\omega_{d-1}^{(d-1)/2}\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}}\\ + 
	&\sum_{k=1}^{(d-3)/2}\left( \omega_{d-1}^k\ketbra{u_k}{u_{d-1-k}} + \omega_{d-1}^{-k}\ketbra{x_{d-1-k}}{x_k}\right)\\ 
	\Psi_0(a_4) = &\ketbra{u_0}{u_0} +\ketbra{u_{(d-1)/2}}{u_{(d-1)/2}} \\+
	 &\sum_{k=1}^{(d-3)/2}\left(\ketbra{u_{d-1-k}}{u_k} + \ketbra{u_k}{u_{d-1-k}}\right),
\end{align*}
The representation of the rest of the generators
can be constructed from $\Psi_0(a_1)$, $\Psi_0(a_2)$, $\Psi_0(a_3)$ and $\Psi_0(a_4)$ following the defining 
conjugacy relations in $\Pg_0$.
\end{definition}
We can check that
\begin{align*}
	&\Psi_0(a_1a_2) =  \sum_{j=0}^{d-2} \omega_d^{r^j} \ketbra{x_{r^j}}{x_{r^j}} =
	\tO \\
	&\Psi_0(a_3a_4) =  \sum_{j=0}^{d-2} \ketbra{x_{r^{j-1}}}{x_{r^j}} = \tU,
\end{align*}
where $\tO$ and $\tU$ are defined in \cref{eq:O,eq:U}.
We can also check that the representation $\Psi_0$ satisfies the relations
that $\Psi_0(a_3a_4) \Psi_0(a_2) \Psi_0(a_4a_3) = \Psi_0(a_2)$ by first observing that
	\begin{align*}
		\Psi_0(a_3)\Psi_0(a_4)= \sum_{j=0}^{d-2} \ketbra{x_{r^{j-1}}}{x_{r^j}}.
	\end{align*}
Then we can calculate  
	\begin{align*}
		&\Psi_0(a_3)\Psi_0(a_4) \Psi_0(a_2) \Psi_0(a_4) \Psi_0(a_3)\\
		= &\sum_{j=0}^{(d-3)/2} \Psi_0(a_3)\Psi_0(a_4) (\ketbra{x_{r^j}}{x_{r^{d-1-j}}}+\ketbra{x_{r^{d-1-j}}}{x_{r^j}}) \Psi_0(a_4)\Psi_0(a_3)\\
		=& \sum_{j=0}^{(d-3)/2} \ketbra{x_{r^{j-1}}}{x_{r^{d-2-j}}}+\ketbra{x_{r^{d-2-j}}}{x_{r^{j-1}}}\\ 
		=& \Psi_0(a_2).
	\end{align*}

Next, following the proof of Lemma $29$ of \cite{slofstra2017}, we construct a representation of $\Pg_1$.
\begin{definition}
\label{def:rep_g1}
The representation $\Psi_1$ of $\Pg_1$ on $W_2 \x W_{d-1}$ is defined by
\begin{itemize}
\item for $f_0$
\begin{align*}
	&\Psi_1(f_0) = X_{W_2} \x \1_{W_{d-1}};
\end{align*}
\item
for $i = 1 \dots r+5$,
\begin{align*}
	&\Psi_1(a_i) = \1_{W_2} \x \Psi_0(a_i) \\
	&\Psi_1(b_i) = \ketbra{x_1}{x_1} \x \Psi_0(a_i) + \ketbra{x_2}{x_2} \x \1_{W_{d-1}} \\
	&\Psi_1(c_i) =\ketbra{x_1}{x_1} \x \1_{W_{d-1}} + \ketbra{x_2}{x_2} \x \Psi_0(a_i) \\
	&\Psi_1(d_i) =  X_{W_2} \x \Psi_0(a_i);
\end{align*}
\item for $(i,j,k) \in C(r)$,
\begin{align*}
	\Psi_1(h_{jk}) = \ketbra{x_1}{x_1} \x \Psi_0(a_j) + \ketbra{x_2}{x_2} \x \Psi_0(a_k).
\end{align*}
\end{itemize}
\end{definition}

Finally, we follow the proof of Proposition $27$ of \cite{slofstra2017} to construct the 
representation $\Psi$ of $\Gamma$.
\begin{definition}
\label{def:rep_gamma}
The representation $\Psi$ of $\Gamma$ on $W_2 \x W_2 \x W_{d-1}$ is defined by:
\begin{itemize}
\item for $s \in \{a_i, b_i, c_i, d_i\}_{i=1}^{r+5}$,
\begin{align*}
	\Psi(s) = \1_{W_2} \x \Psi_1(s);
\end{align*}
\item for $f_i$, $g_i$, $m_i$ with $i = 0,1,2$ and $J$
\begin{align*}
    &\Psi(f_0) = \1 \x X_{W_2} \x \1_{W_{d-1}} \\
	&\Psi(f_1) = X_{W_2} \x X_{W_2} \x \1_{W_{d-1}} \\
	&\Psi(f_2) = X_{W_2} \x \1_{W_2} \x \1_{W_{d-1}} \\
	&\Psi(g_0) = \1 \x Z_{W_2} \x \1_{W_{d-1}} \\
	&\Psi(g_1) = Z_{W_2} \x Z_{W_2} \x \1_{W_{d-1}} \\
	&\Psi(g_2) = Z_{W_2} \x \1_{W_2} \x \1_{W_{d-1}} \\
	& \Psi(m_0) = Z_{W_2} \x X_{W_2} \x \1_{W_{d-1}}\\
	& \Psi(m_1) = X_{W_2} \x Z_{W_2} \x \1_{W_{d-1}}\\
	& \Psi(m_2) = Y_{W_2} \x Y_{W_2} \x \1_{W_{d-1}}\\
	& \Psi(J) = - \1_{W_2} \x \1_{W_2} \x \1_{W_{d-1}};
\end{align*}
\item for the generators involved in $R_i$
with $1 \leq i \leq r+5$ :
\begin{align*}
	\Psi(p_{i,1}) &= X_{W_2} \x \Psi_1(b_i) \\
	\Psi(p_{i,2}) &= \ketbra{x_1}{x_2} \x \Psi_1(b_if_0) + \ketbra{x_2}{x_1} \x \Psi_1(f_0b_i)\\
	\Psi(p_{i,3}) &= \ketbra{x_1}{x_1} \x \Psi_1(b_if_0b_i) + \ketbra{x_2}{x_2} \x \Psi_1(f_0) \\
	\Psi(p_{i,4}) &= \ketbra{x_1}{x_1} \x \Psi_1(b_ic_i) + \ketbra{x_2}{x_2} \x \1_{W_2} \x \1_{W_{d-1}}\\
    \Psi(p_{i,5}) &= \ketbra{x_1}{x_1} \x \Psi_1(b_i) + \ketbra{x_2}{x_2} \x \Psi_1(c_i);
\end{align*}
\item for the generators involved in $R_{\bc}$ with $\bc = (i,j,k) \in C(r)$:
\begin{align*}
	\Psi(q_{\bc, 1}) &= X_{W_2} \x \Psi_1(d_i) \\
	\Psi(q_{\bc, 2}) &= X_{W_2} \x \Psi_1(b_j) \\
	\Psi(q_{\bc, 3}) &= \ketbra{x_1}{x_2} \x  \Psi_1(b_jd_i) + \ketbra{x_2}{x_1} \x \Psi_1(d_ib_j)\\
	\Psi(q_{\bc, 4}) &= \ketbra{x_1}{x_1} \x \Psi_1(b_jd_ib_j) + \ketbra{x_2}{x_2} \x  \Psi_1(d_i)\\
	\Psi(q_{\bc, 5}) &=  \ketbra{x_1}{x_1} \x \Psi_1(b_jc_k) + \ketbra{x_2}{x_2} \x \1_{W_2} \x \1_{W_{d-1}}\\
	\Psi(q_{\bc, 6}) &= \ketbra{x_1}{x_1} \x \Psi_1(b_j) + \ketbra{x_2}{x_2} \x \Psi_1(c_{k}).
\end{align*}
\end{itemize}
\end{definition}
The validation of $\Psi_1$ and $\Psi$ were done in the proofs of 
Lemma $29$ and Proposition $27$ in \cite{slofstra2017}.
%========================================
\section{The proof of \cref{thm:selftest} }
\label{sec:w_chsh_proof}
%========================================
This proof follows the same line of argument in Appendix A of \cite{bamps2015}.
We first find two sum-of-square decompositions of $2\sqrt{\alpha^2+1} \1 - \I_\alpha$,
where $\I_\alpha$ is expressed in terms of $\{M_x\}$ and $\{N_y\}$.
The decompositions allow us to determine some key relations between Alice and Bob's observables
and their shared state, which will be used to draw the conclusion.

\begin{proof}
The first step is to find a sum-of-square decomposition of 
the following Bell expression
\begin{align}
	\bar{\I}_\alpha = 2\sqrt{\alpha^2+1} \1 - \I_\alpha
	= \frac{2}{\sin(\mu)} \1 - \frac{\cos(\mu)}{\sin(\mu)}(M_1N_1+M_1N_2) -  M_2N_1 + M_2N_2.
\end{align} 
With the notation $c:= \cos(\mu)$, $s := \sin(\mu)$, 
we define
\begin{align*}
	&Z_A := M_1 && X_A := M_2\\
	&Z_B := \frac{N_1+N_2}{2c} && X_B := \frac{N_1-N_2}{2s}.
\end{align*}
The two SOS decompositions that we use are
\begin{align}
	\label{eq:sos1}&\bar{\I}_\alpha = \frac{s \bar{\I}_\alpha^2 + 4sc^2(Z_AX_B+X_AZ_B)^2}{4},\\
	\label{eq:sos2}&\bar{\I}_\alpha = \frac{c^2}{s}(Z_A-Z_B)^2 + s(X_A-X_B)^2.
\end{align}
The verification is omitted here. From the SOS decomposition, we can prove \cref{eq:za-zb,eq:xa-xb,eq:xazb,eq:zaxb,eq:zaxa,eq:zaxaxbzb}.

We first define
\begin{align*}
	&S_1 = \frac{\sqrt{s}}{2} \bar{\I}_\alpha, &&
	S_2 = \sqrt{s}c(Z_AX_B+ X_AZ_B),\\
	&S_3 = \frac{c}{\sqrt{s}}(Z_A-Z_B), &&
	S_4 = \sqrt{s}(X_A-X_B)
\end{align*}
then 
\begin{align}
\bar{\I}_\alpha = S_1^2 + S_2^2 = S_3^2 + S_4^2.
\end{align}
The fact that the quantum strategy $(\ket{\psi}, \{M_x\}_{x=1,2}, \{N_{y }\}_{y=1,2}$ achieves that 
$\ip{\bar{\I}_\alpha} \leq \epsilon$ implies that
\begin{align*}
    \bra{\psi} S_i^2 \ket{\psi} \leq \ep,
\end{align*}
and equivalently,
\begin{align*}
   \norm{S_i \ket{\psi}} \leq \se 
\end{align*}
for $i = 1,2,3,4$.
From the definitions of $S_i$'s, we can get that 
\begin{align}
	\label{eq:base1} &\norm{(X_AZ_B+X_BZ_A)\ket{\psi}} \leq \frac{1}{c\sqrt{s}} \se\\
	\label{eq:base2} &\norm{(Z_A-Z_B)\ket{\psi}} \leq \frac{\sqrt{s}}{c} \se\\
	\label{eq:base3} &\norm{(X_A-X_B)\ket{\psi}} \leq \frac{1}{\sqrt{s}} \se,
\end{align}
which proves \cref{eq:za-zb,eq:xa-xb}.
\Cref{eq:base1,eq:base2} give us that 
\begin{align*}
	\norm{ [Z_A(\1+X_B) - (\1-X_A)Z_B] \ket{\psi} } 
	\leq \norm{(X_AZ_B+X_BZ_A)\ket{\psi}} + \norm{(Z_A-Z_B)\ket{\psi}}
	\leq \frac{s+1}{c\sqrt{s}} \se,
\end{align*}
which proves \cref{eq:xazb}.
Similarly, \cref{eq:base1,eq:base3} give us that 
\begin{align*}
	\norm{[X_A(\1+Z_B) - X_B(\1-Z_A)] \ket{\psi}} \leq \frac{c+1}{c\sqrt{s}} \se,
\end{align*}
which proves \cref{eq:zaxb}.
Since $Z_AX_A +X_A Z_A = \frac{S_2}{c\sqrt{s}} + \frac{\sqrt{s}\tilde{X}_AS_3}{c} + \frac{\tilde{Z}_AS_4}{\sqrt{s}}$, we can deduce that
\begin{align*}
	\norm{ (Z_A X_A+X_AZ_A)\ket{\psi}} \leq \frac{1+c+s}{c\sqrt{s}} \se,
\end{align*}
as in \cref{eq:zaxa}.
To prove \cref{eq:zaxaxbzb}, we switch to the approximate relation form and derive that 
\begin{align}
	X_AZ_A \ket{\psi} &\appd{\frac{1+c+s}{c\sqrt{s}} \se}  -Z_AX_A \ket{\psi} \\
	&\appd{\frac{\sqrt{s}}{c} \se} -Z_AX_B \ket{\psi} \\
	&\appd{\frac{1}{s^{3/2} }\se} -X_BZ_B\ket{\psi},
\end{align}
where in the last line we use the fact $\norm{X_B} \leq 1/s$.

Now we introduce the isometries $\Phi_A$ and $\Phi_B$ mentioned in the theorem.
They are the same as the ones used in \cite{bamps2015}.
To construct $\Phi_A$ and $\Phi_B$ we need to regularize $Z_B$ and $X_B$ to make sure the corresponding operations 
are unitary in the isometries. We define $Z_B^\ast$ to be the operator obtained from $Z_B$ by changing all the $0$-eigenvalues
to $1$, then we define
\begin{align*}
Z_B' := Z_B^\ast |Z_B^\ast|^{-1},
\end{align*}
where $|Z_B^\ast|$ is obtained from $Z_B^\ast$ by replacing all negative eigenvalues by its absolute value.
In a similar way, we define $X_B^\ast$ and $X_B'$.
On Alice's side, since $Z_A$ and $X_A$ are unitaries already,
we define
$Z_A' := Z_A$ and $X_A' = X_A$.
The isometries are illustrated in the figure below.
\begin{figure}[H]
\center
        \begin{tikzpicture}[thick]
        %
        % `operator' will only be used by Hadamard (H) gates here.
        % `phase' is used for controlled phase gates (dots).
        % `surround' is used for the background box.
        \tikzstyle{operator} = [draw,fill=white,minimum size=1.5em] 
        \tikzstyle{phase} = [fill,shape=circle,minimum size=5pt,inner sep=0pt]
        \tikzstyle{surround} = [fill=blue!10,thick,draw=black,rounded corners=2mm]
        %
        % Bracket
        \draw[decorate,decoration={brace,mirror},thick] (0,-1) to
    	node[midway,left] (bracket1) {$\ket{\tpsi}$}
    	(0,-2);
        % Qubits
        \node at (0,0) (q1) {$\ket{0}$};
        \node at (0,-1) (q2) {};
        \node at (0,-2) (q3) {};
        \node at (0,-3) (q4) {$\ket{0}$};
        %
        % Column 1
        \node[operator] (op11) at (1,0) {H} edge [-] (q1);
        \node[operator] (op14) at (1,-3) {H} edge [-] (q4);
        %
        % Column 3
        \node[phase] (phase11) at (2,0) {} edge [-] (op11);
	\node[operator] (op22) at (2,-1) {$Z_A'$} edge [-] (q2);
	\node[operator] (op23) at (2, -2) {$Z_B'$} edge[-] (q3);
        \node[phase] (phase14) at (2,-3) {} edge [-] (op14);
        \draw[-] (phase11) -- (op22);
        \draw[-] (phase14) -- (op23);
        %
        % Column 4
        \node[operator] (op31) at (3,0) {H} edge [-] (phase11);
        \node[operator] (op34) at (3,-3) {H} edge [-] (phase14);
        %
        % Column 5
        \node[phase] (phase21) at (4,0) {} edge [-] (op31);
	\node[operator] (op42) at (4,-1) {$X_A'$} edge [-] (op22);
	\node[operator] (op43) at (4, -2) {$X_B'$} edge[-] (op23);
        \node[phase] (phase24) at (4,-3) {} edge [-] (op34);
        \draw[-] (phase21) -- (op42);
        \draw[-] (phase24) -- (op43);
        %
        % Column 6
        \node (end2) at (5,-1) {} edge [-] (op42);
        \node (end3) at (5,-2) {} edge [-] (op43);
        %
        % Bracket
        \draw[decorate,decoration={brace},thick] (5,-1) to
    	node[midway,right] (bracket) {$\ket{aux}$}
    	(5,-2);
        %
        % Column 7
        \node (end1) at (6.5,-0.5) {} edge [-] (phase21);
        \node (end4) at (6.5,-2.5) {} edge [-] (phase24);
        % Dashed line
        \draw[loosely dotted] (0,-1.5) -- (5,-1.5);
        % Bracket
        \draw[decorate,decoration={brace},thick,] (6.5,-0.5) to
    	node[midway,right] (bracket2) {$\frac{\ket{00}+\ket{11}}{\sqrt{2}}$}
    	(6.5,-2.5);
        %
        % Background Box
        \begin{pgfonlayer}{background} 
        \node[surround] (background) [fit = (q1) (op14) (bracket1)(bracket2)] {};
        \end{pgfonlayer}
        %
        \end{tikzpicture}
	\caption{The isometries $\Phi_A$ and $\Phi_B$.}
\end{figure}
To bound $e_{xy} := \norm{ (\Phi_A\x\Phi_B) (\tM_x \x \tN_y) \ket{\tpsi} - \ket{junk} \x (M_x\x N_y) \ket{\EPR{2}}}$,
there are some intermediate steps. Since the derivations are the same as in \cite{bamps2015}, 
we only record the key relations here.
\begin{align*}
	&\norm{(Z_B' - Z_B) \ket{\tpsi}} \leq \frac{\sqrt{s}}{c} \sqrt{\epsilon},\\
	&\norm{(Z_B' - Z_A') \ket{\tpsi}} \leq 2 \frac{\sqrt{s}}{c}\sqrt{\epsilon}, \\
	&\norm{(X_B' - X_B) \ket{\tpsi}} \leq \frac{c+1}{s^{3/2}} \sqrt{\epsilon} := \delta_1 \sqrt{\epsilon},\\
	&\norm{(X_B'Z_B'+ Z_B'X_B')\ket{\tpsi}} \leq [\frac{2\sqrt{s}}{c} + \frac{2}{\sqrt{s}} + 2\delta_1 + 
	(\sqrt{2}+\frac{1}{c})(2\frac{\sqrt{s}}{c}+\frac{1+c+s}{c\sqrt{s}})]\sqrt{\epsilon}
	:= \delta_2 \sqrt{\epsilon}.
\end{align*}
Then we can calculate that 
\begin{align*}
	&e_{00} = e_{10} = 2\delta_2 \sqrt{\epsilon}\\
	&e_{20} = 2(\frac{1+c+s}{c\sqrt{s}} + \delta_2) \sqrt{\epsilon}\\
	&e_{01} = e_{02} = e_{11} = e_{12} = [\sqrt{s}+s(2\frac{1+c+s}{c\sqrt{s}} + \delta_1) + 2\delta_2]\sqrt{\epsilon}\\
	&e_{21} = e_{22} = [2\frac{1+c+s}{c\sqrt{s}} + \sqrt{s}+s(2\frac{1+c+s}{c\sqrt{s}} + \delta_1) + 2\delta_2]\sqrt{\epsilon},
\end{align*}
so an upper bound of the error is that
\begin{align}
 e_{xy} \in O((\frac{1}{c^2s^{1/2}} + \frac{1}{s^{3/2}})\sqrt{\epsilon})
 \quad\quad 
 \forall x,y \in \{0,1,2\}.
\end{align}
\end{proof}
\end{document}